### 2022-12-26 manuscript/main.tex
wrote more

\documentclass[12pt, letterpaper, ]{article}
\usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
\addbibresource{total_biblio.bib}
\addbibresource{datasets.bib}
\usepackage{caption}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{url}
\graphicspath{./images/}
\geometry{
        letterpaper,
        top = 1.27cm,
        bottom = 1.27cm,
        left = 2.54cm,
        right = 2.54cm,
        }
\title{MSc Thesis}
\author{Isaac Peetoom Heida}
\date{December 2022}
\doublespace

\begin{document}

\input{title.tex}

\tableofcontents

\section{Abstract}

As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying. New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains. One of the promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance. With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions. Our ability to mobilize silicon-based cropping strategies is dependent on a thorough understanding of the ultimate and proximate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake. In this thesis, I attempt to extend recent advances in our understanding of silicon ecology in cereal crops, taking an integrative approach towards untangling silicon accumulation in cereal crops. I used a genome-wide association study to identify genetic markers associated with high silicon content, and performed a greenhouse experiment to test for patterns of rapid silicon accumulation in four cereal crops.

[GWAS results, Rapid Si results]

\section*{Lay Summary}

Silicon, a naturally occurring element, provides tremendous benefits to plant health, but is not widely utilized in agriculture. One of the main factors limiting it's application in agriculture is our poor understanding of the exact dynamics of how plants absorb and use silicon from the soil. I identified genetic traits that are associated with high silicon content in a relative of bread wheat, as well as demonstrated that cereal crops (e.g. wheat, barley, oats) have the ability to rapidly uptake silicon from the soil. This rapid uptake means that silicon may be a highly effective defence against insect pests. Combining this results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content. This development could provide an environmentally friendly strategy to maintain output to feed a growing human population.

\section*{Preface}
(This was taken nearly verbatim from Matt's thesis, so need to go back through to make sure I am not plagarizing)

The research presented in this thesis is original and unpublished. Isaac Peetoom Heida and Dr. Juli Carrillo conceptualized and developed the experiment presented in Chapter 1. Isaac Peetoom Heida, Dr. Juli Carrillo and Dr. Gurcharn Singh Brar conceptualized and developed the experiement presented in Chapter 2.
Isaac Peetoom Heida developed the question and methodology for Chapter 1. Dr. Aaron Beattie, Dr. Mazen Aljarrah, and Dr. Gurcharn Brar provided seeds for the experiment. Isaac Peetoom Heida designed and set up the experiment, processed and analysed the samples, and performed the statistical analysis. Dr. Shaun Barker and the Mineral Deposit Research Unit of the University of British Columbia provided facilities and expertise for the XRF analysis of the tissue samples. (but I paid them, do I still thank?)
For chapter 2, Isaac Peetoom Heida and Dr. Gurcharn Brar designed the experiment. Isaac Peetoom Heida let plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda. Isaac Peetoom Heida led the sample harvest, processing and analysis. Kelly Wang, Grace Wang, and Chelsea Gowton assisted with the sample harvesting. Dr. Daria [last name], Paul Fisher, Lucas Friesen, Katie Pryer, Dr. Kinga Treder, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation. 
Chapters two and three of this thesis will be submitted to peer-reviewed journals for publication. For the purposes of this manuscript, actions are depicted in the singular first person. 
\section{Introduction}

\subsection{Ag good, climate change bad, silicon helps}

As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying. New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains. One of the promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance. With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.

\subsection{Silicon in nature}

Silicon is super-abundant in the earth’s crust, with Silicon dioxide (SiO2) comprising about 60\% of the crust by mass. Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant. When supplemented with silicon, plants are generally more resistant to stress. Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). Silicon is deposited throughout the plant body, forming amorphous masses of silica, with highly variable geometries (\cite{piperno_phytoliths_2006}). Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}). Recent evidence points to interactions with gene expression regulators. Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}). The resulting model from this research posits that much of silicon’s beneficial role arises from phytolith interactions with compounds adverse to plant growth, limiting the penetration of the compounds into cells where damage can be caused. For herbivore defence, phytoliths serve a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

\subsection{Silicon in soils and roots}

Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems. Within the soil environment, silicon commonly exists in both crystalline (geologic) forms as well as amorphous (biogenic) forms (\cite{haynes_contemporary_2014}). Amorphous silicates can derive from previous plant material that has decayed in the soil, but also from marine and aquatic organisms such as diatoms. Globally, the silicon cycle has silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments. The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the plant silicon being recycled from previous plant material (\cite{de_tombeur_plants_2020}). In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients for uptake. Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root (\cite{de_tombeur_silicon_2021-1}). This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation my improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).

\subsection{Silicon transporters}

One of the most important advances in plant-silicon research was understanding the mechanisms through which silicon is acquired and transported into the plant. Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}). While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families. Research in rice has identified four proteins that transport silicon into and through the plant body. Two of these proteins (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}). Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}). The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice \cite{yamaji_transporter_2008}. Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate \cite{christian_breeding_2022}. However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant. Further investigations into how genotypic variation is reflected in the silicon content of plants can aid in the discovery of new genes involved in silicon accumulation, and may provide targets for silicon breeding programs.

\subsection{Silicon in leaves}

Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}). Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure for these bodies to yield certain functions to the plant (\cite{piperno_phytoliths_2006}). In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}). This trait has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation reduces the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}). In the leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf. This toughness, and the hardness of the phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics like stem phytoliths (\cite{stromberg_functions_2016}). Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}). As phytoliths deposit in the leaves, polymerization is aided by interactions with proteins in the cell wall, which control sites of nucleation (\cite{nawaz_phytolith_2019}). Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (Waterman et al. 2021b). Thus, silicon-based defences in crop plants may be one of the first lines of active defence, providing rapid and sensitive responses to herbivory.

\subsection{Looking back and looking forward}

Much of today’s plant silicon work is indebted to the pioneering work (\cite{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by (\cite{takahashi_possibility_1990}). Epstein’s seminal (\cite{epstein_silicon_1999}) paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers extending the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world. Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade. Rice, maize, wheat, and barley alone account for 1/3 of the worlds’ total cultivated landmass (\cite{faostat}), and are all domesticated grass species. Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich substrates (\cite{haynes_contemporary_2014,meena_case_2014}). Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{} Fraysse et al. 2009). Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate production systems, particularly in wheat and barley. This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficeincy in which these temperate crops utilize the ample silicon available in their soils. Our ability to integrate silicon as a tool for production improvement in dry-land grain production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which dry-land cereals utilize silicon in pest-protection.

\section{Chapter 1: Identifying rapid silicon accumulation in cereal crops}

During times of crisis, having an effective and timely response can limit damage and speed recovery. To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant. Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant. In most scenarios, induced plant defences are activated in response to an external cue, and build in intensity over time, with defensive hormones peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}). Despite this rapid hormonal response, actual defensive compounds are slower to accumulate, taking (need to figure out if this is true). Many defensive responses are also context dependent, where the identity of the damaging actor, the severity of damage, and a host of other factors interact to determine the final defensive response. The most effective defensive strategies are those that can either prevent herbivory outright, or can mount a rapid response to limit damage. These same strategies are also the most promising for crop production, where pest damage represents both an economic and food security cost. Integrating better natural plant defences into crop production systems may be key to reducing the environmental impact of agriculture, but hinges upon a thorough understanding of plant defensive physiology.

One of the most promising avenues for new crop defence is the harnessing of silicon. Silicon acts on multiple temporal and physiological scales, delivering broad spectrum resistance to pests, pathogens, and abiotic stressors. Soluble silicon taken up from the soil is deposited predominantly in the leaf epidermis, where it forms solid granules that increase the toughness of the tissue, reducing herbivore digestive efficiency. Plant silicon is expressed latently, but also increases in response to herbivory. Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even preempting some chemical defences (\cite{waterman_short-term_2021}). This rapid action makes silicon accumulation a promising trait for future crop development. Despite the novel results, this pattern has so far been observed in just one species, and only under artificial herbivory via the application of methyl-jasmonate. Though a useful tool for herbivory research, methyl-jasmonate application fails to reproduce a complete herbivory signal for the plant and thus observed changes to plant defence may not be representative of a true herbivory scenario (\cite{strauss_direct_2002}). Testing for this rapid silicon accumulation across a variety of grain crops, and under both simulated (methyl-jasmonate) and real herbivory is a crucial first step towards integrating rapid silicification into our understanding of plant defence and crop protection.

Plant silicon research has mostly focused on members of the grass family (Poaceae) due to their exceptional silicon content within the plant kingdom, as well as the economic importance of domesticated grass species. Domesticated crops differ significantly from their wild relatives, due to effects of strong selective pressure imposed by humans (\cite{chen_crop_2015}). Most domesticated crops show much lower genetic diversity than their wild ancestors (\cite{hafeez_creation_2021, smith_domestication_2019}). Initial selection for a few individuals with favourable traits creates a genetic bottleneck, and the majority of allelic diversity is lost. Subsequent selection by humans for agronomically relevant traits can result in concurrent losses of adaptations to natural environments, as the traits that maximize human value (eg. yield, ease of harvest) can come at the cost of ecologically relevant traits such as defence (\cite{whitehead_domestication_2017, chen_crop_2015}). Indeed, in the context of silicon, we can detect clear signals of domestication across the Poaceae family, where wild ancestors consistently have higher baseline silicon content than their domesticated descendants (\cite{simpson_still_2017}). Due to the effects of selection on plant defence it becomes crucial to test new developments in the silicon-defence literature in modern crop species, both to validate their utility towards agricultural production, and in this case to gather further observations on the dynamics of silicon-based defences in the first hours after herbivory.

In this study, I test four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory. In a glasshouse environment, I grew bread wheat (\textit{Triticum aestivum}), oats (\textit{Avena sativa}), barley (\textit{Hordeum vulgare}) and Triticale ($\times$ \textit{Triticosecale}), and tested the following hypotheses:
\begin{enumerate}
        \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species silicon content would show responses to herbivory consistent with this process.
        \item Due to different phylogeny and domestication history, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
        \item Due to the different cues involved when comparing true herbivory damage and methyl-jasmonate induced defensive induction, the tested species would show different patterns of short-term silicon accumulation in response to cricket (\textit{Acheta domesticus}) herbivory and methyl-jasmonate application. 
\end{enumerate}
This study is a thematic replication of Waterman et al.’s 2021 paper (\cite{waterman_short-term_2021}), but attempts to extend the findings to commercially important grain crops. The findings of this study will refine our understanding of the prevalence of rapid silicification in the Poaceae, and will help to inform the value of potential applications of silicon-based defences into grain crops.

\subsection{Methods}

\subsubsection{Plant Growth and Treatments}

To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley. Cultivars were selected on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, I germinated seeds in germination trays filled with moist sand. After two days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid. Though potting mix and the water used contain some amount of plant available silicon, the silicic acid was added to ensure that there would be no silicon limitation to the plants. The pots were bottom watered on flood tables with nutrient solution. Each plant was assigned to one of three herbivory treatments: control, simulated herbivory, or true herbivory. Simulated herbivory was achieved by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while true herbivory was provided by crickets housed in water-pik tubes. Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial. Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.

\subsubsection{Sample Harvest and Preparation}

24 hours after treatment application, I harvested leaf material by clipping the above-ground portion of the plants. Leaf tissue was flash frozen in liquid nitrogen and freeze dried, and stored in a -80ºC freezer. After transferring to 2mL microcentrifuge tubes, I flash froze the leaf tissue and ground for 30 seconds in a tissuelyser bead mill.

\subsubsection{Silicon analysis}

To measure the silicon content of the leaf tissue, I followed a modified version of the benchtop XRF method (\cite{reidinger_rapid_2012}). I pressed ground leaf tissue in a hydraulic press, using a 13mm die at 11 tons of pressure. I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon.

\subsubsection{Statistical Analysis}

To answer all three of our questions, I used a bayesian hierarchical model. [Insert text about how great these models are]. I specified a hierarchical model using the following model:

\[y_i \sim Normal(\hat{y}_i, \sigma)\]
\[\hat{y_i} = \alpha + \alpha_{[i]cultivar} + \beta_{[j]induction} + \beta_{[k]species} + \beta_{[jk]species \times induction}\]

I ran the model using \verb|Turing.jl| in Julia [version number] [cite]. Using 4 chains and 1000 sample iterations, I sampled the posterior distribution using a No U-turn Sampler with 1000 warm up iterations and a target acceptance rate of 0.65 (Hoffman and Gelman 2014). I tested our model structure on simulated data, to ensure it returned accurate parameter estimates. I used the Gelman-Rubin statistic (\( \hat{R} \)) (Gelman and Rubin 1992) and effective sample size to diagnose the convergence of our chains. I verified model fits using posterior predictive checks in \verb|ArviZ.jl| [cite]

\subsection{Results}

Among our cultivars, silicon content ranged from [x]\% to [y]\% of dry mass. [Species a] had the highest amount of silicon at [z]\%, while [species b] had the lowest silicon content at [z]\%. Overall, methyl-jasmonate and insect treatments increased plant silicon by [x] and [y]\% respectively. Our model showed that both species and treatment type had effects on the plant silicon content. Parameter estimates and 90\% credible intervals are summarized in Table \ref{Tab:params}. Species and induction treatments had an interaction.

\subsection{Discussion}

\subsection{Acknowledgements}

\subsection{Data Availability}

\subsection{Figures and Tables}

\begin{table}[h]
        \centering
        \caption{Credible Interval and Parameter estimates for the hierarchical model. Parameters are estimated against a baseline of Induction: None and Species: Barley.}
        \label{Tab:params}
        \begin{tabular} { c | c | c }
                \hline
                Parameter & Credible Interval & Parameter Estimate \\
                \hline
                Insect & -4.200 -- -4.100 & -4.108 \\
                Methyl Jasmonate & -4.20 -- -4.15 & -4.17 \\
                Oats & -5.00 -- -4.95 & -4.97 \\
                Wheat & -4.60 -- -4.5 & -4.51 \\
                Triticale & -4.13 -- -4.08 & -4.11 \\
                \hline
        \end{tabular}
\end{table}

\begin{figure}[h]
        \includegraphics[width = \textwidth]{images/induction_plot.png}
        \centering
        \caption{The effects of crop species and induction treatment on leaf silicon content. Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets \textit{Acheta domesticus}. Leaves were sampled 24 hours after treatment, and were analyzed using XRF.}
\end{figure}

\clearpage

\section{Chapter 2: Genetic drivers of silicon accumulation in a wild ancestor of wheat}
\subsection{Introduction}
With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production. Over the past thirty years, research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops. Cereal crops are globally important, covering over 1/3 of the world’s arable land, making up over 50\% of the daily caloric intake for most people. Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (>0.75\% total dry weight). Silicon is highly abundant in many soils globally, and is the second most abundant element in the earth’s crust, behind only oxygen. It’s high expression in cereals, high abundance in many soils, and incredible broad spectrum effects on plant vigor and stress tolerance have make it a tantalizing target for improvements in agricultural yield and sustainability. 
Silicon is now considered a quasi-essential nutrient for plant growth. Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.

Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress. For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms. On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress. In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants. A current hypothesis explaining the broad-spectrum activity of silicon is presented in {CITE}, where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, yielding net positive increases in plant performance [I could be more specific if needed]. Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts [weak ending to this sentence].

Plants gather silicon from the soil solution, using a suite of transporter proteins to pump it into their vascular systems and then transport it throughout the body.  Variation in the relative expression of these transporters, as well as differences in the development of the end points for silicon deposition (silica cells), may drive phenotypic variation among individuals. Additionally, individuals may vary in their ability to scavenge silicon from the soil. The soluble form of silicon, silicic acid (SiOH4) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM. Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilization of silicon in decaying plant material. Weathering of silicates releases a host of plant nutrients including Al, Si, Fe, and P. Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms. Plant roots can release carboxylates and phytosiderophores to weather P and Si out of soil minerals. Along with Si and P mobilization, Mn is often released, and taken up by plants roots. Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants, yet so far I are unaware of any studies looking for quantitative variation among genotypes of leaf Mn. If I could identify regions of the plant genome associated with variation in root weathering activity, I may be able to target this trait in breeding programs that improve nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.

The use of x-ray fluorescence (XRF) to quantify plant silicon has greatly reduced the costs, danger, and processing time of for studies focussing on this topic. XRF works by using low-power x-rays to excite elements in the sample, and measures the resulting emitted light. One of the most exciting features of XRF is the fact that it can analyse multiple elements at once, allowing for broad characterization of the sample for most elements heavier than aluminum. Though XRF is an established technique to measure plant Si, its may also be used to measure other metals of interest, including manganese. In this study I use XRF to quantify variation in Si and Mn content among a diversity panel of a wild ancestor of bread wheat, Aegilops tauschii. This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link Si and Mn variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
\subsection{Methods}
\subsubsection{Plant Growing Conditions}
For this experiment, I used a the L2 panel of Aegilops tauschii from (Gaurav et al. 2021). grown at three different sites. Two of the sites were outdoors, with planting occurring in the fall, while the third site was a glasshouse, where I vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment. For full site details see Supplementary Table S1. Using 151 accessions, I started trays of seedlings in glasshouse or growth chamber environments. At approximately 8 weeks after germination, seedlings were transplanted to their field sites. For each environment, I started four replicates of each accession. I planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements. Each block was a 16 m2 square, with plants arranged ~35 cm apart. Shortly after transplanting to the field sites, I applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets. Field transplantation took place on the 15th of October 2022 and the 16th of December 2022. For the glasshouse set, I started seedlings in growth chambers in January 2022. After 12 weeks, I moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for 8 weeks. I then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with [amount] of silicic acid (Tixosil 68B, Solvay). To ensure a comparable life stage at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads. 
\subsubsection{Plant Harvest and Sample Preparation}
When the plants had reached maturity, I harvested the entire above-ground portion of each plant. For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while I harvested the glasshouse plants between the 19th and 21st of September 2022. I placed harvested material in labelled paper bags, and dried it in drying ovens at 60ºC for 48 hours. To harvest leaf material for analysis, I selected stems with flower heads, and removed the three leaves closest to the flowers. Since portions of the plant have different silicon contents (Dai et al. 2005), I chose a consistent set of leaves to minimize introduced variations. I picked leaves until approximately 200mg of dry leaf was collected. I then washed leaves in distilled water to remove any soil residues which might introduce silicon, and re-dried the samples at 60ºC for 48 hours. Dried, clean leaves were then packed into 2 ml microcentrifuge tubes with zircon grinding pellets, flash frozen by immersion in liquid nitrogen, and ground in a tissuelyser ball mill for 30s. The resulting leaf powder was stored sealed until XRF analysis. 
\subsubsection{Sample Analysis}
To analyse the silicon and manganese content of the accessions, I followed the XRF procedure presented in (Reidinger et al. 2012). In short, I pressed leaf powder into 13mm diameter pellets at 11 tonnes of pressure. The resulting pellets were analysed in an Olympus Vanta p-XRF device mounted in a bench stand. I used a read time of 30s to ensure accurate measurements.

\subsection{Results}

Of the approximately 1700 plants planted, 1300 produced enough leaf material for analysis. Silicon content in \textit{Aegilops tuaschii} ranged from 0.756\% to 0.865\%. The various growing environments drove a large amount of variation in silicon content. Overall, my analysis revealed four regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}). One of these genomes was on chromosome 4S, near a known gene analogue to \textit{Lsi1}, a silicon transporter protein. My results for manganese content are less clear. I detected no regions that met the threshold for significance, though there were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}). Within the plants, silicon and manganese content were correlated ($R^2$ = 0.15, p = 0.049) (Figure 3).

\subsection{Discussion}

\subsection{Acknowledgements}

\subsection{Data Availability}

\subsection{Tables and Figures}

\begin{figure}[h]
        \includegraphics{images/Manhattan_Plot.png}
        \centering
        \caption{This is an example Manhattan Plot from the GWAS output. The real figure will show associations with silicon content}
        \label{Fig:si_peak_plot}
\end{figure}
\begin{figure}[h]
        \includegraphics{images/Manhattan_Plot.png}
        \centering
        \caption{This is another Manhattan Plot, this time showing associations with manganese content}
        \label{Fig:mn_peak_plot}
\end{figure}

\begin{figure}[h]
        \includegraphics{images/si_mn_regression.png}
        \centering
        \caption{This is the regression comparing Si to Mn content in the leaf tissue}
        \label{Fig:mn_si_regression}
\end{figure}

\clearpage

\section{References}
\printbibliography

\end{document}

### 2022-12-26 manuscript/title.tex
wrote more

\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

        \textbf{The ecological and genetic drivers of silicon accumulation in cereal crops}

        \vspace{1cm}

        by

        \vspace{1cm}

        \textbf{Isaac Peetoom Heida}

        \vfill

        \textsc{a thesis submitted in partial fulfillment of the requirements for the degree of}

        \vspace{1cm}

        \textsc{master of science}

        in

        \textsc{the faculty of graduate and postdoctoral studies}

        (Plant Science)

        \textsc{the university of british columbia}

        (Vancouver)

        \copyright Isaac Peetoom Heida, 2023
    \end{center}
\end{titlepage}

### 2022-12-27 manuscript/main.tex
is this the first draft done?

@@
 
-As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying.
+As global agricultural production strains strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying.
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
@@
 
-[GWAS results, Rapid Si results]
+[to add: results, conclusion]
 
-\section*{Lay Summary}
+\section{Lay Summary}
 
@@
 
-\section*{Preface}
+\section{Preface}
+
 (This was taken nearly verbatim from Matt's thesis, so need to go back through to make sure I am not plagarizing)
@@
 Juli Carrillo and Dr.
-Gurcharn Singh Brar conceptualized and developed the experiement presented in Chapter 2.
+Gurcharn Singh Brar conceptualized and developed the experiment presented in Chapter 2.
 Isaac Peetoom Heida developed the question and methodology for Chapter 1.
@@
 
-\subsection{Ag good, climate change bad, silicon helps}
+\subsection{A case for silicon in agriculture}
 
-As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying.
+As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, researchers are leaving no stone unturned in the search for new technologies for sustainable improvement in crop production.
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
-One of the promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance.
+Over the past 20 years, plant-silicon relations has emerged as a promising field that may safeguard crop performance and security within a changing biosphere.
 With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
@@
 
-Silicon is super-abundant in the earth’s crust, with Silicon dioxide (SiO2) comprising about 60\% of the crust by mass.
+Silicon abounds in the earth’s crust, with silicon dioxide (SiO\textsubscript{2}) comprising about 60\% of the crust by mass.
 Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant.
+Silicon can be deposited throughout the whole plant, forming amorphous masses of silica, with highly variable geometries (\cite{piperno_phytoliths_2006}).
+Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation.
 When supplemented with silicon, plants are generally more resistant to stress.
@@
 Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}).
-Silicon is deposited throughout the plant body, forming amorphous masses of silica, with highly variable geometries (\cite{piperno_phytoliths_2006}).
-Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}).
-Recent evidence points to interactions with gene expression regulators.
 Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}).
-The resulting model from this research posits that much of silicon’s beneficial role arises from phytolith interactions with compounds adverse to plant growth, limiting the penetration of the compounds into cells where damage can be caused.
-For herbivore defence, phytoliths serve a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
+The recently put-forward apoplastic barrier hypothesis suggests that the various benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition (\cite{coskun_controversies_2019}).
+Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector proteins, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
 Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
@@
 Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems.
-Within the soil environment, silicon commonly exists in both crystalline (geologic) forms as well as amorphous (biogenic) forms (\cite{haynes_contemporary_2014}).
+Within the soil environment, silicon commonly exists in both crystalline (geologic) and amorphous (biogenic) forms (\cite{haynes_contemporary_2014}).
 Amorphous silicates can derive from previous plant material that has decayed in the soil, but also from marine and aquatic organisms such as diatoms.
-Globally, the silicon cycle has silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments.
-The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the plant silicon being recycled from previous plant material (\cite{de_tombeur_plants_2020}).
-In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients for uptake.
-Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root (\cite{de_tombeur_silicon_2021-1}).
-This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation my improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).
+Globally, the silicon cycle involves silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments.
+The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil(\cite{de_tombeur_plants_2020}).
+In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients, including silicon, for uptake.
+Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
+This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation may improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).
 
@@
 Two of these proteins (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}).
+Homologues of these proteins have been identified in other cereal crops, and additional analogous silicon transporter proteins have been discovered in Cucurbitaceae (\cite{reynolds_silicon_2016}).
 Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}).
@@
 Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
-Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure for these bodies to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
+Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these bodies to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
 In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}).
-This trait has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation reduces the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}).
-In the leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf.
-This toughness, and the hardness of the phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics like stem phytoliths (\cite{stromberg_functions_2016}).
+The use of silicon as a structural component represents a highly energetically efficient strategy, as silicon is ~10x cheaper on an energy unit basis to produce than lignin (\cite{stromberg_functions_2016}).
+Stem silicification has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation has been found to reduce the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}).
+In leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf (\cite{simpson_still_2017}).
+This overall toughness, and abundance of phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}).
 Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}).
 As phytoliths deposit in the leaves, polymerization is aided by interactions with proteins in the cell wall, which control sites of nucleation (\cite{nawaz_phytolith_2019}).
-Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (Waterman et al.
-2021b).
-Thus, silicon-based defences in crop plants may be one of the first lines of active defence, providing rapid and sensitive responses to herbivory.
+Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (\cite{waterman_short-term_2021}).
+Thus, silicon-based defences in crop plants may be one of the first lines of active anti-herbivore defence, providing rapid and sensitive responses to herbivory.
 
@@
 
-Much of today’s plant silicon work is indebted to the pioneering work (\cite{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by (\cite{takahashi_possibility_1990}).
-Epstein’s seminal (\cite{epstein_silicon_1999}) paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers extending the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world.
-Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade.
-Rice, maize, wheat, and barley alone account for 1/3 of the worlds’ total cultivated landmass (\cite{faostat}), and are all domesticated grass species.
-Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich substrates (\cite{haynes_contemporary_2014,meena_case_2014}).
-Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{} Fraysse et al.
-2009).
-Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate production systems, particularly in wheat and barley.
-This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficeincy in which these temperate crops utilize the ample silicon available in their soils.
+Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by Takahashi et al.
+(\citeyear{takahashi_possibility_1990}).
+Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world.
+Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}).
+Rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}), and are all domesticated grass species.
+Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}).
+Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}).
+Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate dry-land production systems, particularly in wheat and barley.
+This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils.
 Our ability to integrate silicon as a tool for production improvement in dry-land grain production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which dry-land cereals utilize silicon in pest-protection.
@@
 
+\subsection{Introduction}
+
 During times of crisis, having an effective and timely response can limit damage and speed recovery.
-To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant.
-Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant.
-In most scenarios, induced plant defences are activated in response to an external cue, and build in intensity over time, with defensive hormones peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}).
-Despite this rapid hormonal response, actual defensive compounds are slower to accumulate, taking (need to figure out if this is true).
-Many defensive responses are also context dependent, where the identity of the damaging actor, the severity of damage, and a host of other factors interact to determine the final defensive response.
-The most effective defensive strategies are those that can either prevent herbivory outright, or can mount a rapid response to limit damage.
+To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
+Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
+In most scenarios, induced plant defences are activated in response to an external cue, and build in intensity over time, with defensive hormone signals peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}).
+Despite this rapid hormonal response, actual defensive phenotypes are slower to emerge, often operating on the scale of days or generations (\cite{karban_induced_1989}).
+Many defensive responses are also context dependent, where the identity of the damaging actor, the severity of damage, and a host of other factors interact to determine the final defensive response (\cite{waterman_simulated_2019}).
+The most effective defensive strategies should be those that can either prevent herbivory outright, or can mount a rapid response to limit damage.
 These same strategies are also the most promising for crop production, where pest damage represents both an economic and food security cost.
@@
 
-One of the most promising avenues for new crop defence is the harnessing of silicon.
-Silicon acts on multiple temporal and physiological scales, delivering broad spectrum resistance to pests, pathogens, and abiotic stressors.
-Soluble silicon taken up from the soil is deposited predominantly in the leaf epidermis, where it forms solid granules that increase the toughness of the tissue, reducing herbivore digestive efficiency.
-Plant silicon is expressed latently, but also increases in response to herbivory.
+One of the most promising avenues for new crop defence is the harnessing of silicon (\cite{reynolds_silicon_2016}).
+Silicon acts on multiple temporal and physiological scales, delivering broad spectrum resistance to pests, pathogens, and abiotic stressors (\cite{cooke_consistent_2016,coskun_controversies_2019}).
+Soluble silicon taken up from the soil is deposited predominantly in the leaf epidermis, where it forms solid granules that increase the toughness of the tissue, reducing herbivore digestive efficiency (\cite{cooke_is_2011}).
+Plant silicon is expressed latently, but also increases in response to herbivory (\cite{takahashi_possibility_1990, }).
 Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even preempting some chemical defences (\cite{waterman_short-term_2021}).
@@
 Despite the novel results, this pattern has so far been observed in just one species, and only under artificial herbivory via the application of methyl-jasmonate.
-Though a useful tool for herbivory research, methyl-jasmonate application fails to reproduce a complete herbivory signal for the plant and thus observed changes to plant defence may not be representative of a true herbivory scenario (\cite{strauss_direct_2002}).
+Though a useful tool for herbivory research, methyl-jasmonate application fails to reproduce a complete herbivory signal for the plant, thus observed changes to plant defence may not be representative of a true herbivory scenario (\cite{strauss_direct_2002}).
 Testing for this rapid silicon accumulation across a variety of grain crops, and under both simulated (methyl-jasmonate) and real herbivory is a crucial first step towards integrating rapid silicification into our understanding of plant defence and crop protection.
 
-Plant silicon research has mostly focused on members of the grass family (Poaceae) due to their exceptional silicon content within the plant kingdom, as well as the economic importance of domesticated grass species.
+Plant silicon research has mostly focused on members of the grass family (Poaceae) due to their exceptional silicon content within the plant kingdom, as well as the economic importance of domesticated grass species (\cite{reynolds_silicon_2016}).
 Domesticated crops differ significantly from their wild relatives, due to effects of strong selective pressure imposed by humans (\cite{chen_crop_2015}).
@@
 Indeed, in the context of silicon, we can detect clear signals of domestication across the Poaceae family, where wild ancestors consistently have higher baseline silicon content than their domesticated descendants (\cite{simpson_still_2017}).
-Due to the effects of selection on plant defence it becomes crucial to test new developments in the silicon-defence literature in modern crop species, both to validate their utility towards agricultural production, and in this case to gather further observations on the dynamics of silicon-based defences in the first hours after herbivory.
+Due to the effects of selection on plant defence it becomes crucial to test new developments in the silicon-defence literature in modern crop species, both to validate their utility towards agricultural production, and to gather further observations on the dynamics of silicon-based defences in the first hours after herbivory.
 
@@
 \end{enumerate}
-This study is a thematic replication of Waterman et al.’s 2021 paper (\cite{waterman_short-term_2021}), but attempts to extend the findings to commercially important grain crops.
+This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, but attempts to extend the findings to commercially important grain crops.
 The findings of this study will refine our understanding of the prevalence of rapid silicification in the Poaceae, and will help to inform the value of potential applications of silicon-based defences into grain crops.
@@
 At the start of the experiment, I germinated seeds in germination trays filled with moist sand.
-After two days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
-Though potting mix and the water used contain some amount of plant available silicon, the silicic acid was added to ensure that there would be no silicon limitation to the plants.
+After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
+Though potting mix and fresh water contain some amount of plant available silicon, the silicic acid was added to ensure that there would be no silicon limitation to the plants.
+I randomized the location of each pot within the growing space.
 The pots were bottom watered on flood tables with nutrient solution.
-Each plant was assigned to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
+I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
 Simulated herbivory was achieved by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al.
@@
 
-24 hours after treatment application, I harvested leaf material by clipping the above-ground portion of the plants.
-Leaf tissue was flash frozen in liquid nitrogen and freeze dried, and stored in a -80ºC freezer.
-After transferring to 2mL microcentrifuge tubes, I flash froze the leaf tissue and ground for 30 seconds in a tissuelyser bead mill.
+24 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e.
+no cricket exposure) leaves.
+Leaf tissue was flash frozen in liquid nitrogen, freeze dried, and stored in a -80ºC freezer.
+To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube.
+I then flash froze the tubes in liquid nitrogen and ground the leaf tissue in a tissuelyser bead mill for 30 seconds at 60 Hz.
 
@@
 
-To measure the silicon content of the leaf tissue, I followed a modified version of the benchtop XRF method (\cite{reidinger_rapid_2012}).
-I pressed ground leaf tissue in a hydraulic press, using a 13mm die at 11 tons of pressure.
+To measure the silicon content of the leaf tissue, I followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
+I pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
 I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon.
+The pellet die and the analyzer were wiped clean after each use to minimize contamination between samples.
 
@@
 I ran the model using \verb|Turing.jl| in Julia [version number] [cite].
-Using 4 chains and 1000 sample iterations, I sampled the posterior distribution using a No U-turn Sampler with 1000 warm up iterations and a target acceptance rate of 0.65 (Hoffman and Gelman 2014).
+Using 4 chains and 1000 sample iterations, I sampled the posterior distribution using a No U-turn Sampler with 1000 warm up iterations and a target acceptance rate of 0.65 (\cite{hoffman_no-u-turn_2014}).
 I tested our model structure on simulated data, to ensure it returned accurate parameter estimates.
-I used the Gelman-Rubin statistic (\( \hat{R} \)) (Gelman and Rubin 1992) and effective sample size to diagnose the convergence of our chains.
+I used the Gelman-Rubin statistic (\( \hat{R} \)) (\cite{gelman_inference_1992}) and effective sample size to diagnose the convergence of our chains.
 I verified model fits using posterior predictive checks in \verb|ArviZ.jl| [cite]
@@
 \section{Chapter 2: Genetic drivers of silicon accumulation in a wild ancestor of wheat}
+
 \subsection{Introduction}
+
 With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production.
-Over the past thirty years, research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops.
-Cereal crops are globally important, covering over 1/3 of the world’s arable land, making up over 50\% of the daily caloric intake for most people.
-Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (>0.75\% total dry weight).
-Silicon is highly abundant in many soils globally, and is the second most abundant element in the earth’s crust, behind only oxygen.
+Over the past thirty years, research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
+Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}).
+Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\>0.75\% total dry weight) (\cite{reynolds_silicon_2016}).
+Silicon is highly abundant in many soils globally, and is the second most abundant element in the earth’s crust, behind only oxygen (\cite{ma_functions_2003}).
 It’s high expression in cereals, high abundance in many soils, and incredible broad spectrum effects on plant vigor and stress tolerance have make it a tantalizing target for improvements in agricultural yield and sustainability.
-Silicon is now considered a quasi-essential nutrient for plant growth.
 Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.
@@
 Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress.
-For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms.
-On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress.
-In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants.
-A current hypothesis explaining the broad-spectrum activity of silicon is presented in {CITE}, where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, yielding net positive increases in plant performance [I could be more specific if needed].
-Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts [weak ending to this sentence].
+For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005, katz_silicon_2019}).
+On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}).
+In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}).
+A current hypothesis explaining the broad-spectrum activity of silicon is presented in Coskun et al.
+(\citeyear{coskun_controversies_2019}), where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance [I could be more specific if needed].
+Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts.
+Finding ways to improve crops towards increased silicon use efficiency is key to harnessing the benefits that plant silicon can confer.
 
-Plants gather silicon from the soil solution, using a suite of transporter proteins to pump it into their vascular systems and then transport it throughout the body.
+Plants gather silicon from the soil solution, using a suite of transporter proteins to pump it into their vascular systems and then transport it throughout the body (\cite{reynolds_silicon_2016}).
 Variation in the relative expression of these transporters, as well as differences in the development of the end points for silicon deposition (silica cells), may drive phenotypic variation among individuals.
 Additionally, individuals may vary in their ability to scavenge silicon from the soil.
-The soluble form of silicon, silicic acid (SiOH4) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM.
-Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilization of silicon in decaying plant material.
-Weathering of silicates releases a host of plant nutrients including Al, Si, Fe, and P.
+The soluble form of silicon, silicic acid (SiOH4) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM (\cite{epstein_anomaly_1994}).
+Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilization of silicon in decaying plant material (\cite{de_tombeur_silicon_2021-1}).
+Weathering of silicates releases a host of plant nutrients including Al, Si, Fe, and P (\cite{de_tombeur_silicon_2021-1}).
 Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms.
@@
 Along with Si and P mobilization, Mn is often released, and taken up by plants roots.
-Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants, yet so far I are unaware of any studies looking for quantitative variation among genotypes of leaf Mn.
-If I could identify regions of the plant genome associated with variation in root weathering activity, I may be able to target this trait in breeding programs that improve nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
+Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far I are unaware of any studies looking for quantitative variation among genotypes of leaf Mn.
+If we could identify regions of the plant genome associated with variation in root weathering activity, we may be able to target this trait in breeding programs that improve nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
 
-The use of x-ray fluorescence (XRF) to quantify plant silicon has greatly reduced the costs, danger, and processing time of for studies focussing on this topic.
+The use of x-ray fluorescence (XRF) to quantify plant silicon has greatly reduced the costs, danger, and processing time of for studies focussing on this topic (\cite{reidinger_rapid_2012}).
 XRF works by using low-power x-rays to excite elements in the sample, and measures the resulting emitted light.
@@
 This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link Si and Mn variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
+
 \subsection{Methods}
+
 \subsubsection{Plant Growing Conditions}
-For this experiment, I used a the L2 panel of Aegilops tauschii from (Gaurav et al.
-2021).
-grown at three different sites.
-Two of the sites were outdoors, with planting occurring in the fall, while the third site was a glasshouse, where I vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment.
+
+For this experiment, I used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites.
+Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where I vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment.
 For full site details see Supplementary Table S1.
 Using 151 accessions, I started trays of seedlings in glasshouse or growth chamber environments.
-At approximately 8 weeks after germination, seedlings were transplanted to their field sites.
+At approximately eight weeks after germination, seedlings were transplanted to their field sites.
 For each environment, I started four replicates of each accession.
 I planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements.
-Each block was a 16 m2 square, with plants arranged ~35 cm apart.
+Each outdoor block was a 16 m\textsuperscript{2} square, with plants arranged $\sim$35 cm apart.
 Shortly after transplanting to the field sites, I applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets.
 Field transplantation took place on the 15th of October 2022 and the 16th of December 2022.
-For the glasshouse set, I started seedlings in growth chambers in January 2022.
-After 12 weeks, I moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for 8 weeks.
+For the glasshouse environment, I started seedlings in growth chambers in January 2022.
+After 12 weeks, I moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks.
 I then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with [amount] of silicic acid (Tixosil 68B, Solvay).
-To ensure a comparable life stage at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.
+Pots were arranged using the same randomized block design, adapted to fit on two flood tables.
+To ensure a comparable life stage accross environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.
+
 \subsubsection{Plant Harvest and Sample Preparation}
+
 When the plants had reached maturity, I harvested the entire above-ground portion of each plant.
@@
 To harvest leaf material for analysis, I selected stems with flower heads, and removed the three leaves closest to the flowers.
-Since portions of the plant have different silicon contents (Dai et al.
-2005), I chose a consistent set of leaves to minimize introduced variations.
+Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), I chose a consistent set of leaves to minimize introduced variation.
 I picked leaves until approximately 200mg of dry leaf was collected.
 I then washed leaves in distilled water to remove any soil residues which might introduce silicon, and re-dried the samples at 60ºC for 48 hours.
-Dried, clean leaves were then packed into 2 ml microcentrifuge tubes with zircon grinding pellets, flash frozen by immersion in liquid nitrogen, and ground in a tissuelyser ball mill for 30s.
+Dried, clean leaves were then packed into 2 ml microcentrifuge tubes with zircon grinding pellets, flash frozen by immersion in liquid nitrogen, and ground in a tissuelyser ball mill for 30 seconds at 60 Hz.
 The resulting leaf powder was stored sealed until XRF analysis.
+
 \subsubsection{Sample Analysis}
-To analyse the silicon and manganese content of the accessions, I followed the XRF procedure presented in (Reidinger et al.
-2012).
-In short, I pressed leaf powder into 13mm diameter pellets at 11 tonnes of pressure.
-The resulting pellets were analysed in an Olympus Vanta p-XRF device mounted in a bench stand.
+
+To analyse the silicon and manganese content of the accessions, I followed the XRF procedure presented in Reidinger et al.
+(\citedate{reidinger_rapid_2012}).
+In short, I pressed leaf powder into 13mm diameter pellets at 11 tonnes of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
 I used a read time of 30s to ensure accurate measurements.
+To minimize cross-contamination between samples, I wiped clean the pellet press and XRF device after each sample.
 
@@
 
-Of the approximately 1700 plants planted, 1300 produced enough leaf material for analysis.
-Silicon content in \textit{Aegilops tuaschii} ranged from 0.756\% to 0.865\%.
+Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis.
+Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{value}]\% to [\textit{value}\%].
 The various growing environments drove a large amount of variation in silicon content.
-Overall, my analysis revealed four regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}).
+Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}).
 One of these genomes was on chromosome 4S, near a known gene analogue to \textit{Lsi1}, a silicon transporter protein.

### 2023-01-20 manuscript/main.tex
added line numbers

@@
 \usepackage{url}
+\usepackage{lineno}
+\linenumbers
 \graphicspath{./images/}
@@
 
-\subsubsection{Plant Growth and Treatments}
+\subsubsection{Plant growth and experimental treatments}
 
@@
 
-\subsubsection{Sample Harvest and Preparation}
+\subsubsection{Sample harvest and preparation}
 
@@
 
-\subsubsection{Statistical Analysis}
+\subsubsection{Statistical analysis}
 
@@
 
-\subsubsection{Plant Growing Conditions}
+\subsubsection{Plant growing conditions}
 
@@
 
-\subsubsection{Plant Harvest and Sample Preparation}
+\subsubsection{Plant harvest and sample preparation}
 
@@
 
-\subsubsection{Sample Analysis}
+\subsubsection{Sample analysis}
 
@@
 To minimize cross-contamination between samples, I wiped clean the pellet press and XRF device after each sample.
+
+\subsubsection{Statistical analysis}
+
+To perfrom the GWAS analysis, I followed the methodology and code published in Gaurav et al.
+\citedate{gaurav_population_2021}.
+For brevity, this methodology only describes the steps we took using the data generated from Gaurav et al.
+\citedate{gaurav_population_2021}.
+For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
+Prior to the linear regression to assess genetic associations, we first tested our phenotypic data for block effects within sites.
+For each site, we tested a linear model to see if measured leaf silicon content was significantly associated with block.
+We tested the model $y \sim Block + Accession$.
+When we validated that block had a negligible effect ($p>0.1$), we reparameterized leaf silicon measurements at the site level.
+We converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si.
+I did this to eliminate a possible confounding effect of inter-site differences that might reduce the power of the GWAS linear regression.
+
+As per Gaurav et al.
+\citedate{gaurav_population_2021}, to reduce the computational intensity of my analysis, I prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative [word choice?].
 

### 2023-01-27 manuscript/main.tex
added Daria's last name
updated Preface, abstract, lay summary

@@
 
-As global agricultural production strains strains under degrading soil fertility and increasing losses due to climate change, research interest in new avenues for production improvement is intensifying.
+As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, there is growing research interest in new avenues for production improvement.
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
@@
 Our ability to mobilize silicon-based cropping strategies is dependent on a thorough understanding of the ultimate and proximate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
-In this thesis, I attempt to extend recent advances in our understanding of silicon ecology in cereal crops, taking an integrative approach towards untangling silicon accumulation in cereal crops.
-I used a genome-wide association study to identify genetic markers associated with high silicon content, and performed a greenhouse experiment to test for patterns of rapid silicon accumulation in four cereal crops.
+In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common canadian crops, as well as using a  genome-wide association study to identify genetic markers associated with high silicon content.
 
@@
 
-Silicon, a naturally occurring element, provides tremendous benefits to plant health, but is not widely utilized in agriculture.
+Silicon provides tremendous benefits to plant health, but is not widely utilized in agriculture, despite it being highly abundance and naturally occurring in most soils.
 One of the main factors limiting it's application in agriculture is our poor understanding of the exact dynamics of how plants absorb and use silicon from the soil.
-I identified genetic traits that are associated with high silicon content in a relative of bread wheat, as well as demonstrated that cereal crops (e.g.
+I identified genetic traits associated with high silicon content in \textit{Aegilops tauschii}, a relative of bread wheat, as well as demonstrated that cereal crops (e.g.
 wheat, barley, oats) have the ability to rapidly uptake silicon from the soil.
 This rapid uptake means that silicon may be a highly effective defence against insect pests.
-Combining this results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content.
+Combining these results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content.
 This development could provide an environmentally friendly strategy to maintain output to feed a growing human population.
@@
 Isaac Peetoom Heida and Dr.
-Juli Carrillo conceptualized and developed the experiment presented in Chapter 1.
+Juli Carrillo, with assistance from Dr.
+Jean-Thomas Cornelis, conceptualized and developed the experiment presented in Chapter Two.
 Isaac Peetoom Heida, Dr.
 Juli Carrillo and Dr.
-Gurcharn Singh Brar conceptualized and developed the experiment presented in Chapter 2.
-Isaac Peetoom Heida developed the question and methodology for Chapter 1.
+Gurcharn Singh Brar, with input from Dr.
+Jean-Thomas Cornelis, conceptualized and developed the experiment presented in Chapter Three.
+Isaac Peetoom Heida developed the question and methodology for Chapter Two.
 Dr.
@@
 Shaun Barker and the Mineral Deposit Research Unit of the University of British Columbia provided facilities and expertise for the XRF analysis of the tissue samples.
-(but I paid them, do I still thank?)
-For chapter 2, Isaac Peetoom Heida and Dr.
-Gurcharn Brar designed the experiment.
-Isaac Peetoom Heida let plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda.
+For Chapter Three, Isaac Peetoom Heida led plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda.
 Isaac Peetoom Heida led the sample harvest, processing and analysis.
@@
 Dr.
-Daria [last name], Paul Fisher, Lucas Friesen, Katie Pryer, Dr.
+Daria Reshetniak, Paul Fisher, Lucas Friesen, Katie Pryer, Dr.
 Kinga Treder, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
-Chapters two and three of this thesis will be submitted to peer-reviewed journals for publication. For the purposes of this manuscript, actions are depicted in the singular first person. 
+Chapters Two and Three of this thesis will be submitted to peer-reviewed journals for publication. For the purposes of this manuscript, actions are depicted in the singular first person.
+
 \section{Introduction}

### 2023-01-27 manuscript/main.tex
eliminated passive voice like a boss

@@
 
-Silicon abounds in the earth’s crust, with silicon dioxide (SiO\textsubscript{2}) comprising about 60\% of the crust by mass.
+Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass.
 Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant.
-Silicon can be deposited throughout the whole plant, forming amorphous masses of silica, with highly variable geometries (\cite{piperno_phytoliths_2006}).
+Silicates are found in a variety of forms, and vary in their plant availability.
+Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available.
+As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths.
+Plant phytoliths are amorphous masses of silica, generally similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}).
+Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations.
+The highly variable morphology among clades, with the conservation of morphology within clades suggests active selection upon the structure of phytoliths.
+The morphology of phytoliths also varies between organs in the plant.
+Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory.
+
 Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation.
@@
 To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley.
-Cultivars were selected on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
+I selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
 At the start of the experiment, I germinated seeds in germination trays filled with moist sand.
 After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
-Though potting mix and fresh water contain some amount of plant available silicon, the silicic acid was added to ensure that there would be no silicon limitation to the plants.
+Though potting mix and fresh water contain some amount of plant available silicon, I added the silicic acid to ensure that there would be no silicon limitation to the plants.
 I randomized the location of each pot within the growing space.
-The pots were bottom watered on flood tables with nutrient solution.
+A flood table bottom watered the pots with nutrient solution.
 I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
-Simulated herbivory was achieved by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al.
-2021b), while true herbivory was provided by crickets housed in water-pik tubes.
+I simulated herbivory by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al.
+2021b), while crickets housed in water-pik tubes provided true herbivory.
 Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial.
@@
 no cricket exposure) leaves.
-Leaf tissue was flash frozen in liquid nitrogen, freeze dried, and stored in a -80ºC freezer.
+I flash froze leaf tissue liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer.
 To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube.
-I then flash froze the tubes in liquid nitrogen and ground the leaf tissue in a tissuelyser bead mill for 30 seconds at 60 Hz.
+I cryoground the leaf tissue at -80ºC in a tissuelyser bead mill for 30 seconds at 60 Hz.
 
@@
 I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon.
-The pellet die and the analyzer were wiped clean after each use to minimize contamination between samples.
+After each use, I wiped the pellet die and the analyzer clean to minimize contamination between samples.
 

### 2023-01-27 manuscript/main.tex
herbivory terminology changes

@@
 
-During times of crisis, having an effective and timely response can limit damage and speed recovery.
 To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
+In the broadest of terms, plant defenses can be categorized as either tolerance of resistance strategies.
+Resistance strategies involve attempts to limit tissue loss through interference with herbivory, while tolerance strategies result in increased regrowth after tissue loss.
 Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
@@
 
-In this study, I test four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory.
+In this study, I tested four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory.
 In a glasshouse environment, I grew bread wheat (\textit{Triticum aestivum}), oats (\textit{Avena sativa}), barley (\textit{Hordeum vulgare}) and Triticale ($\times$ \textit{Triticosecale}), and tested the following hypotheses:
 \begin{enumerate}
-        \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species silicon content would show responses to herbivory consistent with this process.
+        \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species silicon content would show a significant increase in silicon content within 18 hours of the herbivory treatment applications.
         \item Due to different phylogeny and domestication history, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
@@
 
-To answer all three of our questions, I used a bayesian hierarchical model. [Insert text about how great these models are]. I specified a hierarchical model using the following model:
+To answer test all three of our hypotheses, I used a bayesian hierarchical model. [Insert text about how great these models are]. I specified a hierarchical model using the following model:
 

### 2023-03-15 manuscript/main.tex
catching up commit

@@
 Silicates are found in a variety of forms, and vary in their plant availability.
-Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available.
-As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths.
-Plant phytoliths are amorphous masses of silica, generally similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}).
+Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available \cite{fraysse_surface_2009}.
+As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths \cite{de_tombeur_plants_2020}.
+Uptake and usage of silicon is not uniform throughout the vascular plants.
+Plant clades vary in their expression of silicon transporter proteins and in the relative abundance of phyotliths in their tissues.
+Plant phytoliths are amorphous masses of silica, chemically similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}).
 Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations.
@@
 
-Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation.
+Silicon deposition may create apoplastic barriers that seal and toughen the plant tissue, which may benefit the plant in a multitude of ways (\cite{coskun_controversies_2019}).
+Under this hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
+Empirical evidence shows that silicon is effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}).
+The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
+Indeed, when supplemented with silicon plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
+Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
+
+%Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation.
 When supplemented with silicon, plants are generally more resistant to stress.
@@
 Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}).
-The recently put-forward apoplastic barrier hypothesis suggests that the various benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition (\cite{coskun_controversies_2019}).
-Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector proteins, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
+The recently put-forward apoplastic barrier hypothesis suggests that the benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition on the apoplast (\cite{coskun_controversies_2019}).
+Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.
+The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
 Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
@@
 While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families.
-Research in rice has identified four proteins that transport silicon into and through the plant body.
-Two of these proteins (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}).
-Homologues of these proteins have been identified in other cereal crops, and additional analogous silicon transporter proteins have been discovered in Cucurbitaceae (\cite{reynolds_silicon_2016}).
+Research in rice has identified four gene products that transport silicon into and through the plant body.
+Two of these (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}).
+Orthologs of these proteins have been identified in other cereal crops, and additional analogous silicon transporter proteins have been discovered in the Cucurbitaceae (\cite{reynolds_silicon_2016}).
 Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}).
@@
 Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
-Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these bodies to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
+Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
 In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}).
@@
 Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}).
-As phytoliths deposit in the leaves, polymerization is aided by interactions with proteins in the cell wall, which control sites of nucleation (\cite{nawaz_phytolith_2019}).
+As phytoliths develop in the leaves, polymerization of monosilicic acid is aided by interactions with proteins in the cell wall, which act as sites of nucleation (\cite{nawaz_phytolith_2019}).
 Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (\cite{waterman_short-term_2021}).
@@
 This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils.
-Our ability to integrate silicon as a tool for production improvement in dry-land grain production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which dry-land cereals utilize silicon in pest-protection.
+Our ability to integrate silicon as a tool for production improvement in crop production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which crops utilize silicon in pest-protection.
 
@@
 
-24 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e.
+18 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e.
 no cricket exposure) leaves.
-I flash froze leaf tissue liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer.
+I harvested three fully expanded upper leaves the plants, and split the plants in half (COME BACK HERE!!) I flash froze leaf tissue in liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer.
 To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube.
@@
 \citedate{gaurav_population_2021}.
-For brevity, this methodology only describes the steps we took using the data generated from Gaurav et al.
+For brevity, this methodology only describes the steps I took using the data generated from Gaurav et al.
 \citedate{gaurav_population_2021}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
-Prior to the linear regression to assess genetic associations, we first tested our phenotypic data for block effects within sites.
+Prior to the linear regression to assess genetic associations, I first tested my phenotypic data for block effects within sites.
 For each site, we tested a linear model to see if measured leaf silicon content was significantly associated with block.
-We tested the model $y \sim Block + Accession$.
-When we validated that block had a negligible effect ($p>0.1$), we reparameterized leaf silicon measurements at the site level.
-We converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si.
+I tested the model $Si \sim Block + Accession$.
+When I validated that block had a non-significant effect ($p>0.1$), I reparameterized leaf silicon measurements at the site level; I converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si.
 I did this to eliminate a possible confounding effect of inter-site differences that might reduce the power of the GWAS linear regression.
@@
 Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{value}]\% to [\textit{value}\%].
-The various growing environments drove a large amount of variation in silicon content.
+Site accounted for a large amount of variation in silicon content, suggesting a strong environmental effect on the silicon phenotype.
 Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}).

### 2023-03-15 src/cereal_induction.jl
started working with real data

@@ -40,10 +40,10 @@ end
 ### Prep data for model use
 include("turing_model.jl")
 analysis_data = CSV.read("data/fake_data/fake_data.csv", DataFrame)
-ad_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], interaction=false)
+analysis_data_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], interaction=false)
 #To incorporate Control & Barley into the intercept
 begin
-    ad_spread = ad_spread[:, Not([:treat1_Control, :treat2_Barley])]
+    analysis_data_spread = analysis_data_spread[:, Not([:treat1_Control, :treat2_Barley])]
 end
 
 #To center the data run this
@@ -111,5 +111,47 @@ plotting_data = CSV.read("data/fake_data/fake_data.csv", DataFrame)
 mp = @df plotting_data groupedboxplot(:Induction, :response, group = :Species)
 plot!(mp, xlabel = "Induction Treatment", ylabel = "Silicon Content (%)")
 png(mp, "induction_plot")
+test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = repeat(["D", "E", "F"], inner = 9, outer = 1), yvals = rand(Normal(0,1), 27))
+tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
+
+
+####
+
+# Running a the above models on my real data
+
+####
+
+analysis_data = CSV.read("data/real_data/cleaned_si_df.csv", DataFrame)
+ad_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], interaction=false)
+begin
+    ad_spread = ad_spread[:, Not([:treat2_Control, :treat1_Barley])]
+end
+
+#To center the data run this
+begin
+    cols_list = names(ad_spread)
+    scale_vals = DataFrame()
+    ads_centered = rescalecols(df=ad_spread, collist=[:Si_ppm], centers = scale_vals)
+end
+adsc_idx = stringcoltoint(df=ad_spread, stringcol=:Genotype, intcol=:idx)
+y_vals = Float64.(adsc_idx.Si_ppm)
+preds = Matrix(adsc_idx[:, 8:13])
+idx = Int.(adsc_idx.idx)
+my_model = randomintercept_regression(y_vals, preds, idx)
+num_chains = 4
[…]

### 2023-03-15 src/data_compiliaton.jl

@@ -1,0 +1,26 @@
+using CSV
+using DataFrames
+
+#=
+This script is written to combine herbivory data with the raw output of the pXRF analysis. We generated herbivory data by observing
+which cricket-exposed plants had suffered damage after the 18h exposure period. Damage was categorized into three classes: 
+undamaged, moderately damaged (leaf chewed by not severed), severly damaged (leaf severed). The pXRF data were generated
+using an Olympus Vanta pXRF device mounted in a benchtop stand. Prior to analysis, we pressed leaf tissue powder into a 13mm
+pellet using a hydraulic press and die set. We pressed the powder to 300 bar of pressue, resulting in a stable pellet with a 
+uniform surface. To quantify silicon, we used a 45 second scan time, a time that balanced throughput with accuracy (increasing 
+scan time by a factor of two resulted in only a 7% reduction in StDev of the measurement).
+=#
+
+herbivory_data = CSV.read("data/real_data/induction_data.csv", DataFrame)
+xrf_data = CSV.read("data/real_data/xrf_data.csv", DataFrame)
+xrf_colnames = names(xrf_data)
+xrf_si_slice = xrf_data[:, ["Si Concentration", "Sample ID"]] #select only the Si and ID columns
+rename!(xrf_si_slice, "Si Concentration" => "Si_ppm", "Sample ID" => "uniqueID") #rename so you don't have to deal with spaces
+combined_data = leftjoin(herbivory_data, xrf_si_slice, on = :uniqueID) #left join to move Si into the main dataframe
+filtered_combined_data = filter(row -> (ismissing(row.Damage_rank) || row.Damage_rank != "N") && !ismissing(row.Si_ppm), combined_data )
+
+CSV.write("data/real_data/cleaned_si_df.csv", filtered_combined_data)
+
+
+
+

### 2023-03-15 src/R_linear_model.R
re-ran models, added an R model to double-check

@@ -1,0 +1,12 @@
+install.packages("lmerTest")
+install.packages("tidyverse")
+install.packages("emmeans")
+library(lmerTest)
+library(tidyverse)
+library(emmeans)
+si_data <- as.data.frame(read.csv("data/real_data/cleaned_si_df.csv"))
+my_mod <- lmer(Si_ppm ~ Induction * Species + (1|Genotype), si_data)
+summary(my_mod)
+anova(my_mod)
+emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
+emmeans(my_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")

### 2023-03-15 src/cereal_induction.jl

@@ -143,7 +143,8 @@ chains = sample(my_model, NUTS(0.6), MCMCThreads(), 1_000, num_chains)
 summarystats(chains) |> DataFrame |> println
 plt = plot(chains)
 
-fit(MixedModel, @formula(Si_ppm ~ Induction * Species + (1|Genotype)), analysis_data)
+centered_data = rescalecols(df=analysis_data, collist=[:Si_ppm], centers = scale_vals)
+lm1 = fit(MixedModel, @formula(Si_ppm ~ Induction * Species + (1|Genotype)), centered_data)
 
 fm = @formula(Si_ppm ~ Induction + Species + (1|Genotype))
 model = turing_model(fm, analysis_data)

### 2023-03-17 src/R_linear_model.R
added biomass data

@@ -4,9 +4,11 @@ install.packages("emmeans")
 library(lmerTest)
 library(tidyverse)
 library(emmeans)
-si_data <- as.data.frame(read.csv("data/real_data/cleaned_si_df.csv"))
+si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 my_mod <- lmer(Si_ppm ~ Induction * Species + (1|Genotype), si_data)
 summary(my_mod)
 anova(my_mod)
 emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
-emmeans(my_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
+emmeans(my_mod, list(pairwise ~ Induction * Species), adjust = "tukey", type = "response")
+ggplot(si_data, aes(x = Induction, y = Si_ppm)) +
+geom_col()

### 2023-03-17 src/data_compiliaton.jl

@@ -13,13 +13,17 @@ scan time by a factor of two resulted in only a 7% reduction in StDev of the mea
 
 herbivory_data = CSV.read("data/real_data/induction_data.csv", DataFrame)
 xrf_data = CSV.read("data/real_data/xrf_data.csv", DataFrame)
+biomass_data = CSV.read("data/real_data/biomass_no106.csv", DataFrame)
 xrf_colnames = names(xrf_data)
 xrf_si_slice = xrf_data[:, ["Si Concentration", "Sample ID"]] #select only the Si and ID columns
 rename!(xrf_si_slice, "Si Concentration" => "Si_ppm", "Sample ID" => "uniqueID") #rename so you don't have to deal with spaces
-combined_data = leftjoin(herbivory_data, xrf_si_slice, on = :uniqueID) #left join to move Si into the main dataframe
-filtered_combined_data = filter(row -> (ismissing(row.Damage_rank) || row.Damage_rank != "N") && !ismissing(row.Si_ppm), combined_data )
+combined_data_1 = leftjoin(herbivory_data, xrf_si_slice, on = :uniqueID) #left join to move Si into the main dataframe
+combined_data_2 = leftjoin(combined_data_1, biomass_data, on = :uniqueID)
 
-CSV.write("data/real_data/cleaned_si_df.csv", filtered_combined_data)
+#filtered_combined_data = filter(row -> (ismissing(row.Damage_rank) || row.Damage_rank != "N") && !ismissing(row.Si_ppm), combined_data )
+
+#CSV.write("data/real_data/cleaned_si_df.csv", filtered_combined_data)
+CSV.write("data/real_data/biomass_si_data.csv", combined_data_2)
 
 
 

### 2023-03-19 manuscript/main.tex
merging bib files

@@
 \usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
-\addbibresource{total_biblio.bib}
-\addbibresource{datasets.bib}
+\addbibresource{merged_biblio2.bib}
+\addbibresource{internet_stuff.bib}
+%\addbibresource{induction.bib}
+%\addbibresource{gwas_silicon.bib}
 \usepackage{caption}
@@
 no cricket exposure) leaves.
-I harvested three fully expanded upper leaves the plants, and split the plants in half (COME BACK HERE!!) I flash froze leaf tissue in liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer.
-To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube.
-I cryoground the leaf tissue at -80ºC in a tissuelyser bead mill for 30 seconds at 60 Hz.
+I harvested three fully expanded upper leaves the plants, and split the plants in half along the midvein of the leaf.
+I placed one half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
+I plaed the other half of the tissue into a coin envelope, oven dried it for 4 days at 60\deg C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz).
+After freeze drying, the samples for phenolic processing were ground under the same conditions.
 
@@
 I pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
-I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon.
-After each use, I wiped the pellet die and the analyzer clean to minimize contamination between samples.
+I then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
+After each use, I cleaned the pellet die and XRF analyzer to minimize contamination between samples.
 
@@
 
-To answer test all three of our hypotheses, I used a bayesian hierarchical model. [Insert text about how great these models are]. I specified a hierarchical model using the following model:
+To answer test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}).
+Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels (\cite{carmona_plant_2011}).
+We found a negative correlation between plant size and silicon content (\beta = -0.067 \pm 0.017, p < 0.001), and thus included plant size as a covariate in our final model.
+Thus, we tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
 
-\[y_i \sim Normal(\hat{y}_i, \sigma)\]
-\[\hat{y_i} = \alpha + \alpha_{[i]cultivar} + \beta_{[j]induction} + \beta_{[k]species} + \beta_{[jk]species \times induction}\]
+\[Si ~ Species * Induction + Biomass + (1|Genotype)\]
+%[Insert text about how great these models are]. I specified a hierarchical model using the following model:
 
-I ran the model using \verb|Turing.jl| in Julia [version number] [cite].
+%\[y_i \sim Normal(\hat{y}_i, \sigma)\]
+%\[\hat{y_i} = \alpha + \alpha_{[i]cultivar} + \beta_{[j]induction} + \beta_{[k]species} + \beta_{[jk]species \times induction}\]
+
+%I ran the model using \verb|Turing.jl| in Julia [version number] [cite].
 Using 4 chains and 1000 sample iterations, I sampled the posterior distribution using a No U-turn Sampler with 1000 warm up iterations and a target acceptance rate of 0.65 (\cite{hoffman_no-u-turn_2014}).

### 2023-03-19 src/cereal_induction.jl

@@ -117,42 +117,52 @@ tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
 
 ####
 
-# Running a the above models on my real data
+# Running the above models on my real data
 
 ####
 
-analysis_data = CSV.read("data/real_data/cleaned_si_df.csv", DataFrame)
-ad_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], interaction=false)
-begin
-    ad_spread = ad_spread[:, Not([:treat2_Control, :treat1_Barley])]
-end
+include("linear_models.jl")
+include("turing_model.jl")
+analysis_data = CSV.read("data/real_data/biomass_si_data.csv", DataFrame)
 
-#To center the data run this
-begin
-    cols_list = names(ad_spread)
-    scale_vals = DataFrame()
-    ads_centered = rescalecols(df=ad_spread, collist=[:Si_ppm], centers = scale_vals)
-end
-adsc_idx = stringcoltoint(df=ad_spread, stringcol=:Genotype, intcol=:idx)
-y_vals = Float64.(adsc_idx.Si_ppm)
-preds = Matrix(adsc_idx[:, 8:13])
-idx = Int.(adsc_idx.idx)
-my_model = randomintercept_regression(y_vals, preds, idx)
-num_chains = 4
-chains = sample(my_model, NUTS(0.6), MCMCThreads(), 1_000, num_chains)
-summarystats(chains) |> DataFrame |> println
-plt = plot(chains)
-
+si_mass_model = biomass_si_regression(df=analysis_data, interaction = false)
+biomass_si_scatter = plot(analysis_data.mass_g, analysis_data.Si_ppm/10000, group = analysis_data.Species, seriestype=:scatter, smooth=true)
+plot!(biomass_si_scatter, xlabel = "Aboveground Biomass", ylabel = "Si content (%)")
+scale_vals = DataFrame()
+ads_centered = rescalecols(df=analysis_data, collist=[:Si_ppm], centers = scale_vals)
+full_model = fullcenteredglm(df=analysis_data)
+my_turing_model, center_vals = HMturingmodel(df = analysis_data, interaction = true)
+insect_model = insectonlymodel(df=analysis_data, interaction=true)
+insect_slice = analysis_data[(analysis_data.Induction .== "Insect"), :]
[…]

### 2023-03-19 src/data_compiliaton.jl

@@ -12,13 +12,30 @@ scan time by a factor of two resulted in only a 7% reduction in StDev of the mea
 =#
 
 herbivory_data = CSV.read("data/real_data/induction_data.csv", DataFrame)
-xrf_data = CSV.read("data/real_data/xrf_data.csv", DataFrame)
+xrf_data = CSV.read("data/real_data/xrf_data_86_cleaned.csv", DataFrame)
 biomass_data = CSV.read("data/real_data/biomass_no106.csv", DataFrame)
 xrf_colnames = names(xrf_data)
-xrf_si_slice = xrf_data[:, ["Si Concentration", "Sample ID"]] #select only the Si and ID columns
-rename!(xrf_si_slice, "Si Concentration" => "Si_ppm", "Sample ID" => "uniqueID") #rename so you don't have to deal with spaces
-combined_data_1 = leftjoin(herbivory_data, xrf_si_slice, on = :uniqueID) #left join to move Si into the main dataframe
-combined_data_2 = leftjoin(combined_data_1, biomass_data, on = :uniqueID)
+xrf_si_slice = xrf_data[:, ["Si Concentration", "Mn Concentration", "Sample ID"]] #select only the Si, Mn, and ID columns
+rename!(xrf_si_slice, "Si Concentration" => "Si_ppm", "Sample ID" => "uniqueID", "Mn Concentration" => "Mn_ppm") #rename so you don't have to deal with spaces
+combined_data_1 = innerjoin(herbivory_data, xrf_si_slice, on = :uniqueID) #inner join to move Si into the main dataframe
+# The biomass data had two samples for 106, as did the XRF data. Unable to tell which was which so dropp it from the data set.
+filter!(row -> row.uniqueID != 106, combined_data_1)
+combined_data_2 = innerjoin(combined_data_1, biomass_data, on = :uniqueID)
+
+function damageCat(x,y) 
+    #This function lumps the Moderate and Severe Damage Classes into one "Damaged" category
+    #and assigns "Undamaged" to all other plants in the "Insect" induction treatment
+    #and assigns <missing> to all other plants
+    if (x == "Insect" && y == "N")
+        return "Undamaged"
+    elseif (x == "Insect" && (y == "S" || y == "M"))
+        return "Damaged"
+    else
+        return missing
+    end
+end
+
+combined_data_2.isDamaged = damageCat.(combined_data_2.Induction, combined_data_2.Damage_rank)
 
[…]

### 2023-03-19 src/linear_models.jl

@@ -1,0 +1,95 @@
+using GLM
+using MixedModels
+using DataFrames
+
+
+"""
+This model is used to test for differences in the silicon content of plants in the insect treatment.
+
+It filters the inputted DataFrame (in the same format as the CSV) to only select the Insect treated
+plants.
+
+Depending on the interaction parameter, it then runs one of these two 
+linear mixed effects models: 
+
+    Si_ppm ~ Species + Damage + mass_g + (1|Genotype) #interaction = false
+
+    Si_ppm ~ Species * Damage + mass_g + (1|Genotype) #interaction = false
+
+"""
+function insectonlymodel(;df::DataFrame, interaction::Bool) 
+    df1 = filter(row -> row.Induction == "Insect", df)
+    if interaction 
+        interactivemodel = fit(MixedModel, @formula(Si_ppm ~ Species * isDamaged + mass_g + (1|Genotype)), df1)
+        return interactivemodel
+    else
+        additivemodel = fit(MixedModel, @formula(Si_ppm ~ Species + isDamaged + mass_g + (1|Genotype)), df1)
+        return additivemodel
+    end
+end
+
+"""
+This model is used to test for an effect of our induction treatments on the silicon content of the plants.
+
+It filters out the non-damaged plants in the insect treatment group, as these plants received a minimal amount of damage.
+
+It then runs a model with the following formula:
+
+    Si_ppm ~ Species * Induction + mass_g + (1|Genotype)
+"""
+function fullcenteredglm(;df::DataFrame)
+    df1 = filter(row -> (ismissing(row.isDamaged)|| row.isDamaged != "Undamaged"), df)
+    outputmodel = fit(MixedModel, @formula(Si_ppm ~ Species * Induction + mass_g + (1|Genotype)), df1)
+    return outputmodel
+end
+"""
+This model tests for a significat correlation between biomass and silicon content.
+It can test the model with/without an interaction term between the species and biomass treatments
+
+If interaction = true the following formula is used:
+
+    Si_ppm ~ mass_g * Species + (1|Genotype))
+
+If interaction = false the following formula is used:
+
+    Si_ppm ~ mass_g + Species + (1|Genotype))
+"""
+function biomass_si_regression(;df::DataFrame, interaction::Bool)
+    if interaction
+        outputmodel = fit(MixedModel, @formula(Si_ppm ~ mass_g * Species + (1|Genotype)), df)
[…]

### 2023-03-19 src/turing_model.jl

@@ -30,27 +30,27 @@ using Plots, StatsPlots;
 # σᵦ ~ Exponential(1)
 
 function spreadvars(;df::DataFrame, treat_types::Vector{Symbol}, interaction::Bool) 
-  _df = df
+  copy_df = copy(df, copycols = true)
   #spread data to wide format
   length(treat_types) > 2 ? println("Fxn can only handle two treatment types") :
   if interaction == true
-    for tr1 in unique(_df[:, treat_types[1]])
-      _df[:, "treat1_$tr1"] = ifelse.(_df[:, treat_types[1]] .== tr1, 1, 0)
-      for tr2 in unique(_df[:, treat_types[2]])
-        _df[:, "treat2_$tr2"] = ifelse.(_df[:,treat_types[2]] .== tr2, 1,0)
+    for tr1 in unique(copy_df[:, treat_types[1]])
+      copy_df[:, "treat1_$tr1"] = ifelse.(copy_df[:, treat_types[1]] .== tr1, 1, 0)
+      for tr2 in unique(copy_df[:, treat_types[2]])
+        copy_df[:, "treat2_$tr2"] = ifelse.(copy_df[:,treat_types[2]] .== tr2, 1,0)
         #Create columns for interactions
-        _df[:, "$(tr1)x$(tr2)"] = ifelse.((_df[:,treat_types[1]] .== tr1 .&& _df[:,treat_types[2]] .== tr2), 1,0 )
+        copy_df[:, "$(tr1)x$(tr2)"] = ifelse.((copy_df[:,treat_types[1]] .== tr1 .&& copy_df[:,treat_types[2]] .== tr2), 1,0 )
       end
     end
   else 
-    for tr1 in unique(_df[:, treat_types[1]])
-      _df[:, "treat1_$tr1"] = ifelse.(_df[:, treat_types[1]] .== tr1, 1, 0)
-      for tr2 in unique(_df[:, treat_types[2]])
-        _df[:, "treat2_$tr2"] = ifelse.(_df[:,treat_types[2]] .== tr2, 1,0)
+    for tr1 in unique(copy_df[:, treat_types[1]])
+      copy_df[:, "treat1_$tr1"] = ifelse.(copy_df[:, treat_types[1]] .== tr1, 1, 0)
+      for tr2 in unique(copy_df[:, treat_types[2]])
[…]

### 2023-03-19 manuscript/main.tex
fixed compilation issue

@@
 I placed one half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
-I plaed the other half of the tissue into a coin envelope, oven dried it for 4 days at 60\deg C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz).
+I plaed the other half of the tissue into a coin envelope, oven dried it for 4 days at 60$\deg$ C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz).
 After freeze drying, the samples for phenolic processing were ground under the same conditions.
@@
 To answer test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}).
-Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels (\cite{carmona_plant_2011}).
-We found a negative correlation between plant size and silicon content (\beta = -0.067 \pm 0.017, p < 0.001), and thus included plant size as a covariate in our final model.
+Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels \cite{carmona_plant_2011}.
+We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
 Thus, we tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:

### 2023-03-19 manuscript/main.tex
created summary statistics

@@
 
-To answer test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}).
+To test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}).
 Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels \cite{carmona_plant_2011}.
@@
 
-\[Si ~ Species * Induction + Biomass + (1|Genotype)\]
+\[Si \sim Species * Induction + Biomass + (1|Genotype)\]
 %[Insert text about how great these models are]. I specified a hierarchical model using the following model:
@@
 \subsection{Results}
-
-Among our cultivars, silicon content ranged from [x]\% to [y]\% of dry mass.
-[Species a] had the highest amount of silicon at [z]\%, while [species b] had the lowest silicon content at [z]\%.
+Oats had the lowest silicon content at 0.34\%.
 Overall, methyl-jasmonate and insect treatments increased plant silicon by [x] and [y]\% respectively.

### 2023-03-19 src/cereal_induction.jl

@@ -115,15 +115,32 @@ test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = 
 tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
 
 
-####
+#############
 
 # Running the above models on my real data
 
-####
+#############
 
 include("linear_models.jl")
 include("turing_model.jl")
 analysis_data = CSV.read("data/real_data/biomass_si_data.csv", DataFrame)
+
+#############
+
+# Some summary stats on the data
+
+#############
+analysis_data_noinduction = filter(row -> row.Induction == "Control", analysis_data)
+analysis_data_noinduction.Si_ppm = analysis_data_noinduction.Si_ppm./10000
+gdat_geno = groupby(analysis_data_noinduction, :Genotype)
+summ_stats_geno = combine(gdat_geno, :Si_ppm => mean, :Si_ppm => std, nrow)
+summ_stats_geno.Si_ppm_se = summ_stats_geno.Si_ppm_std ./ sqrt.(summ_stats_geno.nrow)
+CSV.write("./data/real_data/summary_statistics_genotype.csv", summ_stats_geno)
+
+gdat_spp = groupby(analysis_data_noinduction, :Species)
+summ_stats_spp = combine(gdat_spp, :Si_ppm => mean, :Si_ppm => std, nrow)
+summ_stats_spp.Si_ppm_se = summ_stats_spp.Si_ppm_std ./ sqrt.(summ_stats_spp.nrow)
+CSV.write("./data/real_data/summary_statistics_species.csv", summ_stats_spp)
 
 si_mass_model = biomass_si_regression(df=analysis_data, interaction = false)
 biomass_si_scatter = plot(analysis_data.mass_g, analysis_data.Si_ppm/10000, group = analysis_data.Species, seriestype=:scatter, smooth=true)

### 2023-03-21 manuscript/main.tex
worked on discussion

@@
 \usepackage{lineno}
+\usepackage{gensymb}
 \linenumbers
@@
 I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
-I simulated herbivory by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al.
+I simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al.
 2021b), while crickets housed in water-pik tubes provided true herbivory.
@@
 Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
+Prior to harvest, I recorded whether the crickets had initiated feeding on the plants by visual inspection of the leaves for missing tissue.
 
@@
 
-18 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e.
-no cricket exposure) leaves.
-I harvested three fully expanded upper leaves the plants, and split the plants in half along the midvein of the leaf.
-I placed one half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
-I plaed the other half of the tissue into a coin envelope, oven dried it for 4 days at 60$\deg$ C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz).
-After freeze drying, the samples for phenolic processing were ground under the same conditions.
+18 hours after treatment application, I harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
+I placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+I placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
+After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.
 
@@
 
-To test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}).
-Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels \cite{carmona_plant_2011}.
+Despite the starvation, some crickets did not initiate feeding during the exposure period.
+I filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model.
+Prior to running our full model, I first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels \cite{carmona_plant_2011}.
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
-Thus, we tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
+To test all three of our hypotheses, I used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
 
@@
 
+\subsubsection{Software used}
+
+I compiled the final dataset using \verb|DataFrames.jl| (\cite{bogumil_kaminski_2023_7632427}) in Julia 1.8.5 (\cite{bezanson2017julia}).
+I implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}).
+I tested the full mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
+I generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).
+
 \subsection{Results}
-Oats had the lowest silicon content at 0.34\%.
-Overall, methyl-jasmonate and insect treatments increased plant silicon by [x] and [y]\% respectively.
-Our model showed that both species and treatment type had effects on the plant silicon content.
+Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}).
+Amongst species, oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
+Counter to my predictions, I failed to find strong support for inducible increases in silicon content among the tested plant species.
+Despite a small p-value in the ANOVA table of the model ($p = 0.046, F = 3.14, df = 2,185.9$), the model showed only moderate support for an effect of methyl-jasmonate application ($\beta = 0.079 \pm 0.043, p = 0.070$), and no support for an effect of cricket exposure on silicon content ($\beta = 0.038 \pm 0.050, p = 0.453$).
+I found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$).
+Post-hoc tukey tests revealed that this was driven primarily by wheat's response to my induction treatments, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
+%Our model showed that both species and treatment type had effects on the plant silicon content.
 Parameter estimates and 90\% credible intervals are summarized in Table \ref{Tab:params}.
@@
 \subsection{Discussion}
+
+Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
+The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
+In this study, I attempted to demonstrate rapid silicifcation in four cereal crops.
+I failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure.
+In our study, the variation among our cultivars within species was similar in magnitude to variation among species (Figure \ref{Fig:baseline_si}), possibly obscuring the effects of our treatments.
+This study also differed from previous studies demonstrating rapid silicon uptake in a number of ways.
+The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
+In my study, I grew plants in potting soil amended with solid silicic acid.
+In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
+The theoretical maximum concentration of silicic acid in soil solution is 2 mM, however observed concentrations can be much below that.
+In my growing conditions, I applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
+Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to meet the demands of rapid silicon uptake.
 
@@
 \begin{figure}[h]
+        \includegraphics[width = \textwidth]{images/spp_si_content.png}
+        \centering
+        \caption{Baseline (uninduced) silicon content in the cereal cultivars used in this study. Cultivar species is notated in parentheses in the legend.}
+        \label{Fig:baseline_si}
+\end{figure}
+
+\begin{figure}[h]
         \includegraphics[width = \textwidth]{images/induction_plot.png}
@@
 Leaves were sampled 24 hours after treatment, and were analyzed using XRF.}
+        \label{Fig:induction}
 \end{figure}

### 2023-03-21 src/R_linear_model.R

@@ -5,10 +5,18 @@ library(lmerTest)
 library(tidyverse)
 library(emmeans)
 si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
-my_mod <- lmer(Si_ppm ~ Induction * Species + (1|Genotype), si_data)
+si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
+my_mod <- lmer(Si_ppm ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
 summary(my_mod)
 anova(my_mod)
 emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
 emmeans(my_mod, list(pairwise ~ Induction * Species), adjust = "tukey", type = "response")
-ggplot(si_data, aes(x = Induction, y = Si_ppm)) +
-geom_col()
+emmeans(my_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
+ggplot(si_filtered, aes(x = Induction, y = Si_ppm)) +
+geom_col() +
+theme_classic(base_size = 22)
+
+insect_data <- si_data[si_data$Induction == "Insect",]
+insect_data$isDamaged <- as.factor(insect_data$isDamaged)
+insect_data$Si_percent <- insect_data$Si_ppm/10000
+logit_mod <- glmer(isDamaged ~ Si_percent + Species + (1|Genotype), insect_data, family = "binomial")

### 2023-03-21 src/cereal_induction.jl

@@ -137,14 +137,25 @@ summ_stats_geno = combine(gdat_geno, :Si_ppm => mean, :Si_ppm => std, nrow)
 summ_stats_geno.Si_ppm_se = summ_stats_geno.Si_ppm_std ./ sqrt.(summ_stats_geno.nrow)
 CSV.write("./data/real_data/summary_statistics_genotype.csv", summ_stats_geno)
 
+begin
+    spp_boxplots = @df analysis_data_noinduction groupedboxplot(:Species, :Si_ppm, group = :Genotype)
+    plot!(spp_boxplots, ylabel = "Leaf Si content (%)", xlabel = "Species")
+    plot!(size = (800,600)) #width, height
+    png(spp_boxplots, "images/spp_si_content")
+end
+
 gdat_spp = groupby(analysis_data_noinduction, :Species)
 summ_stats_spp = combine(gdat_spp, :Si_ppm => mean, :Si_ppm => std, nrow)
 summ_stats_spp.Si_ppm_se = summ_stats_spp.Si_ppm_std ./ sqrt.(summ_stats_spp.nrow)
 CSV.write("./data/real_data/summary_statistics_species.csv", summ_stats_spp)
 
 si_mass_model = biomass_si_regression(df=analysis_data, interaction = false)
-biomass_si_scatter = plot(analysis_data.mass_g, analysis_data.Si_ppm/10000, group = analysis_data.Species, seriestype=:scatter, smooth=true)
-plot!(biomass_si_scatter, xlabel = "Aboveground Biomass", ylabel = "Si content (%)")
+begin
+    biomass_si_scatter = plot(analysis_data.mass_g, analysis_data.Si_ppm/10000, group = analysis_data.Species, seriestype=:scatter, smooth=true)
+    plot!(biomass_si_scatter, xlabel = "Aboveground Biomass (g)", ylabel = "Leaf Si content (%)")
+    plot!(size = (800,600))
+    png(biomass_si_scatter, "images/biomass_regression")
+end
 scale_vals = DataFrame()
 ads_centered = rescalecols(df=analysis_data, collist=[:Si_ppm], centers = scale_vals)
 full_model = fullcenteredglm(df=analysis_data)

### 2023-04-16 manuscript/main.tex
added parameter table, fixed 1st person sing.

@@
 \usepackage{gensymb}
+\usepackage{siunitx}
 \linenumbers
@@
 
-In this study, I tested four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory.
-In a glasshouse environment, I grew bread wheat (\textit{Triticum aestivum}), oats (\textit{Avena sativa}), barley (\textit{Hordeum vulgare}) and Triticale ($\times$ \textit{Triticosecale}), and tested the following hypotheses:
+In this study, we tested four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory.
+In a glasshouse environment, we grew bread wheat (\textit{Triticum aestivum}), oats (\textit{Avena sativa}), barley (\textit{Hordeum vulgare}) and Triticale ($\times$ \textit{Triticosecale}), and tested the following hypotheses:
 \begin{enumerate}
@@
 
-To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley.
-I selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
-At the start of the experiment, I germinated seeds in germination trays filled with moist sand.
-After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
-Though potting mix and fresh water contain some amount of plant available silicon, I added the silicic acid to ensure that there would be no silicon limitation to the plants.
-I randomized the location of each pot within the growing space.
+To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley.
+We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
+At the start of the experiment, we germinated seeds in germination trays filled with moist sand.
+After four days, we transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
+Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be no silicon limitation to the plants.We randomized the location of each pot within the growing space.
 A flood table bottom watered the pots with nutrient solution.
-I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
-I simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al.
+We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
+We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al.
 2021b), while crickets housed in water-pik tubes provided true herbivory.
-Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial.
-Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
-Prior to harvest, I recorded whether the crickets had initiated feeding on the plants by visual inspection of the leaves for missing tissue.
+Prior to introduction to the the plants, we acclimated crickets by feeding them on the same species used in this trial.
+Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
+Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visual inspection of the leaves for missing tissue.
 
@@
 
-18 hours after treatment application, I harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
-I placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
-I placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
+18 hours after treatment application, we harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three \SI{3.2}{\milli\metre} diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
 After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.
@@
 
-To measure the silicon content of the leaf tissue, I followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
-I pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
-I then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
-After each use, I cleaned the pellet die and XRF analyzer to minimize contamination between samples.
+To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
+We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a \SI{13}{\milli\metre} die to create a pellet.
+We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
+After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.
+
+\subsection{Phenolic analysis}
+
+To measure the response of phenolics to our treatments, we used the Fast Blue BB method \cite{pico_2020_fastblue}.
+To prepare our samples, we took \SI{0.075}{\gram} of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three \SI{3.2}{\milli\metre} chrome steel beads at 30Hz for 60 seconds.
+To the leaf powder, we added 1ml of 1\% formic acid in methanol, sonicated for 15 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2ml  microcentrifuge tube.
+We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
+To the resulting extract, we added \SI{100}{\micro\litre}
 
@@
 Despite the starvation, some crickets did not initiate feeding during the exposure period.
-I filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model.
-Prior to running our full model, I first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels \cite{carmona_plant_2011}.
+We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model.
+Prior to running our full model, we first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels \cite{carmona_plant_2011}.
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
-To test all three of our hypotheses, I used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
+To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
 
@@
 
-I compiled the final dataset using \verb|DataFrames.jl| (\cite{bogumil_kaminski_2023_7632427}) in Julia 1.8.5 (\cite{bezanson2017julia}).
-I implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}).
-I tested the full mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
-I generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).
+We compiled the final dataset using \verb|DataFrames.jl| (\cite{bogumil_kaminski_2023_7632427}) in Julia 1.8.5 (\cite{bezanson2017julia}).
+We implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}).
+We tested the full mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
+We generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).
 
@@
 Amongst species, oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
-Counter to my predictions, I failed to find strong support for inducible increases in silicon content among the tested plant species.
+Counter to my predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species.
 Despite a small p-value in the ANOVA table of the model ($p = 0.046, F = 3.14, df = 2,185.9$), the model showed only moderate support for an effect of methyl-jasmonate application ($\beta = 0.079 \pm 0.043, p = 0.070$), and no support for an effect of cricket exposure on silicon content ($\beta = 0.038 \pm 0.050, p = 0.453$).
-I found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$).
+We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$).
 Post-hoc tukey tests revealed that this was driven primarily by wheat's response to my induction treatments, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
@@
 The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
-In this study, I attempted to demonstrate rapid silicifcation in four cereal crops.
-I failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure.
+In this study, we attempted to demonstrate rapid silicifcation in four cereal crops.
+We failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure.
 In our study, the variation among our cultivars within species was similar in magnitude to variation among species (Figure \ref{Fig:baseline_si}), possibly obscuring the effects of our treatments.
@@
 The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
-In my study, I grew plants in potting soil amended with solid silicic acid.
+In my study, we grew plants in potting soil amended with solid silicic acid.
 In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
-The theoretical maximum concentration of silicic acid in soil solution is 2 mM, however observed concentrations can be much below that.
-In my growing conditions, I applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
+The theoretical maximum concentration of silicic acid in soil solution is \SI{2}{\milli\mole}, however observed concentrations can be much below that.
+In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
 Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to meet the demands of rapid silicon uptake.
@@
 
-\begin{table}[h]
+\begin{table}[ht]
         \centering
-        \caption{Credible Interval and Parameter estimates for the hierarchical model. Parameters are estimated against a baseline of Induction: None and Species: Barley.}
+        \caption{\verb{emmeans} results of pairwise comparisons between groups.
+p-values are tested using the \verb{mvt} adjustment, against the  multivariate normal distribution, a less conservative approach than typical bonferroni corrections.}
         \label{Tab:params}
-        \begin{tabular} { c | c | c }
+        \begin{tabular}{lrrrrr}
             \hline
-                Parameter & Credible Interval & Parameter Estimate \\
+            contrast & estimate & SE & df & t.ratio & p.value \\
             \hline
-                Insect & -4.200 -- -4.100 & -4.108 \\
-                Methyl Jasmonate & -4.20 -- -4.15 & -4.17 \\
-                Oats & -5.00 -- -4.95 & -4.97 \\
-                Wheat & -4.60 -- -4.5 & -4.51 \\
-                Triticale & -4.13 -- -4.08 & -4.11 \\
-                \hline
+            Control Barley - Insect Barley & -373.9 & 514 & 186.37 & -0.727 & 0.9997 \\
+            Control Barley - MeJA Barley & -793.4 & 449 & 186.05 & -1.769 & 0.7656 \\
+            Control Barley - Control Oats & 723.0 &1126 & 10.08 & 0.642 & 0.9997 \\
+            Control Barley - Insect Oats & 678.6 &1169 & 11.68 & 0.581 & 0.9999 \\
+            Control Barley - MeJA Oats & 215.6 &1127 & 10.12 & 0.191 & 1.0000 \\
+            Control Barley - Control Triticale &-2321.0 &1155 & 11.08 & -2.009 & 0.6139 \\
+            Control Barley - Insect Triticale &-2516.2 &1185 & 12.19 & -2.124 & 0.5467 \\
+            Control Barley - MeJA Triticale &-2813.2 &1151 & 10.94 & -2.445 & 0.3819 \\
+            Control Barley - Control Wheat &-3485.7 &1136 & 10.43 & -3.068 & 0.1735 \\
+            Control Barley - Insect Wheat &-1638.1 &1149 & 10.92 & -1.425 & 0.9054 \\
+            Control Barley - MeJA Wheat &-3044.7 &1130 & 10.20 & -2.695 & 0.2854 \\
+            Insect Barley - MeJA Barley & -419.5 & 508 & 186.31 & -0.825 & 0.9992 \\
+            Insect Barley - Control Oats & 1096.9 &1148 & 10.86 & 0.956 & 0.9928 \\
+            Insect Barley - Insect Oats & 1052.5 &1190 & 12.55 & 0.884 & 0.9963 \\
+            Insect Barley - MeJA Oats & 589.5 &1149 & 10.91 & 0.513 & 1.0000 \\
+            Insect Barley - Control Triticale &-1947.0 &1175 & 11.85 & -1.657 & 0.8083 \\
+            Insect Barley - Insect Triticale &-2142.3 &1202 & 12.94 & -1.782 & 0.7439 \\
+            Insect Barley - MeJA Triticale &-2439.2 &1169 & 11.68 & -2.086 & 0.5702 \\
+            Insect Barley - Control Wheat &-3111.8 &1156 & 11.16 & -2.693 & 0.2775 \\
+            Insect Barley - Insect Wheat &-1264.1 &1169 & 11.68 & -1.081 & 0.9824 \\
+            Insect Barley - MeJA Wheat &-2670.8 &1150 & 10.97 & -2.322 & 0.4421 \\
+            MeJA Barley - Control Oats & 1516.4 &1120 & 9.87 & 1.354 & 0.9262 \\
+            MeJA Barley - Insect Oats & 1472.0 &1164 & 11.49 & 1.265 & 0.9514 \\
+            MeJA Barley - MeJA Oats & 1009.0 &1122 & 9.92 & 0.900 & 0.9951 \\
+            MeJA Barley - Control Triticale &-1527.6 &1148 & 10.83 & -1.330 & 0.9350 \\
+            MeJA Barley - Insect Triticale &-1722.8 &1177 & 11.88 & -1.464 & 0.8932 \\
+            MeJA Barley - MeJA Triticale &-2019.8 &1143 & 10.66 & -1.767 & 0.7490 \\
+            MeJA Barley - Control Wheat &-2692.4 &1129 & 10.17 & -2.385 & 0.4163 \\
+            MeJA Barley - Insect Wheat & -844.7 &1142 & 10.66 & -0.739 & 0.9990 \\
+            MeJA Barley - MeJA Wheat &-2251.3 &1123 & 9.97 & -2.004 & 0.6186 \\
+            Control Oats - Insect Oats & -44.4 & 552 & 186.14 & -0.080 & 1.0000 \\
+            Control Oats - MeJA Oats & -507.4 & 448 & 186.03 & -1.133 & 0.9870 \\
+            Control Oats - Control Triticale &-3043.9 &1136 & 10.40 & -2.679 & 0.2909 \\
+            Control Oats - Insect Triticale &-3239.2 &1158 & 11.17 & -2.798 & 0.2414 \\
+            Control Oats - MeJA Triticale &-3536.1 &1126 & 10.08 & -3.140 & 0.1563 \\
+            Control Oats - Control Wheat &-4208.7 &1114 & 9.65 & -3.779 & 0.0707 \\
+            Control Oats - Insect Wheat &-2361.1 &1129 & 10.16 & -2.092 & 0.5698 \\
+            Control Oats - MeJA Wheat &-3767.7 &1112 & 9.60 & -3.387 & 0.1198 \\
+            Insect Oats - MeJA Oats & -463.0 & 557 & 186.23 & -0.831 & 0.9991 \\
+            Insect Oats - Control Triticale &-2999.5 &1185 & 12.26 & -2.532 & 0.3359 \\
+            Insect Oats - Insect Triticale &-3194.8 &1207 & 13.17 & -2.646 & 0.2826 \\
+            Insect Oats - MeJA Triticale &-3491.7 &1176 & 11.97 & -2.968 & 0.1818 \\
+            Insect Oats - Control Wheat &-4164.3 &1164 & 11.49 & -3.578 & 0.0764 \\
+            Insect Oats - Insect Wheat &-2316.6 &1178 & 12.03 & -1.967 & 0.6373 \\
+            Insect Oats - MeJA Wheat &-3723.3 &1161 & 11.39 & -3.206 & 0.1339 \\
+            MeJA Oats - Control Triticale &-2536.5 &1139 & 10.49 & -2.228 & 0.4960 \\
+            MeJA Oats - Insect Triticale &-2731.8 &1161 & 11.28 & -2.354 & 0.4264 \\
+            MeJA Oats - MeJA Triticale &-3028.7 &1129 & 10.17 & -2.683 & 0.2894 \\
+            MeJA Oats - Control Wheat &-3701.3 &1116 & 9.74 & -3.315 & 0.1322 \\
+            MeJA Oats - Insect Wheat &-1853.6 &1131 & 10.26 & -1.639 & 0.8139 \\
+            MeJA Oats - MeJA Wheat &-3260.3 &1115 & 9.69 & -2.924 & 0.2210 \\
+            Control Triticale - Insect Triticale & -195.3 & 559 & 186.25 & -0.350 & 1.0000 \\
+            Control Triticale - MeJA Triticale & -492.2 & 512 & 186.60 & -0.961 & 0.9967 \\
+            Control Triticale - Control Wheat &-1164.8 &1136 & 10.39 & -1.026 & 0.9872 \\
+            Control Triticale - Insect Wheat & 682.9 &1151 & 10.93 & 0.593 & 0.9999 \\
+            Control Triticale - MeJA Wheat & -723.8 &1136 & 10.38 & -0.637 & 0.9997 \\
+            Insect Triticale - MeJA Triticale & -296.9 & 548 & 187.32 & -0.542 & 1.0000 \\
+            Insect Triticale - Control Wheat & -969.5 &1154 & 11.05 & -0.840 & 0.9974 \\
+            Insect Triticale - Insect Wheat & 878.1 &1170 & 11.62 & 0.751 & 0.9990 \\
+            Insect Triticale - MeJA Wheat & -528.5 &1156 & 11.11 & -0.457 & 1.0000 \\
+            MeJA Triticale - Control Wheat & -672.6 &1124 & 10.00 & -0.598 & 0.9999 \\
+            MeJA Triticale - Insect Wheat & 1175.1 &1139 & 10.54 & 1.031 & 0.9866 \\
+            MeJA Triticale - MeJA Wheat & -231.5 &1125 & 10.03 & -0.206 & 1.0000 \\
+            Control Wheat - Insect Wheat & 1847.7 & 479 & 186.44 & 3.855 & 0.0057 \\
+            Control Wheat - MeJA Wheat & 441.0 & 443 & 186.04 & 0.995 & 0.9956 \\
+            Insect Wheat - MeJA Wheat &-1406.6 & 480 & 186.45 & -2.931 & 0.0982 \\
         \end{tabular}
@@
 Along with Si and P mobilization, Mn is often released, and taken up by plants roots.
-Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far I are unaware of any studies looking for quantitative variation among genotypes of leaf Mn.
+Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far we are unaware of any studies looking for quantitative variation among genotypes of leaf Mn.
 If we could identify regions of the plant genome associated with variation in root weathering activity, we may be able to target this trait in breeding programs that improve nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
@@
 Though XRF is an established technique to measure plant Si, its may also be used to measure other metals of interest, including manganese.
-In this study I use XRF to quantify variation in Si and Mn content among a diversity panel of a wild ancestor of bread wheat, Aegilops tauschii.
+In this study we use XRF to quantify variation in Si and Mn content among a diversity panel of a wild ancestor of bread wheat, Aegilops tauschii.
 This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link Si and Mn variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
@@
 
-For this experiment, I used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites.
-Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where I vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment.
+For this experiment, we used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites.
+Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment.
 For full site details see Supplementary Table S1.
-Using 151 accessions, I started trays of seedlings in glasshouse or growth chamber environments.
+Using 151 accessions, we started trays of seedlings in glasshouse or growth chamber environments.
 At approximately eight weeks after germination, seedlings were transplanted to their field sites.
-For each environment, I started four replicates of each accession.
-I planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements.
+For each environment, we started four replicates of each accession.
+We planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements.
 Each outdoor block was a 16 m\textsuperscript{2} square, with plants arranged $\sim$35 cm apart.
-Shortly after transplanting to the field sites, I applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets.
+Shortly after transplanting to the field sites, we applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets.
 Field transplantation took place on the 15th of October 2022 and the 16th of December 2022.
-For the glasshouse environment, I started seedlings in growth chambers in January 2022.
-After 12 weeks, I moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks.
-I then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with [amount] of silicic acid (Tixosil 68B, Solvay).
+For the glasshouse environment, we started seedlings in growth chambers in January 2022.
+After 12 weeks, we moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks.
+We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with [amount] of silicic acid (Tixosil 68B, Solvay).
 Pots were arranged using the same randomized block design, adapted to fit on two flood tables.
@@
 
-When the plants had reached maturity, I harvested the entire above-ground portion of each plant.
-For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while I harvested the glasshouse plants between the 19th and 21st of September 2022.
-I placed harvested material in labelled paper bags, and dried it in drying ovens at 60ºC for 48 hours.
-To harvest leaf material for analysis, I selected stems with flower heads, and removed the three leaves closest to the flowers.
-Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), I chose a consistent set of leaves to minimize introduced variation.
-I picked leaves until approximately 200mg of dry leaf was collected.
-I then washed leaves in distilled water to remove any soil residues which might introduce silicon, and re-dried the samples at 60ºC for 48 hours.
-Dried, clean leaves were then packed into 2 ml microcentrifuge tubes with zircon grinding pellets, flash frozen by immersion in liquid nitrogen, and ground in a tissuelyser ball mill for 30 seconds at 60 Hz.
-The resulting leaf powder was stored sealed until XRF analysis.
+When the plants had reached maturity, we harvested the entire above-ground portion of each plant.
+For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022.
+We placed harvested material in labelled paper bags, and dried it in drying ovens at 60ºC for 48 hours.
+To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers.
+Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation.
+We picked leaves until approximately 200mg of dry leaf was collected.
+Some plants did not yield enough leaf tissue to meet to 200mg threshold.
+To reduce costs and increase the amount of biomass availabe per genotype, we pooled leaf material from within sites.
+From each individual, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
+We packed dried leaves samples were 2 ml microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
+We stored the resulting leaf powder sealed until XRF analysis.
 
@@
 
-To analyse the silicon and manganese content of the accessions, I followed the XRF procedure presented in Reidinger et al.
+To analyse the silicon and manganese content of the accessions, we followed the XRF procedure presented in Reidinger et al.
 (\citedate{reidinger_rapid_2012}).
-In short, I pressed leaf powder into 13mm diameter pellets at 11 tonnes of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
-I used a read time of 30s to ensure accurate measurements.
-To minimize cross-contamination between samples, I wiped clean the pellet press and XRF device after each sample.
+In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
+For beam 1 (Mn), we used a 20 second read time, while for beam 2 (Si), we used a 50 second read time.
+Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy.
+For each pellet, we took two technical replicates, scanning each side of the pellet once.
+To minimize cross-contamination between samples, we cleaned the pellet press and XRF device after each sample.
+We calibrated our measurments against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).
 
@@
 
-To perfrom the GWAS analysis, I followed the methodology and code published in Gaurav et al.
+To prepare our data for 
+To perfrom the GWAS analysis, we followed the methodology and code published in Gaurav et al.
 \citedate{gaurav_population_2021}.
-For brevity, this methodology only describes the steps I took using the data generated from Gaurav et al.
+For brevity, this methodology only describes the steps we took using the data generated from Gaurav et al.
 \citedate{gaurav_population_2021}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
-Prior to the linear regression to assess genetic associations, I first tested my phenotypic data for block effects within sites.
-For each site, we tested a linear model to see if measured leaf silicon content was significantly associated with block.
-I tested the model $Si \sim Block + Accession$.
-When I validated that block had a non-significant effect ($p>0.1$), I reparameterized leaf silicon measurements at the site level; I converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si.
-I did this to eliminate a possible confounding effect of inter-site differences that might reduce the power of the GWAS linear regression.
-
 As per Gaurav et al.
-\citedate{gaurav_population_2021}, to reduce the computational intensity of my analysis, I prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative [word choice?].
+\citedate{gaurav_population_2021}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
+To prepare our
 

### 2023-04-16 src/R_linear_model.R

@@ -1,16 +1,21 @@
 install.packages("lmerTest")
 install.packages("tidyverse")
 install.packages("emmeans")
+install.packages("xtable")
 library(lmerTest)
 library(tidyverse)
+library(xtable)
 library(emmeans)
 si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
 my_mod <- lmer(Si_ppm ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
 summary(my_mod)
 anova(my_mod)
-emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
-emmeans(my_mod, list(pairwise ~ Induction * Species), adjust = "tukey", type = "response")
+emgrid1  <- emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
+emgrid2 <- emmeans(my_mod,  ~ Induction * Species, adjust = "mvt", type = "response")
+my_emm <- pairs(emmeans(my_mod, ~ Induction * Species))
+foo <- pwpm(my_emm, by = NULL, adjust = "bonferroni")
+xtable::xtable(foo)
 emmeans(my_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
 ggplot(si_filtered, aes(x = Induction, y = Si_ppm)) +
 geom_col() +
@@ -20,3 +25,14 @@ insect_data <- si_data[si_data$Induction == "Insect",]
 insect_data$isDamaged <- as.factor(insect_data$isDamaged)
 insect_data$Si_percent <- insect_data$Si_ppm/10000
 logit_mod <- glmer(isDamaged ~ Si_percent + Species + (1|Genotype), insect_data, family = "binomial")
+summary(logit_mod)
+anova(logit_mod)
+
+pigsint.lm <- lm(log(conc) ~ source * factor(percent), data = pigs)
+pigs.sav <- as.list(ref_grid(pigsint.lm))
+pigs.anew <- as.emmGrid(pigs.sav)
+pigsint.emm <- pairs(emmeans(pigsint.lm, ~ percent | source))
+test(pigsint.emm, by = NULL, adjust = "bonferroni")
+typeof(pigsint.emm)
+xtable::xtable(pigsint.emm, type = "response")
+

### 2023-04-18 manuscript/main.tex
more cleaning

@@
+\listfiles
 \documentclass[12pt, letterpaper, ]{article}
@@
 \usepackage{lineno}
-\usepackage{gensymb}
-\usepackage{siunitx}
+
 \linenumbers
@@
 18 hours after treatment application, we harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three \SI{3.2}{\milli\metre} diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three %\SI{3.2}{\milli\metre} diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
@@
 To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
-We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a \SI{13}{\milli\metre} die to create a pellet.
+We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
 We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
@@
 To measure the response of phenolics to our treatments, we used the Fast Blue BB method \cite{pico_2020_fastblue}.
-To prepare our samples, we took \SI{0.075}{\gram} of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three \SI{3.2}{\milli\metre} chrome steel beads at 30Hz for 60 seconds.
+To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three %\SI{3.2}{\milli\metre} chrome steel beads at 30Hz for 60 seconds.
 To the leaf powder, we added 1ml of 1\% formic acid in methanol, sonicated for 15 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2ml  microcentrifuge tube.
@@
 In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
-The theoretical maximum concentration of silicic acid in soil solution is \SI{2}{\milli\mole}, however observed concentrations can be much below that.
+The theoretical maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that.
 In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
@@
         \centering
-        \caption{\verb{emmeans} results of pairwise comparisons between groups.
-p-values are tested using the \verb{mvt} adjustment, against the  multivariate normal distribution, a less conservative approach than typical bonferroni corrections.}
+        \caption{emmeans results of pairwise comparisons between groups.
+p-values are tested using the  adjustment, against the  multivariate normal distribution, a less conservative approach than typical bonferroni corrections.}
         \label{Tab:params}

### 2023-04-19 manuscript/main.tex
updated document with param estimates

@@
 \usepackage{lineno}
+\usepackage{gensymb}
+\usepackage{longtable}
 
@@
 18 hours after treatment application, we harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three %\SI{3.2}{\milli\metre} diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
@@
 
+We analysed phenolics as a way to validate the effectiveness of our methyl-jasmonate applications on defense induction.
 To measure the response of phenolics to our treatments, we used the Fast Blue BB method \cite{pico_2020_fastblue}.
-To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three %\SI{3.2}{\milli\metre} chrome steel beads at 30Hz for 60 seconds.
-To the leaf powder, we added 1ml of 1\% formic acid in methanol, sonicated for 15 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2ml  microcentrifuge tube.
+To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30Hz for 60 seconds.
+To the leaf powder, we added 1ml of 1\% formic acid in methanol, sonicated for 15 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
 We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
-To the resulting extract, we added \SI{100}{\micro\litre}
+To the resulting extract, we added 100$\mu$L of Fast Blue BB solution, allowed the color to develop in dark conditions for one hour, then used a 96-well plate to read the absorbance of [XXX] nm light.
+We compared the readings to a standard curve created using gallic acid.
 
@@
         \centering
+        \caption{ANOVA table for our linear mixed effects model analyzing the effect of defense induction and species identity on leaf silicon content.
+We generated the ANOVA table using the R package car, specifying a type III ANOVA.}
+        \label{Tab:params}
+        \begin{tabular}{lrrrr}
+                \hline
+                \textbf{Effect} & \textbf{F} & \textbf{Df} & \textbf{Df residual} & \textbf{\textit{P} value} \\
+                \hline
+                (Intercept) & 32.4990 & 1 & 24.350 & 6.775e-06*** \\
+                Induction & 1.5671 & 2 & 186.230 & 0.21139 \\   
+                Species & 6.0830 & 3 & 10.308 & 0.01199* \\ 
+                Biomass & 16.5099 & 1 & 190.074 & 7.076e-05*** \\
+                Induction*Species & 2.2514 & 6 & 186.352 & 0.04027* \\
+                \hline
+        \end{tabular}
+\end{table}
+
+\begin{center}
+\begin{longtable}{lrrrrr}
         \caption{emmeans results of pairwise comparisons between groups.
 p-values are tested using the  adjustment, against the  multivariate normal distribution, a less conservative approach than typical bonferroni corrections.}
-        \label{Tab:params}
-        \begin{tabular}{lrrrrr}
+        \label{Tab:emm_params}
+            \\ \hline
+            \textbf{Contrast} & \textbf{Estimate} & \textbf{SE} & \textbf{df} & \textbf{\textit{t} ratio} & \textbf{\textit{p} value} 
+            \\ \hline
+            \endfirsthead
+            \multicolumn{3}{c}{{\bfseries \tablename\ \thetable{} -- continued from previous page}}
+            \\ \hline
+            \textbf{Contrast} & \textbf{Estimate} & \textbf{SE} & \textbf{df} & \textbf{\textit{t} ratio} & \textbf{\textit{p} value} 
+            \\ \hline
+            \endhead
             \hline
-            contrast & estimate & SE & df & t.ratio & p.value \\
+            \endfoot
             \hline
+            \hline
+            \endlastfoot
+
             Control Barley - Insect Barley & -373.9 & 514 & 186.37 & -0.727 & 0.9997 \\
@@
             Insect Wheat - MeJA Wheat &-1406.6 & 480 & 186.45 & -2.931 & 0.0982 \\
-        \end{tabular}
-\end{table}
+\end{longtable}
+\end{center}
 
@@
         \caption{The effects of crop species and induction treatment on leaf silicon content.
-Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets \textit{Acheta domesticus}.
-Leaves were sampled 24 hours after treatment, and were analyzed using XRF.}
+Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
+Leaves were sampled 18 hours after treatment, and were analyzed using XRF.}
         \label{Fig:induction}
@@
 To reduce costs and increase the amount of biomass availabe per genotype, we pooled leaf material from within sites.
-From each individual, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
-We packed dried leaves samples were 2 ml microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
+For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
+Overall, we were left with approximately 115 useable genotypes from each site.
+We packed dried leaves samples were 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
 We stored the resulting leaf powder sealed until XRF analysis.
@@
 In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
-For beam 1 (Mn), we used a 20 second read time, while for beam 2 (Si), we used a 50 second read time.
+For beam 1 (Mn), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time.
 Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy.
@@
 
-To prepare our data for 
+Need to revisit this and figure out exactly what to say.
+
 To perfrom the GWAS analysis, we followed the methodology and code published in Gaurav et al.
@@
 Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis.
-Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{value}]\% to [\textit{value}\%].
-Site accounted for a large amount of variation in silicon content, suggesting a strong environmental effect on the silicon phenotype.
+After pooling, we were left with 359 samples across three environments.
+Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{0.784}]\% to [\textit{11.473}\%].
+There were notable differences in silicon content based on the growing environment.
+The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
 Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}).

### 2023-04-19 src/R_linear_model.R

@@ -2,15 +2,18 @@ install.packages("lmerTest")
 install.packages("tidyverse")
 install.packages("emmeans")
 install.packages("xtable")
-library(lmerTest)
+install.packages("car")
+library(lme4)
 library(tidyverse)
-library(xtable)
+#library(xtable)
 library(emmeans)
+library(car)
 si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
-my_mod <- lmer(Si_ppm ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
+my_mod <- lme4::lmer(Si_ppm ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
 summary(my_mod)
 anova(my_mod)
+Anova(my_mod, type ="III", test.statistic = "F")
 emgrid1  <- emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
 emgrid2 <- emmeans(my_mod,  ~ Induction * Species, adjust = "mvt", type = "response")
 my_emm <- pairs(emmeans(my_mod, ~ Induction * Species))

### 2023-04-25 manuscript/main.tex
another draft version sent to juli

@@
 \listfiles
-\documentclass[12pt, letterpaper, ]{article}
+\documentclass[12pt, letterpaper, ]{report}
 \usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
@@
 \usepackage{longtable}
+\usepackage{titlesec}
+
+% Change chapter heading font size to 14pt
+\titleformat{\chapter}{\Large\bfseries}{\thechapter.}{20pt}{\Large}
+
+% Change section heading font size to 12pt
+\titleformat{\section}{\large\bfseries}{\thesection}{1em}{\large}
+
+% Change subsection heading font size to 10pt
+\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{\normalsize}
 
@@
 
-\section{Abstract}
+\chapter{Abstract}
 
@@
 
-\section{Lay Summary}
+\chapter{Lay Summary}
 
@@
 
-\section{Preface}
+\chapter{Preface}
 
@@
 
-\section{Introduction}
+\chapter{Introduction}
 
-\subsection{A case for silicon in agriculture}
+\section{A case for silicon in agriculture}
 
@@
 
-\subsection{Silicon in nature}
+\section{Silicon in nature}
 
@@
 
-\subsection{Silicon in soils and roots}
+\section{Silicon in soils and roots}
 
@@
 
-\subsection{Silicon transporters}
+\section{Silicon transporters}
 
@@
 
-\subsection{Silicon in leaves}
+\section{Silicon in leaves}
 
@@
 
-\subsection{Looking back and looking forward}
+\section{Looking back and looking forward}
 
@@
 
-\section{Chapter 1: Identifying rapid silicon accumulation in cereal crops}
+\printbibliography
 
-\subsection{Introduction}
+\chapter{Chapter 1: Identifying rapid silicon accumulation in cereal crops}
+
+\section{Introduction}
 
@@
 
-\subsection{Methods}
+\section{Methods}
 
-\subsubsection{Plant growth and experimental treatments}
+\subsection{Plant growth and experimental treatments}
 
@@
 
-\subsubsection{Sample harvest and preparation}
+\subsection{Sample harvest and preparation}
 
@@
 
-\subsubsection{Silicon analysis}
+\subsection{Silicon analysis}
 
@@
 
-\subsection{Phenolic analysis}
+\section{Phenolic analysis}
 
 We analysed phenolics as a way to validate the effectiveness of our methyl-jasmonate applications on defense induction.
-To measure the response of phenolics to our treatments, we used the Fast Blue BB method \cite{pico_2020_fastblue}.
+To measure the response of phenolics to our treatments, we used the Fast Blue BB method (\cite{pico2020systematic}).
 To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30Hz for 60 seconds.
@@
 
-\subsubsection{Statistical analysis}
+\subsection{Statistical analysis}
 
@@
 We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model.
-Prior to running our full model, we first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels \cite{carmona_plant_2011}.
+Prior to running our full model, we first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels (\cite{carmona_plant_2011}).
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
@@
 
-\subsubsection{Software used}
+\subsection{Software used}
 
@@
 
-\subsection{Results}
+\section{Results}
 Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}).
@@
 
-\subsection{Discussion}
+\section{Discussion}
 
@@
 The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
-In my study, we grew plants in potting soil amended with solid silicic acid.
+In my study, we grew plants in potting soil amended with solid silicic acid, similar to Nacimento et al.
+(\citedate{nascimento_silicon_2019}).
 In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
@@
 In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
-Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to meet the demands of rapid silicon uptake.
+Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.
 
-\subsection{Acknowledgements}
+\section{Acknowledgements}
 
-\subsection{Data Availability}
+\section{Data Availability}
 
-\subsection{Figures and Tables}
+\printbibliography
+
+\section{Figures and Tables}
 
@@
 
-\section{Chapter 2: Genetic drivers of silicon accumulation in a wild ancestor of wheat}
+\chapter{Chapter 2: Genetic drivers of silicon accumulation in a wild ancestor of wheat}
 
-\subsection{Introduction}
+\section{Introduction}
 
@@
 
-\subsection{Methods}
+\section{Methods}
 
-\subsubsection{Plant growing conditions}
+\subsection{Plant growing conditions}
 
@@
 
-\subsubsection{Plant harvest and sample preparation}
+\subsection{Plant harvest and sample preparation}
 
@@
 
-\subsubsection{Sample analysis}
+\subsection{Sample analysis}
 
-To analyse the silicon and manganese content of the accessions, we followed the XRF procedure presented in Reidinger et al.
-(\citedate{reidinger_rapid_2012}).
+To analyse the silicon and manganese content of the accessions, we followed the XRF procedure presented in \textcite{reidinger_rapid_2012}.
 In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
@@
 
-To perfrom the GWAS analysis, we followed the methodology and code published in Gaurav et al.
-\citedate{gaurav_population_2021}.
-For brevity, this methodology only describes the steps we took using the data generated from Gaurav et al.
-\citedate{gaurav_population_2021}.
+To look for evidence of root exudation driving silicon content, we tested a correlation between observed leaf silicon and manganese content.
+Prior to the analysis, we plotted histograms of silicon and manganese concentrations.
+We observed right-skew for both elements and applied a log-transformation.
+To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$).
+We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:
+
+$Manganese ~ Silicon + (1|Genotype)$
+
+To perfrom the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
+For brevity, this methodology only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
-As per Gaurav et al.
-\citedate{gaurav_population_2021}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
-To prepare our
+As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
+For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers.
+We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.
 
-\subsection{Results}
+\section{Results}
 
@@
 After pooling, we were left with 359 samples across three environments.
-Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{0.784}]\% to [\textit{11.473}\%].
+Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%.
 There were notable differences in silicon content based on the growing environment.
 The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
-Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}).
-One of these genomes was on chromosome 4S, near a known gene analogue to \textit{Lsi1}, a silicon transporter protein.
+Silicon and manganese showed significant positive correlation ($\beta = 0.278\pm0.051, z = 5.46, p<0.0001$) (Figure \ref{Fig:mn_si_regression}).
+Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}).
 My results for manganese content are less clear.
@@
 
-\subsection{Discussion}
+\section{Discussion}
 
-\subsection{Acknowledgements}
+We observed major differences in silicon content between our growth environments, and did not find conistent correlations between genetic variation and variation in silicon content.
+Despite this, we did find a significant positive correlation between silicon and manganese content, supporting root exudates increasing silicon uptake.
 
-\subsection{Data Availability}
+\section{Acknowledgements}
 
-\subsection{Tables and Figures}
+\section{Data Availability}
+
+\printbibliography
+
+\section{Tables and Figures}
+
+\begin{figure}[h]
+        \includegraphics[scale=0.048]{images/si_mn_regression.png}
+        \centering
+        \caption{Scatter plot comparing leaf silicon and manganese content. Values for both elements were converted to standard scores to allow
+        comparison between environments. Points are color coded according to environment. GH is the glasshouse environment, while SC and TF were two different outdoor plots.}
+        \label{Fig:mn_si_regression}
+\end{figure}
 
@@
 \end{figure}
+
 \begin{figure}[h]
@@
 
-\begin{figure}[h]
-        \includegraphics{images/si_mn_regression.png}
-        \centering
-        \caption{This is the regression comparing Si to Mn content in the leaf tissue}
-        \label{Fig:mn_si_regression}
-\end{figure}
-
 \clearpage
 
-\section{References}
-\printbibliography
-
 \end{document}

### 2023-04-28 src/R_linear_model.R
more data more plots!

@@ -3,23 +3,25 @@ install.packages("tidyverse")
 install.packages("emmeans")
 install.packages("xtable")
 install.packages("car")
+install.packages("Hmisc")
 library(lme4)
 library(tidyverse)
 #library(xtable)
 library(emmeans)
 library(car)
-si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
+library(Hmisc)
+si_data <- as.data.frame(read.csv("data/real_data/si_absorbance_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
-my_mod <- lme4::lmer(Si_ppm ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
-summary(my_mod)
-anova(my_mod)
-Anova(my_mod, type ="III", test.statistic = "F")
-emgrid1  <- emmeans(my_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
-emgrid2 <- emmeans(my_mod,  ~ Induction * Species, adjust = "mvt", type = "response")
-my_emm <- pairs(emmeans(my_mod, ~ Induction * Species))
+si_mod <- lme4::lmer(Si_ppm/10000 ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
+summary(si_mod)
+anova(si_mod)
+Anova(si_mod, type ="III", test.statistic = "F")
+emgrid1  <- emmeans(si_mod, list(pairwise ~ Species), adjust = "tukey", type = "response")
+emgrid2 <- emmeans(si_mod,  ~ Induction * Species, adjust = "mvt", type = "response")
+my_emm <- pairs(emmeans(si_mod, ~ Induction * Species))
 foo <- pwpm(my_emm, by = NULL, adjust = "bonferroni")
 xtable::xtable(foo)
-emmeans(my_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
+emmeans(si_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
 ggplot(si_filtered, aes(x = Induction, y = Si_ppm)) +
 geom_col() +
 theme_classic(base_size = 22)
@@ -27,7 +29,7 @@ theme_classic(base_size = 22)
 insect_data <- si_data[si_data$Induction == "Insect",]
 insect_data$isDamaged <- as.factor(insect_data$isDamaged)
 insect_data$Si_percent <- insect_data$Si_ppm/10000
[…]

### 2023-04-28 src/cereal_induction.jl

@@ -114,7 +114,7 @@ png(mp, "induction_plot")
 test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = repeat(["D", "E", "F"], inner = 9, outer = 1), yvals = rand(Normal(0,1), 27))
 tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
 
-
+more_data = CSV.read("data/")
 #############
 
 # Running the above models on my real data
@@ -156,6 +156,8 @@ begin
     plot!(size = (800,600))
     png(biomass_si_scatter, "images/biomass_regression")
 end
+
+
 scale_vals = DataFrame()
 ads_centered = rescalecols(df=analysis_data, collist=[:Si_ppm], centers = scale_vals)
 full_model = fullcenteredglm(df=analysis_data)
@@ -194,3 +196,38 @@ ad_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], inte
     chains = sample(my_model, NUTS(0.6), MCMCThreads(), 1_000, num_chains)
     summarystats(chains) |> DataFrame |> println
     plot(chains)
+
+phenolic_data = CSV.read("data/real_data/si_absorbance_data.csv", DataFrame)
+
+gboxplot_ab = @df phenolic_data groupedboxplot(:Induction, :mcabsorbance, group = :Species)
+
+ph_si_scatter = plot(phenolic_data.Si_ppm/10000, phenolic_data.mcabsorbance, group=phenolic_data.Species, seriestype=:scatter)
+plot!(ph_si_scatter, xlabel = "Leaf Silicon Content (%)", ylabel = "Leaf Phenolic Content")
+plot!(size = (800,600), dpi = 800)
+png(ph_si_scatter, "manuscript/images/phenolic_silicon_regression")
+
+function standarderror(col) 
+    sd = std(col)
+    denom = sqrt(length(col)-1)
+    se = sd/denom
+    return se
+end
+
+
+gph = groupby(phenolic_data, [:Species])
+test = combine(gph, [:mcabsorbance, :Si_ppm] => ((a, s) -> (sppMEANabsorbance = mean(a), sppSEabsorbance = standarderror(a),sppMEANsi = mean(s), sppSEsi = standarderror(s))) => AsTable)
+
+using Measurements
+
[…]

### 2023-04-28 src/merging_data.jl

@@ -1,0 +1,9 @@
+using CSV
+using DataFrames
+
+abdata = CSV.read("data/real_data/mass_corrected_absorbances.csv", DataFrame)
+sidata = CSV.read("data/real_data/biomass_si_data.csv", DataFrame)
+
+merged = innerjoin(sidata, abdata, on=:uniqueID)
+
+CSV.write("data/real_data/si_absorbance_data.csv", merged)

### 2023-04-28 src/phenolicdataprep.jl

@@ -1,0 +1,81 @@
+using CSV
+using DataFrames
+using Statistics
+begin
+p1m = CSV.File("data/real_data/plate1_matrix.csv", header = false) |> Tables.matrix
+p2m = CSV.File("data/real_data/plate2_matrix.csv", header = false) |> Tables.matrix
+p3m = CSV.File("data/real_data/plate3_matrix.csv", header = false) |> Tables.matrix
+p4m = CSV.File("data/real_data/plate4_matrix.csv", header = false) |> Tables.matrix
+p5m = CSV.File("data/real_data/plate5_matrix.csv", header = false) |> Tables.matrix
+p6m = CSV.File("data/real_data/plate6_matrix.csv", header = false) |> Tables.matrix
+p7m = CSV.File("data/real_data/plate7_matrix.csv", header = false) |> Tables.matrix
+p8m = CSV.File("data/real_data/plate8_matrix.csv", header = false) |> Tables.matrix
+p9m = CSV.File("data/real_data/plate9_matrix.csv", header = false) |> Tables.matrix
+p10m = CSV.File("data/real_data/plate10_matrix.csv", header = false) |> Tables.matrix
+
+matlist = [p1m, p2m, p3m, p4m, p5m, p6m, p7m, p8m, p9m, p10m]
+
+p1k = CSV.File("data/real_data/plate1_key.csv", header = false) |> Tables.matrix
+p2k = CSV.File("data/real_data/plate2_key.csv", header = false) |> Tables.matrix 
+p3k = CSV.File("data/real_data/plate3_key.csv", header = false) |> Tables.matrix
+p4k = CSV.File("data/real_data/plate4_key.csv", header = false) |> Tables.matrix
+p5k = CSV.File("data/real_data/plate5_key.csv", header = false) |> Tables.matrix
+p6k = CSV.File("data/real_data/plate6_key.csv", header = false) |> Tables.matrix
+p7k = CSV.File("data/real_data/plate7_key.csv", header = false) |> Tables.matrix
+p8k = CSV.File("data/real_data/plate8_key.csv", header = false) |> Tables.matrix
+p9k = CSV.File("data/real_data/plate9_key.csv", header = false) |> Tables.matrix
+p10k = CSV.File("data/real_data/plate10_key.csv", header = false) |> Tables.matrix
+
+keylist = [p1k, p2k, p3k, p4k, p5k, p6k, p7k, p8k, p9k, p10k]
+end
+function builddf(ml, kl)
[…]

### 2023-04-29 manuscript/main.tex
added juli's comments

@@
 \usepackage{lineno}
+
 \usepackage{gensymb}
@@
 Juli Carrillo, with assistance from Dr.
-Jean-Thomas Cornelis, conceptualized and developed the experiment presented in Chapter Two.
+Jean-Thomas Cornelis and Dr.
+Gurcharn Singh Brar, conceptualized and developed the experiment presented in Chapter Two.
 Isaac Peetoom Heida, Dr.
@@
 Shaun Barker and the Mineral Deposit Research Unit of the University of British Columbia provided facilities and expertise for the XRF analysis of the tissue samples.
+Dr.
+Simone Castellarin and Dr.
+Joana Pico Carbajo provided facilities and guidance in developing the protocol for phenolic analysis.
+Chelsea Gowton provided indespensable assistance with phenolic sample processing.
 For Chapter Three, Isaac Peetoom Heida led plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda.
@@
 Daria Reshetniak, Paul Fisher, Lucas Friesen, Katie Pryer, Dr.
-Kinga Treder, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
-Chapters Two and Three of this thesis will be submitted to peer-reviewed journals for publication. For the purposes of this manuscript, actions are depicted in the singular first person.
+Kinga Treder, Chelsea Gowten, Dennis Chiu, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
+I use the first person plural for chapters two and three as they are intended for submission to a peer-reviewed journal.
 
@@
 
-As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, researchers are leaving no stone unturned in the search for new technologies for sustainable improvement in crop production.
+As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, new technologies are being developed for sustainable improvement in crop production.
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
@@
 
-Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass.
-Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant.
+Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass \cite{debona_silicons_2017}.
+Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant \cite{debona_silicons_2017}.
 Silicates are found in a variety of forms, and vary in their plant availability.
@@
 The morphology of phytoliths also varies between organs in the plant.
-Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory.
+Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory \cite{parr_phytolith_2011}.
 
@@
 We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
-To the resulting extract, we added 100$\mu$L of Fast Blue BB solution, allowed the color to develop in dark conditions for one hour, then used a 96-well plate to read the absorbance of [XXX] nm light.
+To the resulting extract, we added 100$\mu$L of Fast Blue BB solution, allowed the color to develop in dark conditions for one hour, then used a 96-well plate to read the absorbance of 420 nm light.
 We compared the readings to a standard curve created using gallic acid.

### 2023-04-29 manuscript/main.tex
continued addind juli's feedback

@@
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
-One of the promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance.
+One of the most promising developments to arise over the past 20 years is the potential of use silicon to improve crop plant performance.
 With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
-Our ability to mobilize silicon-based cropping strategies is dependent on a thorough understanding of the ultimate and proximate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
-In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common canadian crops, as well as using a  genome-wide association study to identify genetic markers associated with high silicon content.
+Our ability to mobilize silicon-based cropping strategies depends on a thorough understanding of the proximate and ultimate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
+In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common Canadian crops, and use a  genome-wide association study to identify genetic markers associated with high silicon content.
 
@@
 Combining these results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content.
-This development could provide an environmentally friendly strategy to maintain output to feed a growing human population.
+This development could provide an environmentally friendly strategy to reduce pesticide use while maintaining yields.
 
@@
 As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, new technologies are being developed for sustainable improvement in crop production.
-New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
+New crop technologies will meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
 Over the past 20 years, plant-silicon relations has emerged as a promising field that may safeguard crop performance and security within a changing biosphere.
@@
 
-Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass \cite{debona_silicons_2017}.
-Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant \cite{debona_silicons_2017}.
+Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass (\cite{holland_41_2014}).
+Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant (\cite{debona_silicons_2017}).
 Silicates are found in a variety of forms, and vary in their plant availability.
-Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available \cite{fraysse_surface_2009}.
-As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths \cite{de_tombeur_plants_2020}.
-Uptake and usage of silicon is not uniform throughout the vascular plants.
-Plant clades vary in their expression of silicon transporter proteins and in the relative abundance of phyotliths in their tissues.
+Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available (\cite{fraysse_surface_2009}).
+As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths (\cite{de_tombeur_plants_2020}).
+Uptake and usage of silicon is not uniform throughout the vascular plants and plant clades vary in their expression of silicon transporter proteins and in the relative abundance of phyotliths in their tissues (\cite{ma_chapter_2001}).
 Plant phytoliths are amorphous masses of silica, chemically similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}).
-Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations.
-The highly variable morphology among clades, with the conservation of morphology within clades suggests active selection upon the structure of phytoliths.
+Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations (\cite{piperno_phytoliths_2006}).
+The highly variable morphology among clades, with the conservation of morphology within clades hints at active selection upon the structure of phytoliths.
 The morphology of phytoliths also varies between organs in the plant.
-Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory \cite{parr_phytolith_2011}.
+Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory (\cite{parr_phytolith_2011}).
 
 Silicon deposition may create apoplastic barriers that seal and toughen the plant tissue, which may benefit the plant in a multitude of ways (\cite{coskun_controversies_2019}).
-Under this hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
+Under this apoplastic barrier hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
 Empirical evidence shows that silicon is effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}).

### 2023-04-29 manuscript/main.tex
more juli edits

@@
 The morphology of phytoliths also varies between organs in the plant.
-Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory (\cite{parr_phytolith_2011}).
+Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory (\cite{parr_phytolith_2011,stromberg_functions_2016}).
 
@@
 
-One of the most important advances in plant-silicon research was understanding the mechanisms through which silicon is acquired and transported into the plant.
+One of the most important advances in plant-silicon research was understanding the mechanisms through which plants aquire and transport silicon to specific location within the plant (\cite{coskun_controversies_2019}).
 Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}).
@@
 Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}).
-The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice \cite{yamaji_transporter_2008}.
-Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate \cite{christian_breeding_2022}.
+The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice (\cite{yamaji_transporter_2008}).
+Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate (\cite{christian_breeding_2022}).
 However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant.
@@
 
-Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
+Plants deposit silicon in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
 Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
@@
 This overall toughness, and abundance of phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}).
+Leaf silicon concentrations are negatively correlated with relative growth rate and survival in a number of invertebrate herbivores (\cite{juma_influence_2015,massey_silica_2006,mir_silicon_2019}).
+Phytoliths act by wearing down the mandibles of insect herbivores, thus reducing feeding efficieny (\cite{waterman_short-term_2021-1,mir_silicon_2019}), as well as reducing the digestive efficiency of the herbivore gut (\cite{hunt_novel_2008}).
 Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}).
@@
 
-Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by Takahashi et al.
+Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}), who published a comprehesive study of silica in biotic systems, and the subsequent charting of silicon content across the plant kingdom by Takahashi et al.
 (\citeyear{takahashi_possibility_1990}).
-Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world.
+Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world (\cite{coskun_controversies_2019,hartley_ecology_2016,cooke_is_2011,christian_breeding_2022,de_tombeur_silicon_2021}).
 Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}).
-Rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}), and are all domesticated grass species.
+The domesticated grass species rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}).
 Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}).
-Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}).
-Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate dry-land production systems, particularly in wheat and barley.
-This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils.
+The critical depletion of silicon in these production systems has spurred much research activity targeting the biology of silicon, particularly in rice (\cite{deren_variable_1992-1,dai_genetic_2005,rodrigues_silicon_2004,ye_priming_2013}).
+In most temperate soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}).
+Thus, silicon research has been slower to develop in dryland temperate cereal crops, (but see \cite{rains_active_2006,neu_silicon_2017,ahmad_silicon_2016}).
+Great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils.
 Our ability to integrate silicon as a tool for production improvement in crop production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which crops utilize silicon in pest-protection.
@@
 
-\chapter{Chapter 1: Identifying rapid silicon accumulation in cereal crops}
+\chapter{Chapter 1: Testing rapid silicon accumulation in four cereal crops under real and artificial herbivory}
 
@@
 To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
-In the broadest of terms, plant defenses can be categorized as either tolerance of resistance strategies.
-Resistance strategies involve attempts to limit tissue loss through interference with herbivory, while tolerance strategies result in increased regrowth after tissue loss.
 Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
-In most scenarios, induced plant defences are activated in response to an external cue, and build in intensity over time, with defensive hormone signals peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}).
-Despite this rapid hormonal response, actual defensive phenotypes are slower to emerge, often operating on the scale of days or generations (\cite{karban_induced_1989}).
+In most scenarios, plants induce defences in response to an external cue, and build in intensity over time, with defensive hormone signals peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}).
+Despite this rapid hormonal response, actual defensive phenotypes are slower to emerge, due to the higher costs of synthesizing defensive compounds, or developing new tissues like thorns and spines (\cite{karban_induced_1989}).
 Many defensive responses are also context dependent, where the identity of the damaging actor, the severity of damage, and a host of other factors interact to determine the final defensive response (\cite{waterman_simulated_2019}).
-The most effective defensive strategies should be those that can either prevent herbivory outright, or can mount a rapid response to limit damage.
+The most effective defensive strategies should be those that can either prevent herbivory outright, or rapidly respond to limit damage.
 These same strategies are also the most promising for crop production, where pest damage represents both an economic and food security cost.
@@
 Plant silicon is expressed latently, but also increases in response to herbivory (\cite{takahashi_possibility_1990, }).
-Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even preempting some chemical defences (\cite{waterman_short-term_2021}).
+Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even preempting the production of phenolic compounds, a group of defensive compounds (\cite{waterman_short-term_2021}).
 This rapid action makes silicon accumulation a promising trait for future crop development.
-Despite the novel results, this pattern has so far been observed in just one species, and only under artificial herbivory via the application of methyl-jasmonate.
+Despite the novel results, this pattern has so far been observed in just one species, and only under simulated herbivory through the application of methyl-jasmonate.
 Though a useful tool for herbivory research, methyl-jasmonate application fails to reproduce a complete herbivory signal for the plant, thus observed changes to plant defence may not be representative of a true herbivory scenario (\cite{strauss_direct_2002}).
@@
 \begin{enumerate}
-        \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species silicon content would show a significant increase in silicon content within 18 hours of the herbivory treatment applications.
+        \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species' silicon content would show a significant increase in silicon content within 18 hours of the herbivory treatment applications.
         \item Due to different phylogeny and domestication history, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
@@
 \end{enumerate}
-This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, but attempts to extend the findings to commercially important grain crops.
+This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, which found that silicon accumulated within six hours after methyl-jasmonate application, and was faster to accumualte than phenolic compounds.
+We attempt to extend these findings to commercially important grain crops.
 The findings of this study will refine our understanding of the prevalence of rapid silicification in the Poaceae, and will help to inform the value of potential applications of silicon-based defences into grain crops.
@@
 After four days, we transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
-Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be no silicon limitation to the plants.We randomized the location of each pot within the growing space.
+Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.We randomized the location of each pot within the growing space.
 A flood table bottom watered the pots with nutrient solution.
@@
 2021b), while crickets housed in water-pik tubes provided true herbivory.
-Prior to introduction to the the plants, we acclimated crickets by feeding them on the same species used in this trial.
+Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial.
 Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
-Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visual inspection of the leaves for missing tissue.
+Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.
 
@@
 
-18 hours after treatment application, we harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
-After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.
+After freeze drying, we ground the samples for phenolic processing under the same conditions as the other samples.
 
@@
 
-To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
+To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}).
 We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
 We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
-After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.
+After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples (comment left to include picture.
+Picture of the device or the pellet die?).
 
@@
 To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30Hz for 60 seconds.
-To the leaf powder, we added 1ml of 1\% formic acid in methanol, sonicated for 15 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
+To the leaf powder, we added 1ml of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
 We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
-To the resulting extract, we added 100$\mu$L of Fast Blue BB solution, allowed the color to develop in dark conditions for one hour, then used a 96-well plate to read the absorbance of 420 nm light.
+To the resulting extract, we added 100 $\mu$L of Fast Blue BB solution, vortexed for one minute, added 100 $\mu$L of 5\% NaOH, vortexed for 10 seconds, and allowed the color to develop in dark conditions for one hour.
+We used a 96-well plate to read the absorbance of 420 nm light.
 We compared the readings to a standard curve created using gallic acid.
@@
 
-Despite the starvation, some crickets did not initiate feeding during the exposure period.
-We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model.
-Prior to running our full model, we first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels (\cite{carmona_plant_2011}).
+Despite the starvation, some crickets did not initiate feeding during the exposure period (\sim38\%).
+We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model (n = 33).
+Prior to running our full model, we first tested for an effect of plant biomass of silicon content, as defense levels are influenced by plant size (\cite{carmona_plant_2011}).
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
@@
 We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$).
-Post-hoc tukey tests revealed that this was driven primarily by wheat's response to my induction treatments, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
-%Our model showed that both species and treatment type had effects on the plant silicon content.
-Parameter estimates and 90\% credible intervals are summarized in Table \ref{Tab:params}.
-Species and induction treatments had an interaction.
+Post-hoc tukey tests revealed that this was driven primarily by wheat, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
 
@@
 The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
-In my study, we grew plants in potting soil amended with solid silicic acid, similar to Nacimento et al.
-(\citedate{nascimento_silicon_2019}).
+In my study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}.
 In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
@@
 \section{Acknowledgements}
+
+We thank Dr.
+Aaron Beattie, Dr.
+Mazen Aljarrah, and Dr.
+Gurcharn Singh Brar for providing seeds for this experiment.
+Chelsea Gowton assisted with the harvesting and processing of samples.
+Dennis Chiu assisted with sample preparation and XRF analysis.
+Dr.
+Shaun Barker and [Dr.
+Brian McNulty] assisted with XRF analysis.
+Dr.
+Simone Castellarin and Dr.
+Joana Pico Carbajo advised and assisted with the development of the protocol for phenolic analysis.
 
@@
 With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production.
-Over the past thirty years, research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
+Over the past thirty years, increasing research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
 Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}).
-Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\>0.75\% total dry weight) (\cite{reynolds_silicon_2016}).
-Silicon is highly abundant in many soils globally, and is the second most abundant element in the earth’s crust, behind only oxygen (\cite{ma_functions_2003}).
-It’s high expression in cereals, high abundance in many soils, and incredible broad spectrum effects on plant vigor and stress tolerance have make it a tantalizing target for improvements in agricultural yield and sustainability.
-Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.
+Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\sim0.6 - 10\% total dry weight) (\cite{reynolds_silicon_2016}).
+Silicon's high abundance in many soils, high expression in cereal crops, and incredible broad spectrum effects on plant vigor and stress tolerance have made it a tantalizing target for improvements in agricultural yield and sustainability.
 
 Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress.
+Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.
 For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005, katz_silicon_2019}).

### 2023-04-29 manuscript/main.tex
more juli comments. Prepare for merge conflict

--- 2023-04-29 manuscript/main.tex
@@
 
-18 hours after treatment application, we harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis.
+We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
@@
 
-To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}).
+To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}).
 We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.

### 2023-04-30 manuscript/main.tex
more juli comments

@@
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
-After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.
+After freeze drying, we ground the samples for phenolic processing under the same conditions as the other samples.
 
@@
 We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
-After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.
+After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples (comment left to include picture.
+Picture of the device or the pellet die?).
 

### 2023-05-01 manuscript/main.tex
worked on the stats

--- 2023-04-29 manuscript/main.tex
@@
 We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model (n = 33).
-Prior to running our full model, we first tested for an effect of plant biomass of silicon content, as defense levels are influenced by plant size (\cite{carmona_plant_2011}).
+Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defense levels are influenced by plant size (\cite{carmona_plant_2011}).
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
-To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:
+To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following R model formula:
 
@@
 Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}).
-Amongst species, oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
+We found strong support for a species effect on silicon content ($p=0.012, F = 6.08, df=3,10.3$).
+Oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
 Counter to my predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species.

### 2023-05-01 src/R_linear_model.R

@@ -10,18 +10,38 @@ library(tidyverse)
 library(emmeans)
 library(car)
 library(Hmisc)
-si_data <- as.data.frame(read.csv("data/real_data/si_absorbance_data.csv"))
+library(lmerTest)
+si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
-si_mod <- lme4::lmer(Si_ppm/10000 ~ Induction * Species +mass_g + (1|Genotype), si_filtered)
+si_filtered$Genotype <- as.factor(si_filtered$Genotype)
+si_filtered$Induction <- as.factor(si_filtered$Induction)
+si_filtered$Species <- as.factor(si_filtered$Species)
+si_mod <- lmerTest::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
+summary(aov(Si_ppm/10000 ~ Induction * Species + mass_g, si_filtered))
 summary(si_mod)
-anova(si_mod)
-Anova(si_mod, type ="III", test.statistic = "F")
+lmerTest::anova(si_mod)
+Anova(si_mod, type =2, test.statistic = "F")
+si_mod0 <- lmer(log(Si_ppm) ~  mass_g + (1|Species/Genotype), si_filtered)
+si_mod1 <- lmer(log(Si_ppm) ~ Induction + mass_g + (1|Species/Genotype), si_filtered)
+si_mod2 <- lmer(log(Si_ppm) ~ Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod3 <- lmer(log(Si_ppm) ~ Induction + Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod4 <- lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod5 <- lmer(log(Si_ppm) ~ Induction:Species + mass_g + (1|Species/Genotype), si_filtered)
+fit.formula <- log(Si_ppm) ~ Induction+Species + mass_g
+X <- model.matrix(fit.formula, si_filtered)
+pb01 <- PBmodcomp(si_mod1, si_mod0, nsim = 10000, cl = 4)
+pb02 <- PBmodcomp(si_mod2, si_mod0, nsim = 10000, cl = 4)
[…]

### 2023-05-02 manuscript/main.tex
added more julis comments, different stats tried

@@
 \usepackage{titlesec}
-
+\newcommand{\mycomment}[1]{}
 % Change chapter heading font size to 14pt
@@
 At the start of the experiment, we germinated seeds in germination trays filled with moist sand.
-After four days, we transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid.
+After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid.
 Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.We randomized the location of each pot within the growing space.
@@
 We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2 mL microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
 We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
@@
 To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}).
-We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet.
+We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13 mm die to create a pellet.
 We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
@@
 To measure the response of phenolics to our treatments, we used the Fast Blue BB method (\cite{pico2020systematic}).
-To prepare our samples, we took 0.075G of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30Hz for 60 seconds.
-To the leaf powder, we added 1ml of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 rpm for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
+To prepare our samples, we took 0.075 g of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30 Hz for 60 seconds.
+To the leaf powder, we added 1 mL of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 G for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
 We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
@@
 
-Despite the starvation, some crickets did not initiate feeding during the exposure period (\sim38\%).
+Despite the starvation, some crickets did not initiate feeding during the exposure period ($\sim38\%$).
 We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model (n = 33).
@@
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
-To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following R model formula:
+To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon and phenolics using the following R model formulae:
 
-\[Si \sim Species * Induction + Biomass + (1|Genotype)\]
+\[Silicon \sim Species * Induction + Biomass + (1|Genotype)\]
+\[Phenolics \sim Species * Induction + Biomass + (1|Genotype)\]
+
+We tested for significance using model comparison in \verb|pbkrtest| with 10000 bootstrap iterations.
+
+Though not directly related to our hypotheses, we conducted additional post-hoc tests targetting trade-offs between phenolic and silicon content (\cite{simpson_still_2017}).
+We tested the relationship between silicon and phenolic content using a linear mixed effects model with the following R model formula:
+
+\[Phenolics \sim Silicon + Species + Biomass + (1|Genotype)\]
+
+[I am open to feedback about what this model should be.
+Should we include an induction effect?
+Drop the species effect?
+No matter how I specify it, there is not much support for a significant relationship between Si and Phenolic, despite what the scatter plot suggests.
+Perhaps we just decide to exclude?]
 %[Insert text about how great these models are]. I specified a hierarchical model using the following model:
@@
 We implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}).
-We tested the full mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
+We tested the mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}) and \verb|pbkrtest| (\cite{halekoh_pbkrtest_2014}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
 We generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).
@@
 Post-hoc tukey tests revealed that this was driven primarily by wheat, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
+We found a significant effect of both species and induction on phenolic content [need to finalize stats approach] (Figure \ref{Fig:absorbance_boxplots}).
+We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).
 
@@
 Dr.
-Shaun Barker and [Dr.
-Brian McNulty] assisted with XRF analysis.
+Shaun Barker and Dr.
+Brian McNulty advised and assisted with XRF analysis.
 Dr.
@@
 
+\mycomment{
 \begin{center}
@@
 \end{center}
+}
 
@@
         \centering
-        \caption{The effects of crop species and induction treatment on leaf silicon content.
+        \caption{Boxplots showing leaf silicon content for four cereal species across three defense induction treatments.
 Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
@@
         \label{Fig:induction}
+\end{figure}
+
+\begin{figure}[h]
+        \includegraphics[width = \textwidth]{images/absorbance_boxplots.png}
+        \centering
+        \caption{Boxplots showing leaf phenolic content for four cereal species across three defense induction treatments.
+Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
+Leaves were sampled 18 hours after treatment, and were analyzed using the Fast Blue BB method.}
+        \label{Fig:absorbance_boxplots}
+\end{figure}
+
+\begin{figure}[h]
+        \includegraphics[width = \textwidth]{images/phenolic_silicon_regression.png}
+        \centering
+        \caption{Scatter plot of the relationship between leaf silicon and leaf phenolic concentration}
+        \label{Fig:phe_si_scatter}
 \end{figure}
@@
 Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}).
-Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\sim0.6 - 10\% total dry weight) (\cite{reynolds_silicon_2016}).
+Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content ($\sim0.6 - 10\%$ total dry weight) (\cite{reynolds_silicon_2016}).
 Silicon's high abundance in many soils, high expression in cereal crops, and incredible broad spectrum effects on plant vigor and stress tolerance have made it a tantalizing target for improvements in agricultural yield and sustainability.
@@
 On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}).
-In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}).
-A current hypothesis explaining the broad-spectrum activity of silicon is presented in Coskun et al.
-(\citeyear{coskun_controversies_2019}), where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance [I could be more specific if needed].
+In comparing stressed plants grown in the absence or presence of silicon, silicon-exposed plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}).
+A current hypothesis explaining the broad-spectrum activity of silicon is presented in  \textcite{coskun_controversies_2019}.
+Their apoplastic barrier hypothesis suggests that silicon deposited in the apoplast of plant tissues modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance.
 Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts.
@@
 Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilization of silicon in decaying plant material (\cite{de_tombeur_silicon_2021-1}).
-Weathering of silicates releases a host of plant nutrients including Al, Si, Fe, and P (\cite{de_tombeur_silicon_2021-1}).
+Weathering of silicates releases a host of plant nutrients including aluminium, silicon, iron, and phophorus (\cite{de_tombeur_silicon_2021-1}).
 Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms.
-Plant roots can release carboxylates and phytosiderophores to weather P and Si out of soil minerals.
+Plant roots can release carboxylates and phytosiderophores to weather phosphorus and silicon out of soil minerals.
 Along with Si and P mobilization, Mn is often released, and taken up by plants roots.
@@
 One of the most exciting features of XRF is the fact that it can analyse multiple elements at once, allowing for broad characterization of the sample for most elements heavier than aluminum.
-Though XRF is an established technique to measure plant Si, its may also be used to measure other metals of interest, including manganese.
-In this study we use XRF to quantify variation in Si and Mn content among a diversity panel of a wild ancestor of bread wheat, Aegilops tauschii.
-This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link Si and Mn variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
+Though XRF is an established technique to measure plant silicon, its may also be used to measure other metals of interest, including manganese.
+In this study we use XRF to quantify variation in silicon and manganese content among a diversity panel of a wild ancestor of bread wheat, \textit{Aegilops tauschii}.
+This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link silicon and manganese variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
 
@@
 After 12 weeks, we moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks.
-We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with [amount] of silicic acid (Tixosil 68B, Solvay).
+We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with 0.17g of silicic acid (Tixosil 68B, Solvay).
 Pots were arranged using the same randomized block design, adapted to fit on two flood tables.
-To ensure a comparable life stage accross environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.
+To ensure a comparable life stage across environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.
 
@@
 For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022.
-We placed harvested material in labelled paper bags, and dried it in drying ovens at 60ºC for 48 hours.
+We placed harvested material in labelled paper bags, and dried it a 60ºC drying oven for 48 hours.
 To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers.
@@
 Some plants did not yield enough leaf tissue to meet to 200mg threshold.
-To reduce costs and increase the amount of biomass availabe per genotype, we pooled leaf material from within sites.
+We pooled leaf material from within sites.
 For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
@@
 
-$Manganese ~ Silicon + (1|Genotype)$
+$Manganese \sim Silicon + (1|Genotype)$
 
-To perfrom the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
+To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
 For brevity, this methodology only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.

### 2023-05-02 src/R_linear_model.R

@@ -5,6 +5,7 @@ install.packages("xtable")
 install.packages("car")
 install.packages("Hmisc")
 library(lme4)
+library(pbkrtest)
 library(tidyverse)
 #library(xtable)
 library(emmeans)
@@ -53,21 +54,14 @@ logit_mod <- glmer(Si_percent ~ Si_percent + Species + (1|Genotype), insect_data
 summary(logit_mod)
 anova(logit_mod)
 
-pigsint.lm <- lm(log(conc) ~ source * factor(percent), data = pigs)
-pigs.sav <- as.list(ref_grid(pigsint.lm))
-pigs.anew <- as.emmGrid(pigs.sav)
-pigsint.emm <- pairs(emmeans(pigsint.lm, ~ percent | source))
-test(pigsint.emm, by = NULL, adjust = "bonferroni")
-typeof(pigsint.emm)
-xtable::xtable(pigsint.emm, type = "response")
 
 
-si_mod0 <- lmer(log(Si_ppm) ~  mass_g + (1|Species/Genotype), si_filtered)
-si_mod1 <- lmer(log(Si_ppm) ~ Induction + mass_g + (1|Species/Genotype), si_filtered)
-si_mod2 <- lmer(log(Si_ppm) ~ Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod3 <- lmer(log(Si_ppm) ~ Induction + Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod4 <- lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod5 <- lmer(log(Si_ppm) ~ Induction:Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod0 <- lmer(log(Si_ppm) ~  mass_g + (1|Genotype), si_filtered)
+si_mod1 <- lmer(log(Si_ppm) ~ Induction + mass_g + (1|Genotype), si_filtered)
+si_mod2 <- lmer(log(Si_ppm) ~ Species + mass_g + (1|Genotype), si_filtered)
+si_mod3 <- lmer(log(Si_ppm) ~ Induction + Species + mass_g + (1|Genotype), si_filtered)
+si_mod4 <- lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Genotype), si_filtered)
[…]

### 2023-05-02 src/cereal_induction.jl

@@ -178,8 +178,8 @@ biomass_si_regression = lm(@formula(Si_ppm ~ mass_g * Species), analysis_data)
 gboxplot = @df analysis_data groupedboxplot(:Induction, :Si_ppm/10000, group = :Species)
 gboxplot = @df analysis_data groupeddotplot(:Induction, :Si_ppm/10000, group = :Species)
 plot!(analysis_data.Induction, analysis_data.Si_ppm/10000, group = analysis_data.Species, seriestype=:scatter)
-plot!(gboxplot, xlabel = "Induction Treatment", ylabel = "Silicon Content (%)")
-#png(gboxplot, "images/induction_plot")
+plot!(gboxplot, xlabel = "Induction Treatment", ylabel = "Silicon Content (%)", size = (800,600), dpi = 600)
+png(gboxplot, "manuscript/images/induction_plot")
 test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = repeat(["D", "E", "F"], inner = 9, outer = 1), yvals = rand(Normal(0,1), 27))
 tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
 
@@ -198,10 +198,13 @@ ad_spread = spreadvars(df=analysis_data, treat_types=[:Species,:Induction], inte
     plot(chains)
 
 phenolic_data = CSV.read("data/real_data/si_absorbance_data.csv", DataFrame)
-
-gboxplot_ab = @df phenolic_data groupedboxplot(:Induction, :mcabsorbance, group = :Species)
-
-ph_si_scatter = plot(phenolic_data.Si_ppm/10000, phenolic_data.mcabsorbance, group=phenolic_data.Species, seriestype=:scatter)
+phenolics_filtered = filter(row -> ismissing(row.isDamaged) || row.isDamaged  ≠ "Undamaged", phenolic_data)
+gboxplot_ab = @df phenolics_filtered groupedboxplot(:Induction, :mcabsorbance, group = :Species)
+plot!(xlabel = "Treatment Group", ylabel = "Phenolic Content", dpi = 600, size = (800,600))
+png(gboxplot_ab, "manuscript/images/absorbance_boxplots")
+phenolic_data.Si_pc = phenolic_data.Si_ppm./10000
+phenolic_data.treatmentcombo = phenolic_data.Induction .* phenolic_data.Species
[…]

### 2023-05-02 src/R_linear_model.R
no effect of Si on phenolics

@@ -80,15 +80,15 @@ summary(ab_mod)
 Anova(ab_mod)
 summary(lm(mcabsorbance ~ mass_g, absorbancedf))
 absorbance_mod <- lmer(scale(mcabsorbance) ~ scale(Si_ppm) + scale(mass_g) + Species+ Induction + (1|Genotype) ,REML = FALSE, abdf_filtered)
-
-
-
-absi_mod0 <- stan_lmer(log(mcabsorbance) ~  mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
-absi_mod1 <- stan_lmer(log(mcabsorbance) ~ Induction + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
+control_absorbance_only <- absorbancedf[absorbancedf$Induction == "Control",]
+summary(lmerTest::lmer(scale(log(mcabsorbance)) ~ scale(log(Si_ppm)) + Species + mass_g + (1|Genotype), control_absorbance_only))
+summary(lm(scale(mcabsorbance) ~ scale(Si_ppm), control_absorbance_only))
+absi_mod0 <- lme4::lmer(log(mcabsorbance) ~  mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
+absi_mod1 <- lme4::lmer(log(mcabsorbance) ~ Induction + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
 absi_mod2 <- lme4::lmer(log(mcabsorbance) ~ Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
 absi_mod3 <- lme4::lmer(log(mcabsorbance) ~ Induction + Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
-absi_mod4 <- stan_lmer(log(mcabsorbance) ~ Induction * Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
-absi_mod5 <- stan_lmer(log(mcabsorbance) ~ Induction:Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
+absi_mod4 <- lme4::lmer(log(mcabsorbance) ~ Induction * Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
+absi_mod5 <- lme4::lmer(log(mcabsorbance) ~ Induction:Species + mass_g + (1|Genotype), abdf_filtered, REML = FALSE)
[…]

### 2023-05-02 src/cereal_induction.jl

@@ -203,7 +203,7 @@ gboxplot_ab = @df phenolics_filtered groupedboxplot(:Induction, :mcabsorbance, g
 plot!(xlabel = "Treatment Group", ylabel = "Phenolic Content", dpi = 600, size = (800,600))
 png(gboxplot_ab, "manuscript/images/absorbance_boxplots")
 phenolic_data.Si_pc = phenolic_data.Si_ppm./10000
-phenolic_data.treatmentcombo = phenolic_data.Induction .* phenolic_data.Species
+phenolic_data.treatmentcombo = phenolic_data.Induction .* " " .* phenolic_data.Species
 ph_si_scatter = plot(phenolic_data.Si_pc, phenolic_data.mcabsorbance, group=phenolic_data.Species, seriestype=:scatter)
 plot!(ph_si_scatter, xlabel = "Leaf Silicon Content (%)", ylabel = "Leaf Phenolic Content")
 plot!(size = (800,600), dpi = 800)
@@ -236,12 +236,18 @@ plot!(meansphsiscatter_induction, xlabel = "Leaf Silicon Content (%)", ylabel = 
 plot!(size = (800,600), dpi = 800)
 #png(ph_si_scatter, "manuscript/images/phenolic_silicon_regression")
 
-gph_trt = groupby(phenolic_data, [:treatmentcombo])
-test_trt = combine(gph_trt, [:mcabsorbance, :Si_pc] => ((a, s) -> (sppMEANabsorbance = mean(a), sppSEabsorbance = standarderror(a),sppMEANsi = mean(s), sppSEsi = standarderror(s))) => AsTable)
+gph_trt = groupby(phenolics_filtered, [:treatmentcombo])
+test_trt = combine(gph_trt, [:mcabsorbance, :Si_pc, :Species, :Induction] => ((ab, si, sp, in) -> (sppMEANabsorbance = mean(ab), sppSEabsorbance = standarderror(ab),sppMEANsi = mean(si), sppSEsi = standarderror(si), spp = first(sp), ind = first(in))) => AsTable)
 meansphsiscatter_induction = plot(test_trt.sppMEANsi .± test_trt.sppSEsi, 
     test_trt.sppMEANabsorbance .± test_trt.sppSEabsorbance, 
     seriestype = "scatter", 
     groups = test_trt.treatmentcombo, 
     markersize = 15
 )
+
[…]

### 2023-05-02 manuscript/main.tex
clarified p val for mn ~ si reg in qtl ch

@@
 For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
-Overall, we were left with approximately 115 useable genotypes from each site.
-We packed dried leaves samples were 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
+Overall, we were left with 359 samples across three sites, making approximately 115 useable genotypes per site.
+We packed dried leaf samples in 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
 We stored the resulting leaf powder sealed until XRF analysis.
@@
 In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
-For beam 1 (Mn), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time.
+For beam 1 (manganese), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time.
 Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy.
@@
 
-To look for evidence of root exudation driving silicon content, we tested a correlation between observed leaf silicon and manganese content.
+To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content.
 Prior to the analysis, we plotted histograms of silicon and manganese concentrations.
@@
 To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
-For brevity, this methodology only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.
+For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
@@
 Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}).
+For manganese, we observed numerous peaks in each environment, yet only a few peaks were consistent across environment.
 My results for manganese content are less clear.
-I detected no regions that met the threshold for significance, though there were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}).
-Within the plants, silicon and manganese content were correlated ($R^2$ = 0.15, p = 0.049) (Figure 3).
+(Figure \ref{Fig:mn_peak_plot}).
+Within the plants, silicon and manganese content were correlated ($R^2$ = [How to get R2 from a lmer Model??], p = 0.0043) (Figure 3).
 

### 2023-05-08 manuscript/main.tex
corrected Si Mn reg stats

@@
 
-Need to revisit this and figure out exactly what to say.
-
+To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{rank_plots}).
 To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content.
@@
 Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}).
-For manganese, we observed numerous peaks in each environment, yet only a few peaks were consistent across environment.
-My results for manganese content are less clear.
-(Figure \ref{Fig:mn_peak_plot}).
-Within the plants, silicon and manganese content were correlated ($R^2$ = [How to get R2 from a lmer Model??], p = 0.0043) (Figure 3).
+For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance.
+However, when comparing accross environments, we found no regions that consistently exceeded the threshold significance.
+We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil.
+were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}).
+Within the plants, silicon and manganese content were correlated ($R^2_m = 0.05, R^2_c = 0.35 , p < 0.0001$) (Figure 3).
 
@@
 \section{Tables and Figures}
+
+\begin{figure}[h]
+        \includegraphics[scale=0.048]{images/rank_plots.png}
+        \centering
+        \caption{}
+        \label{Fig:mn_si_regression}
+\end{figure}
 

### 2023-05-08 manuscript/main.tex
added a few things to the results of qtl chapter

@@
 The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
-Silicon and manganese showed significant positive correlation ($\beta = 0.278\pm0.051, z = 5.46, p<0.0001$) (Figure \ref{Fig:mn_si_regression}).
-Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}).
+Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}).
+Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor for manganese content (Figure \ref{Fig:mn_peak_plot}).
 For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance.
@@
 We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil.
-were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}).
-Within the plants, silicon and manganese content were correlated ($R^2_m = 0.05, R^2_c = 0.35 , p < 0.0001$) (Figure 3).
-
 \section{Discussion}

### 2023-05-23 manuscript/main.tex
added a table of the genes matching the peaks

@@
 \documentclass[12pt, letterpaper, ]{report}
-\usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
+\usepackage[style=authoryear, maxnames=2, maxbibnames=99, citestyle=apa]{biblatex}
 \addbibresource{merged_biblio2.bib}
@@
 \usepackage{lineno}
-
+\usepackage{tabularx}
 \usepackage{gensymb}
@@
 At the start of the experiment, we germinated seeds in germination trays filled with moist sand.
-After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid.
-Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.We randomized the location of each pot within the growing space.
+After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil).
+Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.
+We randomized the location of each pot within the growing space.
 A flood table bottom watered the pots with nutrient solution.
@@
 The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
-In this study, we attempted to demonstrate rapid silicifcation in four cereal crops.
-We failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure.
-In our study, the variation among our cultivars within species was similar in magnitude to variation among species (Figure \ref{Fig:baseline_si}), possibly obscuring the effects of our treatments.
-This study also differed from previous studies demonstrating rapid silicon uptake in a number of ways.
-The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
+In this study, we attempted to demonstrate rapid silicification in four cereal crops.
+We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
+
+This study differed from previous studies demonstrating rapid silicon uptake in a number of ways.
+The two previous studies (\cite{waterman_short-term_2021-1,waterman_short-term_2021}) both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
 In my study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}.
@@
 The theoretical maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that.
-In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense.
+In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-available silicon) (\cite{de_tombeur_plants_2020}), so as to avoid soil conditions with low silicon presence.
 Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.
+
+Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}.
+In their study, they tested 8 species groups of wild and domesticated cereal crops, including wheat and barley.
+Our results for wheat and barley silicon and phenolic content more or less follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content.
+\textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species.
+Of course, both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we are still able to interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic status among studies and growing environments.
+In our study, oats had the highest phenolic concentration, and the lowest silicon content.
+The relative difference in the phenolic and silicon content of oats and wheat in our study [was similar] to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study.
+This suggests an important role for shorter term evolution in determining silicon and phenolic content.
 
@@
 Shortly after transplanting to the field sites, we applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets.
-Field transplantation took place on the 15th of October 2022 and the 16th of December 2022.
+Field transplantation took place on the 15th of October 2021 and the 16th of December 2021.
 For the glasshouse environment, we started seedlings in growth chambers in January 2022.
@@
 
-To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{rank_plots}).
+To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{Fig:rank_plots}).
 To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content.
@@
 We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.
+For peaks that were significant in at least two environments, we crossreferenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.
 
@@
 
-Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis.
-After pooling, we were left with 359 samples across three environments.
  Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%.
 There were notable differences in silicon content based on the growing environment.
-The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
+The glasshouse plants averaged $1.450\% \pm 0.032 (SE)$, while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
+Manganese content varied from 36.5 ppm to 1296.5 ppm, and averaged $225.1 \pm 25.45 ppm (SE)$ in the glasshouse environment, and $ 71.05 \pm 2.571 ppm (SE)$ and $ 224.0 \pm 11.01 (SE)$ for the two outdoor growing environments.
+The results summarized in our rank plots (Figure \ref{Fig:rank_plots}) show that rank is variable for both silicon and manganese.
 Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}).
-Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor for manganese content (Figure \ref{Fig:mn_peak_plot}).
+Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}).
 For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance.
 However, when comparing accross environments, we found no regions that consistently exceeded the threshold significance.
-We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil.
+We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil, though absent in the glasshouse environment (Table \ref{tab:peaks_and_genes}).
+
 \section{Discussion}
 
-We observed major differences in silicon content between our growth environments, and did not find conistent correlations between genetic variation and variation in silicon content.
-Despite this, we did find a significant positive correlation between silicon and manganese content, supporting root exudates increasing silicon uptake.
+We observed major differences in silicon and manganese content between our growth environments, and did not find consistent correlations between genetic variation and variation in silicon content.
+Our rank plots revealed that relative content of both silicon and manganese is not stable across environments.
+This suggests strong genotype by environment interactions, likely reduced our ability to detect genetic contribution to the silicon and manganese phenotype.
+Interestingly, we detected a number of peaks for leaf managese content, though many of these were only found in a single environment.
+Though we detected no peaks present in all three growing environments, we detected 11 regions that were significant in both outdoor environments.
+Compared to the glasshouse environment, the outdoor environments likely exposed the plants to a soil environment in which root exudation yeilds nutrients for uptake.
+Root exudation of organic acids and chelating agents liberates phosphates and silicates from inorganic pools (\cite{de_tombeur_silicon_2021-1,lambers_plant_2008}).
+In the glasshouse environment, we grew plants in standard potting mix, which is high in organic matter with a relatively low inorganic component.
+In this soil context, root exudation is unlikely to yeild appreciable nutrient liberation.
+The high manganese content found in the glasshouse plants is most likely due to the watering regime they experienced.
+We watered these plants using municipal tap water.
+In the Vancouver region, this water is sourced from reservoirs in the surrounding mountains, which themselves are filled from ground runoff.
+Nearby groundwater wells have issues with high manganese content (\cite{hu_drinking_2020}).
+A constant supply of soluble manganese in the irrigation the glasshouse plants likely accounts for the elevated manganese content.
 
+Our observations of a correlation between silicon and manganese content lend support to the idea of root-exudate based silicon aquisition (\cite{de_tombeur_silicon_2021-1}).
+It is possible that this positive correlation is due to a nutrient effect more generally
 \section{Acknowledgements}
@@
 
+\begin{table}[ht]
+        \centering
+        \caption{GWAS peaks for manganese content crossreferenced with the \textit{Aegilops tauschii} Aet v5.0 assembly. }
+        \label{tab:peaks_and_genes}
+        \begin{tabular}{|lrrr|}
+        \hline
+        \textbf{Chromosome} & \textbf{Peak Location}                & \textbf{Range}                             & \textbf{Gene Symbol} \\
+        \hline
+        1D         & 201950k-201960k   & No Gene Found                     &                  \\
+        1D         & 402540k-402550k   & 402549667-402554765              & LOC109755172   \\
+        1D         & 402540k-402550k   & 402547029-402549598 & LOC120968250      \\
+        1D         & 418870k-418880k   & No Gene Found                     &                  \\
+        1D         & 418990k-419000k   & No Gene Found                     &                  \\
+        1D         & 419030k-419040k   & No Gene Found                     &                  \\
+        3D         & 575420k-575430k   & 575424894-575428979 & LOC109751197   \\
+        4D         & 85300k-85310k     & No Gene Found                     &                  \\
+        4D         & 85310k-85320k     & No Gene Found                     &                  \\
+        4D         & 85330k-85340k     & No Gene Found                     &                  \\
+        \hline
+        \end{tabular}
+\end{table}
+
 \begin{figure}[h]
@@
         \centering
-        \caption{}
-        \label{Fig:mn_si_regression}
+        \caption{Average rank of silicon and manganese content of 113 genotypes of \textit{Aegilops tauschii} grown in three different environments.
+Error bars represent the standard deviation of rank.}
+        \label{Fig:rank_plots}
 \end{figure}

### 2023-05-24 manuscript/main.tex
im making so much progress!

@@
 \usepackage{titlesec}
+\usepackage{subcaption}
 \newcommand{\mycomment}[1]{}
@@
 Our observations of a correlation between silicon and manganese content lend support to the idea of root-exudate based silicon aquisition (\cite{de_tombeur_silicon_2021-1}).
-It is possible that this positive correlation is due to a nutrient effect more generally
+It is possible that this positive correlation is due to a nutrient effect more generally, however further regressions show that silicon only shows positive correlations with iron and manganese, both of which are mobilized by root exudation (Supplementary Fig [Number]) (\cite{de_tombeur_silicon_2021-1}).
+Unfortunately, the two genes we did detect as associated with manganese content are both non-coding RNA, reducing the likelihood that they are involved in root exudation or manganese uptake.
+Silicon interacts with manganese in the leaf, binding manganese to the cell wall and reducing manganese concentrations in the cell symplast (\cite{rogalla_role_2002}).
+Root exudation is most important under limited nutrients, and silicon-exudate dynamics have been extensively studied in well-weathered soils (\cite{lambers_plant_2008,de_tombeur_shift_2021,de_tombeur_silicon_2021-1}).
+In agricultural soils, it is unlikely that plants would be exposed to such nutrient poor conditions, and so the specific dynamics of root exudation and silicon mobilization deserve further investigation.
+Root exudation also results in increased soil carbon content, and there is interest in implementing cropping and fertilization systems with an explicit focus on improving soil carbon through root exudation (\cite{cornelis_soil_2022}).
+In this context, understanding how increased root exudation may impact both silicon and metal uptake in plants is crucial, and future studies should pursue similar genetic approaches under limited phosphorus regimes to try and detect genes associated with root exudation activity.
+
+Depsite our inability to detect QTLs that influence silicon or manganese content, this field of study still holds immense promise.
+We implemented a study design that we hoped would allow generalizations about the genetic regions we hoped to find.
+However, the use of multiple environments may have obscured QTLs that we may have otherwise detected in a single environment.
+The use of the diversity panel allowed for a rapid and low-cost first search for loci asscociated with silicon and manganese content, but may have also limited our power to detect loci.
+The panel we used was a subset of the larger \textit{Aegilops tauschii} panel, and ours was selected specifically for diversity in disease-resistance genes.
+Utilizing the full panel of 384 genotypes could improve power of future studies.
+Alternatively, future researchers could attempt to create a mapping population.
+By using the silicon and manganese content rankings generated by this study, one could create a cross of a low and high genotype and use a traditional qtl mapping approach with the recombinant population.
+This approach might allow for more sensitivity in the analysis by reducing the overall genetic variation between genotypes in population, highlighting the genetically driven variation in silicon content.
 \section{Acknowledgements}
@@
         \centering
-        \caption{GWAS peaks for manganese content crossreferenced with the \textit{Aegilops tauschii} Aet v5.0 assembly. }
+        \caption{Genomic regions with significant correlation to leaf manganese content, and the corresponding genes in the \textit{Aegilops tauschii} Aet v5.0 assembly.
+The assembly was scanned using the NCBI genome viewer tool.}
         \label{tab:peaks_and_genes}
-        \begin{tabular}{|lrrr|}
+        \begin{tabular}{lrrr}
         \hline
@@
 \begin{figure}[h]
-        \includegraphics{images/Manhattan_Plot.png}
         \centering
-        \caption{This is an example Manhattan Plot from the GWAS output. The real figure will show associations with silicon content}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/tf_si_plot.png}
+                \label{Fig:tf_si_peak}
+        \end{subfigure}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/sc_si_plot.png}
+                \label{Fig:sc_si_peak}
+        \end{subfigure}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/gh_si_plot.png}
+                \label{Fig:gh_si_peak}
+        \end{subfigure}
+        \caption{Manhattan plots of the results of our GWAS analysis for leaf silicon content in \textit{Aegilops tauschii}.
+Red dots indicate a negative correlation, while blue dots denote a positive correlation.
+The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.}
         \label{Fig:si_peak_plot}
@@
 \begin{figure}[h]
-        \includegraphics{images/Manhattan_Plot.png}
         \centering
-        \caption{This is another Manhattan Plot, this time showing associations with manganese content}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/tf_mn_plot.png}
+                \label{Fig:tf_mn_peak}
+        \end{subfigure}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/sc_mn_plot.png}
+                \label{Fig:sc_mn_peak}
+        \end{subfigure}
+        \begin{subfigure}[b]{\textwidth}
+                \centering
+                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/gh_mn_plot.png}
+                \label{Fig:gh_mn_peak}
+        \end{subfigure}
+        \caption{Manhattan plots of the results of our GWAS analysis for leaf manganese content in \textit{Aegilops tauschii}.
+Red dots indicate a negative correlation, while blue dots denote a positive correlation.
+The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.}
         \label{Fig:mn_peak_plot}

### 2023-05-26 manuscript/main.tex
added the experimental schematic

@@
 After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil).
-Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.
+Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to reduce silicon limitation to the plants.
 We randomized the location of each pot within the growing space.
@@
 Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial.
+We applied our treatments 45 days after transplanting.
 Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
@@
 Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
-The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
+The promising results of this work have been highlighted for their potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
 In this study, we attempted to demonstrate rapid silicification in four cereal crops.
-We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
+In our study, we failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
 
@@
 }
+
+\begin{figure}[ht]
+        \includegraphics[width = \textwidth]{images/Induction_schematic.png}
+        \centering
+        \caption{Experimental schematic for this study.
+We grew four species of cereal crops (left box) in a glasshouse.
+After 45 days, we applied one of three treatments to each plant, in a fully randomized manner (centre box).
+To analyse the response of leaf silicon and phenolic content, we used X-ray fluorescence (XRF) and the Fast Blue BB phenolic assay (right box).}
+\end{figure}
 

### 2023-05-29 manuscript/main.tex
added the pedigree table

@@
 \usepackage{subcaption}
+\usepackage{pdflscape}
+\usepackage{booktabs}
+\usepackage{array}
 \newcommand{\mycomment}[1]{}
@@
 \end{figure}
+\clearpage
+
+\section{Supplementary Info}
+\begin{landscape}
+\begin{table}
+        \centering
+        \caption{Pedigrees of the cultivars used in this study}
+        \begin{tabular}{ll>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{4cm}}
+          \toprule
+          Species & Cultivar & Parent 1 & Parent 2 & Parent 1 Pedigree & Parent 2 Pedigree \\
+          \midrule
+          Barley & CDC Austenson & TR358 & 94Ab12271 & Not Available & Not Available \\
+          Barley & AAC Synergy & TR02267 & Newdale & TR253/AC Metcalfe & CDC Stratus//TR236/WM862-6 \\
+          Barley & CDC Copeland & WM861-5 & TR118 & Harrington & \\
+          Oats & AC Morgan & OT526 & OT763 & 17578/Alfred & Fidler/Cascade \\
+          Oats & Camden & SW Betania & Dominik & SW 151-7-95/SW 52-95 & Not Available \\
+          Oats & AC Summit & Ronald & OT299 & W89329/AC Medallion & AC Rebel/Dumont 48 \\
+          Triticale & Bunker & 95P105 & 85L0120006 & Pika-5/Yogui-1 & Spring triticale*n/RL4137 \\
+          Triticale & Tyndal & Nimir-1/Hare-265//Erizo-9 & 88L012 & Nimir-1/Hare-265//Erizo-9 & Pronghorn*n/RL4137 \\
+          Triticale & AB Stampeder & BULL 10/MANATI1//FARA S/CMH84.4414 & POLLMER 2.2.1 $\ast$2//FARA S/CMH84.4414 & BULL 10/MANATI1//FARA S/CMH84.4414 & POLLMER 2.2.1 $\ast$2//FARA S/CMH84.4414 \\
+          Wheat & AAC Brandon & Superb/CDC Osler & ND744 & Superb/CDC Osler & ND2831/Parshall//ND706 \\
+          Wheat & CDC Landmark & Unity & BW864 & McKenzie$\ast$3// BW174$\ast$2/Clark & YAQUI 50-ENANO/3$\ast$KALYANSONA \\
+          Wheat & CDC Superb & BW282 & CDC Go & Not Available & Grandin/SD3055 \\
+          \bottomrule
+        \end{tabular}
+      \end{table}
+
+      \end{landscape}
 
@@
 Red dots indicate a negative correlation, while blue dots denote a positive correlation.
-The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.}
+The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.
+The dotted line represents the bonferroni-corrected significance threshold.}
         \label{Fig:si_peak_plot}
@@
 Red dots indicate a negative correlation, while blue dots denote a positive correlation.
-The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.}
+The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.
+The dotted line represents the bonferroni-corrected significance threshold.}
         \label{Fig:mn_peak_plot}

### 2023-05-31 src/R_linear_model.R
added more models

@@ -4,6 +4,7 @@ install.packages("emmeans")
 install.packages("xtable")
 install.packages("car")
 install.packages("Hmisc")
+install.packages("rstanarm")
 library(lme4)
 library(pbkrtest)
 library(tidyverse)
@@ -12,27 +13,32 @@ library(emmeans)
 library(car)
 library(Hmisc)
 library(lmerTest)
+library(rstanarm)
+options(contrasts = c("contr.helmert", "contr.poly"))
 si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
 si_filtered$Genotype <- as.factor(si_filtered$Genotype)
 si_filtered$Induction <- as.factor(si_filtered$Induction)
 si_filtered$Species <- as.factor(si_filtered$Species)
 si_mod <- lmerTest::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod <- lme4::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Genotype), si_filtered)
+anova(si_mod, type = 2)
 summary(aov(Si_ppm/10000 ~ Induction * Species + mass_g, si_filtered))
 summary(si_mod)
 lmerTest::anova(si_mod)
 Anova(si_mod, type =2, test.statistic = "F")
-si_mod0 <- lmer(log(Si_ppm) ~  mass_g + (1|Species/Genotype), si_filtered)
-si_mod1 <- lmer(log(Si_ppm) ~ Induction + mass_g + (1|Species/Genotype), si_filtered)
-si_mod2 <- lmer(log(Si_ppm) ~ Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod3 <- lmer(log(Si_ppm) ~ Induction + Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod4 <- lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
-si_mod5 <- lmer(log(Si_ppm) ~ Induction:Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod0 <- lmer(log(Si_ppm) ~  mass_g + (1|Genotype), si_filtered)
[…]

### 2023-06-04 manuscript/main.tex
ahhhhhh

--- 2023-05-26 manuscript/main.tex
@@
 
-As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, there is growing research interest in new avenues for production improvement.
+As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, new avenues for sustainable production must be sought for food security.
 New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
-One of the most promising developments to arise over the past 20 years is the potential of use silicon to improve crop plant performance.
+One of the most promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance.
 With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
-Our ability to mobilize silicon-based cropping strategies depends on a thorough understanding of the proximate and ultimate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
+Our ability to mobilize silicon-based crop improvement strategies depends on a thorough understanding of the proximate and ultimate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
 In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common Canadian crops, and use a  genome-wide association study to identify genetic markers associated with high silicon content.
@@
 
-Silicon provides tremendous benefits to plant health, but is not widely utilized in agriculture, despite it being highly abundance and naturally occurring in most soils.
-One of the main factors limiting it's application in agriculture is our poor understanding of the exact dynamics of how plants absorb and use silicon from the soil.
+Silicon provides tremendous benefits to plant health, but is not yet widely utilized in agriculture despite it being highly abundance and naturally occurring in most soils.
+One of the main factors limiting its application in agriculture is our poor understanding of the exact dynamics of how plants absorb and use silicon from the soil.
 I identified genetic traits associated with high silicon content in \textit{Aegilops tauschii}, a relative of bread wheat, as well as demonstrated that cereal crops (e.g.
 wheat, barley, oats) have the ability to rapidly uptake silicon from the soil.
-This rapid uptake means that silicon may be a highly effective defence against insect pests.
+This rapid uptake means that silicon may be a highly effective defence against insect pests, as it can reduce damage caused by chewing insects.
 Combining these results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content.
@@
 Joana Pico Carbajo provided facilities and guidance in developing the protocol for phenolic analysis.
-Chelsea Gowton provided indespensable assistance with phenolic sample processing.
+Chelsea Gowton provided indispensable assistance with phenolic sample processing.
 For Chapter Three, Isaac Peetoom Heida led plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda.
@@
 Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass (\cite{holland_41_2014}).
-Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant (\cite{debona_silicons_2017}).
-Silicates are found in a variety of forms, and vary in their plant availability.
+Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as the plant absorbs water (\cite{debona_silicons_2017}).
+Silicates occur in a variety of forms, and vary in their plant availability.
 Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available (\cite{fraysse_surface_2009}).
-As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths (\cite{de_tombeur_plants_2020}).
-Uptake and usage of silicon is not uniform throughout the vascular plants and plant clades vary in their expression of silicon transporter proteins and in the relative abundance of phyotliths in their tissues (\cite{ma_chapter_2001}).
-Plant phytoliths are amorphous masses of silica, chemically similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}).
-Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations (\cite{piperno_phytoliths_2006}).
-The highly variable morphology among clades, with the conservation of morphology within clades hints at active selection upon the structure of phytoliths.
-The morphology of phytoliths also varies between organs in the plant.
+As soils age, biogenic silicates, such as diatom testes and plant phytholiths, comprise an increasing share of the plant-available silicon (\cite{de_tombeur_plants_2020}).
+Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems.
+Within the soil environment, silicon commonly exists in both crystalline (geologic) and amorphous (biogenic) forms (\cite{haynes_contemporary_2014}).
+Amorphous silicates can derive from previous plant material that has decayed in the soil, but also from marine and aquatic organisms such as diatoms.
+Globally, the silicon cycle involves silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments.
+The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil (\cite{de_tombeur_plants_2020}).
+In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients, including silicon, for uptake.
+Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
+This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation may improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).
+
+Throughout the plant kingdom, clades vary significantly in their silicon content and expression of silicon transporter proteins (\cite{ma_chapter_2001}).
+Plant silicon is often expressed through phytoliths, amorphous masses of silica, chemically similar to opal, that occur throughout the plant body and have highly variable geometries both across clades and between parts of the plant body (\cite{piperno_phytoliths_2006,parr_phytolith_2011}).
+Despite the highly variable geometries overall, phytolith morphology is often consistent within taxonomic clades, hinting at selective pressure for specific morphologies.
 Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory (\cite{parr_phytolith_2011,stromberg_functions_2016}).
-
-Silicon deposition may create apoplastic barriers that seal and toughen the plant tissue, which may benefit the plant in a multitude of ways (\cite{coskun_controversies_2019}).
-Under this apoplastic barrier hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
-Empirical evidence shows that silicon is effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}).
-The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
-Indeed, when supplemented with silicon plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
-Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
 
@@
 
-\section{Silicon in leaves}
+\section{Silicon in plants}
 
@@
 Thus, silicon-based defences in crop plants may be one of the first lines of active anti-herbivore defence, providing rapid and sensitive responses to herbivory.
+
+In addition to the anti-herbivore benefits, silicon deposition seems to provide heightened disease resistance and stress tolerance.
+when supplemented with silicon, plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
+Silicon may create apoplastic barriers that seal and toughen the plant tissue, which might explain the range of plant traits that show a positive response to silicon (\cite{coskun_controversies_2019}).
+Under this apoplastic barrier hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
+Empirical evidence shows that silicon is effective at limiting the growth of plant fungal pathogens (\cite{fauteux_silicon_2005}).
+Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
 
@@
 
-To address acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
+To counter acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
 Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
@@
 Plant silicon is expressed latently, but also increases in response to herbivory (\cite{takahashi_possibility_1990, }).
-Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even preempting the production of phenolic compounds, a group of defensive compounds (\cite{waterman_short-term_2021}).
+Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even outpacing the production of phenolic compounds, a group of defensive compounds (\cite{waterman_short-term_2021}).
 This rapid action makes silicon accumulation a promising trait for future crop development.
@@
         \item Rapid silicon accumulation is a conserved trait in the Poaceae, and the tested species' silicon content would show a significant increase in silicon content within 18 hours of the herbivory treatment applications.
-        \item Due to different phylogeny and domestication history, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
+        \item Due to disparate taxonomic origins, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
         \item Due to the different cues involved when comparing true herbivory damage and methyl-jasmonate induced defensive induction, the tested species would show different patterns of short-term silicon accumulation in response to cricket (\textit{Acheta domesticus}) herbivory and methyl-jasmonate application.
 \end{enumerate}
-This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, which found that silicon accumulated within six hours after methyl-jasmonate application, and was faster to accumualte than phenolic compounds.
+This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, which found that silicon accumulated within six hours after methyl-jasmonate application, and was faster to accumulate than phenolic compounds.
 We attempt to extend these findings to commercially important grain crops.
@@
 
-To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley.
+To test the prevalence of rapid silicon accumulation in Canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley (\ref{Fig:exp_design}).
 We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
 At the start of the experiment, we germinated seeds in germination trays filled with moist sand.
-After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil).
+After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon / kg of soil).
 Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to reduce silicon limitation to the plants.
@@
 A flood table bottom watered the pots with nutrient solution.
-We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory.
+We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or cricket herbivory.
 We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al.
-2021b), while crickets housed in water-pik tubes provided true herbivory.
-Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial.
+2021b), while crickets housed in water-pik tubes provided true herbivory (\ref{Fig:exp_design}).
+Prior to introduction to the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial.
 We applied our treatments 45 days after transplanting.
-Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
+Immediately preceding cricket application, we placed the crickets in their experimental tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.
 Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.
@@
 
-We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application.
+We harvested three fully expanded upper leaves from the plants 18 hours after treatment application, and split the leaves in half along the midvein.
 We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2 mL microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
@@
 We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
-After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples (comment left to include picture.
-Picture of the device or the pellet die?).
+After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.
 
@@
 Despite the starvation, some crickets did not initiate feeding during the exposure period ($\sim38\%$).
-We filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model (n = 33).
-Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defense levels are influenced by plant size (\cite{carmona_plant_2011}).
+We filtered out plants assigned to the insect induction treatment that received no damage, to ensure that they would not confound the model (n = 33).
+Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defense levels can be influenced by plant size (\cite{carmona_plant_2011}).
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
-To test all three of our hypotheses, we used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon and phenolics using the following R model formulae:
+To test all three of our hypotheses, we used linear mixed effects models.
+We tested the responses of leaf silicon and phenolic content to the fixed effects of treatment and species, with a covariate of biomass.
+We used cultivar identity as a random effect.
+We implement these models using the following R model formulae:
 
@@
 To analyse the response of leaf silicon and phenolic content, we used X-ray fluorescence (XRF) and the Fast Blue BB phenolic assay (right box).}
+        \label{Fig:exp_design}
 \end{figure}
@@
         \centering
-        \caption{Baseline (uninduced) silicon content in the cereal cultivars used in this study. Cultivar species is notated in parentheses in the legend.}
+        \caption{Baseline (uninduced) silicon content (\% dry matter)in the cereal cultivars used in this study. Cultivar species is indicated by point color. Error bars are 1 Standard Error.}
         \label{Fig:baseline_si}
@@
 \end{figure}
+\clearpage
+
+\section{Supplementary Info}
+\begin{landscape}
+\begin{table}
+        \centering
+        \caption{Pedigrees of the cultivars used in this study}
+        \begin{tabular}{ll>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{4cm}}
+          \toprule
+          Species & Cultivar & Parent 1 & Parent 2 & Parent 1 Pedigree & Parent 2 Pedigree \\
+          \midrule
+          Barley & CDC Austenson & TR358 & 94Ab12271 & Not Available & Not Available \\
+          Barley & AAC Synergy & TR02267 & Newdale & TR253/AC Metcalfe & CDC Stratus//TR236/WM862-6 \\
+          Barley & CDC Copeland & WM861-5 & TR118 & Harrington & \\
+          Oats & AC Morgan & OT526 & OT763 & 17578/Alfred & Fidler/Cascade \\
+          Oats & Camden & SW Betania & Dominik & SW 151-7-95/SW 52-95 & Not Available \\
+          Oats & AC Summit & Ronald & OT299 & W89329/AC Medallion & AC Rebel/Dumont 48 \\
+          Triticale & Bunker & 95P105 & 85L0120006 & Pika-5/Yogui-1 & Spring triticale*n/RL4137 \\
+          Triticale & Tyndal & Nimir-1/Hare-265//Erizo-9 & 88L012 & Nimir-1/Hare-265//Erizo-9 & Pronghorn*n/RL4137 \\
+          Triticale & AB Stampeder & BULL 10/MANATI1//FARA S/CMH84.4414 & POLLMER 2.2.1 $\ast$2//FARA S/CMH84.4414 & BULL 10/MANATI1//FARA S/CMH84.4414 & POLLMER 2.2.1 $\ast$2//FARA S/CMH84.4414 \\
+          Wheat & AAC Brandon & Superb/CDC Osler & ND744 & Superb/CDC Osler & ND2831/Parshall//ND706 \\
+          Wheat & CDC Landmark & Unity & BW864 & McKenzie$\ast$3// BW174$\ast$2/Clark & YAQUI 50-ENANO/3$\ast$KALYANSONA \\
+          Wheat & CDC Plentiful & BW282 & CDC Go & AC Elsa/AC Barrie & Grandin/SD3055 \\
+          \bottomrule
+        \end{tabular}
+      \end{table}
+
+      \end{landscape}
 

### 2023-06-04 src/cereal_induction.jl

@@ -6,6 +6,10 @@ using MCMCChains
 using GLM
 using MixedModels
 using TuringGLM
+using DataFrames
+using Measurements
+using Statistics
+using ColorSchemes
 
 ### Combine 2019-2021 insured acreages of grain crops
 ### Data sourced from the Canadian Grain Commission https://www.grainscanada.gc.ca/en/grain-research/statistics/varieties-by-acreage/
@@ -130,19 +134,114 @@ analysis_data = CSV.read("data/real_data/biomass_si_data.csv", DataFrame)
 # Some summary stats on the data
 
 #############
+my_names = unique(analysis_data.Genotype)
+official_names = [
+    "CS Camden", 
+    "AC Summit", 
+    "AC Morgan", 
+    "CDC Austenson",
+    "AAC Brandon",
+    "CDC Landmark",
+    "AB Stampeder",
+    "Pronghorn",
+    "CDC Plentiful",
+    "CDC Copeland",
+    "AAC Synergy",
+    "Tyndal" ]
+
+cultivar_dict = Dict(my_names .=> official_names)
+
+transform!(analysis_data, :Genotype => ByRow(x->cultivar_dict[x]) => :OfficialName)
 analysis_data_noinduction = filter(row -> row.Induction == "Control", analysis_data)
 analysis_data_noinduction.Si_ppm = analysis_data_noinduction.Si_ppm./10000
 gdat_geno = groupby(analysis_data_noinduction, :Genotype)
 summ_stats_geno = combine(gdat_geno, :Si_ppm => mean, :Si_ppm => std, nrow)
 summ_stats_geno.Si_ppm_se = summ_stats_geno.Si_ppm_std ./ sqrt.(summ_stats_geno.nrow)
-CSV.write("./data/real_data/summary_statistics_genotype.csv", summ_stats_geno)
+#CSV.write("./data/real_data/summary_statistics_genotype.csv", summ_stats_geno)
 
+gsi_spp = groupby(analysis_data_noinduction, [:OfficialName])
+spp_si_contents_noind = combine(gsi_spp, [:Si_ppm, :Species] => ((si, sp) -> (sppMEANsi = mean(si), sppSEsi = standarderror(si), spp = first(sp))) => AsTable)
+spp_si_contents_noind.GenotypeID = 1:12
+sort!(spp_si_contents_noind, [:spp, :sppMEANsi])
+
[…]

### 2023-06-04 src/plotting.jl

@@ -1,11 +1,0 @@
-using StatsPlots;
-
-using DataFrames, IndexedTables;
-df = DataFrame(a = 1:10, b = 10 .* rand(10), c = 10 .* rand(10))
-@df df plot(:a, [:b :c], colour = [:red :blue])
-@df df scatter(:a, :b, markersize = 4 .* log.(:c .+ 0.1))
-t = table(1:10, rand(10), names = [:a, :b]) # IndexedTable
-@df t scatter(2 .* :b)
-
-
-

### 2023-06-09 manuscript/main.tex
added letters to induction plots

@@
 \usepackage{array}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\fancyfoot[R]{\thepage}
+\fancypagestyle{plain}{%
+    \renewcommand{\headrulewidth}{0pt}%
+    \fancyhf{}%
+    \fancyfoot[R]{\thepage}%
+}
+
 \newcommand{\mycomment}[1]{}
@@
 \begin{document}
-
+\pagenumbering{roman}
 \input{title.tex}
@@
 \tableofcontents
+\newpage
+\listoftables
+\newpage
+\listoffigures
+\chapter*{Abstract}
+\addcontentsline{toc}{chapter}{Abstract}
 
-\chapter{Abstract}
-
-As global agricultural production strains under degrading soil fertility and increasing losses due to climate change, new avenues for sustainable production must be sought for food security.
-New crop technologies must meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
-One of the most promising developments to arise over the past 20 years is the potential of silicon to improve crop plant performance.
+One of the most promising developments to arise over the past 30 years is the potential of silicon to improve crop plant performance.
 With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
@@
 In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common Canadian crops, and use a  genome-wide association study to identify genetic markers associated with high silicon content.
+We failed to observe rapid silicification in the tested crops.
+This may could be a result of a soil environment incapable of supplying sufficient silicon resources for rapid accumulation, or could reflect the effect of herbivore identity on specific defensive outcomes.
+In our genetic analysis, we failed to find genetic markers with significant associations to silicon content, but did find a marker that had a significant correlation with manganese content.
+Future work in this field might develop a mapping population using crosses from high and low silicon genotypes in this study, yielding experimental designs with increased power and thus may find QTLs associated with silicon content.
+Applications silicon research into crop production techniques are still limited by crucial gaps in our understanding of how ecology and genetics control silicon uptake, but this work provides yet another stepping stone towards the broader adoption of silicon into agricultural systems.
 
-[to add: results, conclusion]
-
-\chapter{Lay Summary}
+\chapter*{Lay Summary}
+\addcontentsline{toc}{chapter}{Lay Summary}
 
 Silicon provides tremendous benefits to plant health, but is not yet widely utilized in agriculture despite it being highly abundance and naturally occurring in most soils.
-One of the main factors limiting its application in agriculture is our poor understanding of the exact dynamics of how plants absorb and use silicon from the soil.
-I identified genetic traits associated with high silicon content in \textit{Aegilops tauschii}, a relative of bread wheat, as well as demonstrated that cereal crops (e.g.
-wheat, barley, oats) have the ability to rapidly uptake silicon from the soil.
-This rapid uptake means that silicon may be a highly effective defence against insect pests, as it can reduce damage caused by chewing insects.
-Combining these results with the genetic data, future research can aim towards creating breeding programs to develop cereal crops that can withstand insect damage based on their silicon content.
-This development could provide an environmentally friendly strategy to reduce pesticide use while maintaining yields.
+Silicon acts a defensive compound against insects, and also helps plants deal with stressful growing conditions, but is currently not widely applied in agriculture.
+Using cereal crops and a wild ancestor of wheat, I attempted to identify ecological and genetic factors that drive silicon content.
+We failed to find genetic regions that exerted control over silicon content, but observed large variations in silicon content between the growing sites.
+We did not observe rapid silicon uptake in our tested cereal crops, complicating our understanding of when rapid uptake is achieved by plant species.
+Continuing to build upon this research will help move silicon crop technologies closer to feasibility, and could provide an environmentally friendly strategy to reduce pesticide use while maintaining yields.
 
-\chapter{Preface}
-
-(This was taken nearly verbatim from Matt's thesis, so need to go back through to make sure I am not plagarizing)
+\chapter*{Preface}
+\addcontentsline{toc}{chapter}{Preface}
 
@@
 Chelsea Gowton provided indispensable assistance with phenolic sample processing.
-For Chapter Three, Isaac Peetoom Heida led plot set up and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, and Paul Fisher, and Matt Tsuruda.
+For Chapter Three, Isaac Peetoom Heida led planting, plot setup, and maintenance, with assistance from Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, Paul Fisher, and Matt Tsuruda.
 Isaac Peetoom Heida led the sample harvest, processing and analysis.
@@
 
+\chapter*{Acknowledgements}
+\addcontentsline{toc}{chapter}{Acknowledgements}
+
+This work would not have happened without the continuous support of my committee, the members of the PIEE lab, my friends and family, and the generosity of a host of funding agencies.
+Dr.
+Juli Carrillo has guided me to become a more well-rounded scientist and person, and has always believed in my abilities.
+I thank Dr.
+Jean-Thomas Cornelis for providing the silicon spark to kick this entire project off.
+Dr.
+Gurcharn Singh Brar lent his time and lab to me to support the work, and stoked my passion for grasses from a new angle.
+The members of the PIEE lab were more than generous with their time, and their continued interest in my project and general convivatilty provided the encouragement I needed to keep going with my research.
+In some ways this work is the culmination of a lifetime of support from my family, as they encouraged my curiosity about the natural world and sought out opportunities for me when I was too young (or too dumb) to seek them out for myself.
+Lastly, NSERC, CPBI, and the UBC provided generous scholarships and fellowships in support of my research and personal development.
+
+\chapter*{Decication}
+\addcontentsline{toc}{chapter}{Dedication}
+
+To my Auntie Margaret, for showing me where the wetland was all those years ago.
+
 \chapter{Introduction}
+\pagenumbering{arabic}
 
@@
 New crop technologies will meet increasing public and regulatory demand for environmental sustainability, encouraging scientists to revisit overlooked or relatively unknown techniques that may unlock productivity gains.
-Over the past 20 years, plant-silicon relations has emerged as a promising field that may safeguard crop performance and security within a changing biosphere.
+Over the past 30 years, plant-silicon relations has emerged as a promising field that may safeguard crop performance and security within a changing biosphere.
 With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
 
-\section{Silicon in nature}
+\section{Silicon in Soils}
 
@@
 Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available (\cite{fraysse_surface_2009}).
-As soils age, biogenic silicates, such as diatom testes and plant phytholiths, comprise an increasing share of the plant-available silicon (\cite{de_tombeur_plants_2020}).
+As soils age, biogenic amorphous silicates, such as diatom testes and plant phytholiths, comprise an increasing share of the plant-available silicon (\cite{de_tombeur_plants_2020}).
 Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems.
@@
 The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil (\cite{de_tombeur_plants_2020}).
+
+\section{Plant uptake of silicon}
+
+One of the most important advances in plant-silicon research was understanding the mechanisms through which plants acquire and transport silicon to specific location within the plant (\cite{coskun_controversies_2019}).
+Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}).
 In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients, including silicon, for uptake.
 Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
-This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation may improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).
-
-Throughout the plant kingdom, clades vary significantly in their silicon content and expression of silicon transporter proteins (\cite{ma_chapter_2001}).
-Plant silicon is often expressed through phytoliths, amorphous masses of silica, chemically similar to opal, that occur throughout the plant body and have highly variable geometries both across clades and between parts of the plant body (\cite{piperno_phytoliths_2006,parr_phytolith_2011}).
-Despite the highly variable geometries overall, phytolith morphology is often consistent within taxonomic clades, hinting at selective pressure for specific morphologies.
-Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory (\cite{parr_phytolith_2011,stromberg_functions_2016}).
-
-%Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation.
-When supplemented with silicon, plants are generally more resistant to stress.
-Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
-Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}).
-Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}).
-The recently put-forward apoplastic barrier hypothesis suggests that the benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition on the apoplast (\cite{coskun_controversies_2019}).
-Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.
-The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}).
-Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
-
-\section{Silicon in soils and roots}
-
-Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems.
-Within the soil environment, silicon commonly exists in both crystalline (geologic) and amorphous (biogenic) forms (\cite{haynes_contemporary_2014}).
-Amorphous silicates can derive from previous plant material that has decayed in the soil, but also from marine and aquatic organisms such as diatoms.
-Globally, the silicon cycle involves silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments.
-The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil(\cite{de_tombeur_plants_2020}).
-In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients, including silicon, for uptake.
-Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
-This active scavenging for silicon remains poorly understood, but it may be an important mechanism in plant defence (allowing for increased uptake during a defensive response) and breeding for increased root exudation may improve crop plant performance and nutrient use efficiency (\cite{de_tombeur_silicon_2021}).
-
-\section{Silicon transporters}
-
-One of the most important advances in plant-silicon research was understanding the mechanisms through which plants aquire and transport silicon to specific location within the plant (\cite{coskun_controversies_2019}).
-Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}).
+Once dissolved in soil solution, silicon can be taken up by active or passive transport by the root system.
 While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families.
-Research in rice has identified four gene products that transport silicon into and through the plant body.
+Research in rice has identified four gene products that actively transport silicon into and through the plant body.
 Two of these (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}).
@@
 Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}).
+Throughout the plant kingdom, clades vary significantly in their silicon content and expression of silicon transporter proteins (\cite{ma_chapter_2001}).
 The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice (\cite{yamaji_transporter_2008}).
-Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate (\cite{christian_breeding_2022}).
+Breeding for silicon content, improved exudate-based silicon mobilisation and silicon use-efficiency in crop plants may be crucial to improving crop performance under a changing climate (\cite{christian_breeding_2022,de_tombeur_silicon_2021}).
 However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant.
@@
 
-\section{Silicon in plants}
+\section{Silicon in the plant body}
 
 Plants deposit silicon in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
-Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
+Silicon deposits taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
 In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}).
@@
 Leaf silicon concentrations are negatively correlated with relative growth rate and survival in a number of invertebrate herbivores (\cite{juma_influence_2015,massey_silica_2006,mir_silicon_2019}).
-Phytoliths act by wearing down the mandibles of insect herbivores, thus reducing feeding efficieny (\cite{waterman_short-term_2021-1,mir_silicon_2019}), as well as reducing the digestive efficiency of the herbivore gut (\cite{hunt_novel_2008}).
+Phytoliths act by wearing down the mandibles of insect herbivores, thus reducing feeding efficiency (\cite{waterman_short-term_2021-1,mir_silicon_2019}), as well as reducing the digestive efficiency of the herbivore gut (\cite{hunt_novel_2008}).
 Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}).
-As phytoliths develop in the leaves, polymerization of monosilicic acid is aided by interactions with proteins in the cell wall, which act as sites of nucleation (\cite{nawaz_phytolith_2019}).
+As phytoliths develop in the leaves, polymerisation of monosilicic acid is aided by interactions with proteins in the cell wall, which act as sites of nucleation (\cite{nawaz_phytolith_2019}).
 Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (\cite{waterman_short-term_2021}).
@@
 
-Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}), who published a comprehesive study of silica in biotic systems, and the subsequent charting of silicon content across the plant kingdom by Takahashi et al.
-(\citeyear{takahashi_possibility_1990}).
+Much of today’s plant silicon work is indebted to the pioneering work of \textcite{jones_silica_1967}, who published a comprehensive study of silica in biotic systems, and the subsequent charting of silicon content across the plant kingdom by \textcite{takahashi_possibility_1990}).
 Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world (\cite{coskun_controversies_2019,hartley_ecology_2016,cooke_is_2011,christian_breeding_2022,de_tombeur_silicon_2021}).
@@
 
+Note: this chapter has been prepared with the intention of submission to a peer-reviewed journal, and thus I use the plural first-person throughout.
+
 \section{Introduction}
 
-To counter acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan down to the development of novel compounds to poison those that would try to eat the plant (\cite{agrawal_plant_2006}).
+To counter acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan to the development of novel compounds to poison organisms that try to eat the plant (\cite{agrawal_plant_2006}).
 Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
@@
 We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2 mL microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
-We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
+We placed the other half of the tissue into a microcentrifuge tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
 After freeze drying, we ground the samples for phenolic processing under the same conditions as the other samples.
@@
 We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13 mm die to create a pellet.
-We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
-After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.
+We then placed the pellet in an Olympus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
+After each use, we cleaned the pellet die and XRF analyser to minimize contamination between samples.
 
@@
 
-We analysed phenolics as a way to validate the effectiveness of our methyl-jasmonate applications on defense induction.
+We analysed phenolics as a way to validate the effectiveness of our methyl-jasmonate applications on defence induction.
 To measure the response of phenolics to our treatments, we used the Fast Blue BB method (\cite{pico2020systematic}).
 To prepare our samples, we took 0.075 g of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30 Hz for 60 seconds.
-To the leaf powder, we added 1 mL of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 G for 10 minutes, then pippetted out the supernatant into a new 2 mL  microcentrifuge tube.
+To the leaf powder, we added 1 mL of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 G for 10 minutes, then pipetted out the supernatant into a new 2 mL  microcentrifuge tube.
 We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
@@
 We filtered out plants assigned to the insect induction treatment that received no damage, to ensure that they would not confound the model (n = 33).
-Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defense levels can be influenced by plant size (\cite{carmona_plant_2011}).
+Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defence levels can be influenced by plant size (\cite{carmona_plant_2011}).
 We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model.
@@
 We tested the responses of leaf silicon and phenolic content to the fixed effects of treatment and species, with a covariate of biomass.
-We used cultivar identity as a random effect.
-We implement these models using the following R model formulae:
+We used cultivar identity as a random effect, as we were interested in understanding how silicon and phenolics varied across treatments and species, but not across cultivars.
+We implemented these models using the following R model formulae:
 
@@
 
-We tested for significance using model comparison in \verb|pbkrtest| with 10000 bootstrap iterations.
+We assessed the models using \verb|lmerTest| (\cite{kuznetsova_2017_lmerTest}) with a type III ANOVA.
 
-Though not directly related to our hypotheses, we conducted additional post-hoc tests targetting trade-offs between phenolic and silicon content (\cite{simpson_still_2017}).
+Though not directly related to our hypotheses, we conducted additional post-hoc tests investigating trade-offs between phenolic and silicon content (\cite{simpson_still_2017,waterman_short-term_2021}).
 We tested the relationship between silicon and phenolic content using a linear mixed effects model with the following R model formula:
@@
 \[Phenolics \sim Silicon + Species + Biomass + (1|Genotype)\]
-
-[I am open to feedback about what this model should be.
-Should we include an induction effect?
-Drop the species effect?
-No matter how I specify it, there is not much support for a significant relationship between Si and Phenolic, despite what the scatter plot suggests.
-Perhaps we just decide to exclude?]
-%[Insert text about how great these models are]. I specified a hierarchical model using the following model:
-
-%\[y_i \sim Normal(\hat{y}_i, \sigma)\]
-%\[\hat{y_i} = \alpha + \alpha_{[i]cultivar} + \beta_{[j]induction} + \beta_{[k]species} + \beta_{[jk]species \times induction}\]
-
-%I ran the model using \verb|Turing.jl| in Julia [version number] [cite].
-Using 4 chains and 1000 sample iterations, I sampled the posterior distribution using a No U-turn Sampler with 1000 warm up iterations and a target acceptance rate of 0.65 (\cite{hoffman_no-u-turn_2014}).
-I tested our model structure on simulated data, to ensure it returned accurate parameter estimates.
-I used the Gelman-Rubin statistic (\( \hat{R} \)) (\cite{gelman_inference_1992}) and effective sample size to diagnose the convergence of our chains.
-I verified model fits using posterior predictive checks in \verb|ArviZ.jl| [cite]
 
@@
 Oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
-Counter to my predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species.
-Despite a small p-value in the ANOVA table of the model ($p = 0.046, F = 3.14, df = 2,185.9$), the model showed only moderate support for an effect of methyl-jasmonate application ($\beta = 0.079 \pm 0.043, p = 0.070$), and no support for an effect of cricket exposure on silicon content ($\beta = 0.038 \pm 0.050, p = 0.453$).
-We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$).
-Post-hoc tukey tests revealed that this was driven primarily by wheat, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}).
-We found a significant effect of both species and induction on phenolic content [need to finalize stats approach] (Figure \ref{Fig:absorbance_boxplots}).
+Counter to our predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species.
+Despite a small p-value for an induction treatment effect ($p = 0.011, F = 4.62, df =2, 186$), the only significant pairwise differences between treatments were in wheat, where silicon content was lowest in the insect treatment group (Figure \ref{Fig:induction}).
+We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,186$), likely driven by the stark differences between wheat and the other species' responses to the induction treatments (Figure \ref{Fig:induction}).
+We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction}, Table \ref{Tab:phe_params}).
+Oats had the highest average leaf phenolic content [$val \pm err$], while barley had the lowest average phenolic content [$val \pm err$].
 We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).
@@
 
-Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
+Recent research in inducible silicon plant defences has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
 The promising results of this work have been highlighted for their potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
 In this study, we attempted to demonstrate rapid silicification in four cereal crops.
-In our study, we failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
+We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
 
@@
 The two previous studies (\cite{waterman_short-term_2021-1,waterman_short-term_2021}) both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
-In my study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}.
+In our study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}.
 In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
-The theoretical maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that.
+The maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that.
 In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-available silicon) (\cite{de_tombeur_plants_2020}), so as to avoid soil conditions with low silicon presence.
 Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.
+Furthermore, the observed silicon content of wheat and barley is lower than what is typically found for these species when grown in natural soils (\cite{simpson_still_2017,parr_phytolith_2011}.
+The substrate used in this study is ubiquitous in plant science, particularly for plants grown in pots.
+The lack of silicon availability in this substrate, even with silicon supplementation is notable, and deserves serious thought.
+Silicon plays a role in a myriad of plant physiological responses (\cite{coskun_controversies_2019}), often mediating/reducing the negative effects of stressors.
+Silicon limitation is assumed to be infrequent in global soils (\cite{guntzer_benefits_2012}), yet our data suggest that plants grown in silicon fail to reach typical levels of silicon content.
+We can only speculate how silicon deficiencies may skew observations in potted plant experiments, but we think it would be prudent for researchers to consider adding soluble silicon to their nutrient mixes to emulate realistic silicon availabilities.
+Nearly all cereal production worldwide takes place in soils and we need further work on the temporal aspects of silicon uptake in a range of soil conditions to broaden our understanding of rapid silicon uptake.
+Though rapid silicification retains promise for improved crop protection, our work highlights the need for future research and replication to fully characterize the nature of silicon induction.
 
+We did not see a clear relationship between silicon and phenolic content in our plants, matching the observations of \textcite{waterman_short-term_2021}).
+Though they observed a negative relationship between silicon and phenolics in plants grown without supplemental silicon, this relationship disappeared when they added silicon to the growth media.
+As previously discussed, our study seems to have imparted moderated silicon limitation on the plants, and this may explain the hint of a negative trend observable in the scatterplot (Figure \ref{Fig:phe_si_scatter}).
 Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}.
@@
 \textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species.
-Of course, both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we are still able to interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic status among studies and growing environments.
+Both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we can interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic content among studies and growing environments.
 In our study, oats had the highest phenolic concentration, and the lowest silicon content.
 The relative difference in the phenolic and silicon content of oats and wheat in our study [was similar] to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study.
-This suggests an important role for shorter term evolution in determining silicon and phenolic content.
+Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (eg.
+domestication) and ecological (eg.
+growing environment) drivers that also determine the final defensive phenotype of cereal species.
+Understanding how these various factors constrain the range of defensive expression for these species will help bring focus to efforts to translate silicon ecology into the field of agriculture.
+
+This study was the first that we are aware of to use crickets as a treatment to induce silicon uptake.
+Other studies have used crickets to measure herbivore performance in response to other defensive phenotype manipulations (eg.
+eCO$_2$, caterpillar herbivory) (\cite{biru_contrasting_2022, ryalls_impacts_2017}).
+Plant defensive responses to herbivory are linked to the identity of the herbivore (\cite{afkhami_endophyte-mediated_2009, amsberry_effects_2006}), and it could be that house crickets elicit different silicon responses than more commonly used caterpillar herbivores.
+Other studies have observed silicon increases in response to locust (Orthoptera) damage (\cite{massey_herbivore_2007, garbuzov_interactive_2011}), so it is unlikely that crickets do not induce silicon uptake, but it may be that the temporal dynamics of such a response differ from caterpillar-triggered defensive induction.
+approximately one third of the applied crickets do not initiate feeding within the time frame of the study.
+In \textcite{massey_herbivore_2007}, crickets tended to avoid feeding on plants with higher silicon content.
+This same pattern may have been at play in our own study system, where plants that had relatively high silicon content escaped damage, while the lower silicon plants experienced herbivory.
+Our questions and statistical approach required undamaged plants assigned to the insect treatment to be excluded.
+This exclusion may in part explain our observations of the low silicon content in the wheat-insect treatment group.
+Unfortunately, our study lacks the replication (and thus statistical power) to confirm if there were significant differences between the silicon content in the damaged and undamaged plants within the insect treatments.
+A future study that quantified the effect of pre-herbivore-exposure silicon content on the likelihood or extent of herbivory would help to answer this question.
+
+Though we did not find support for rapid silicon uptake in response to herbivory, we have added to our knowledge of silicon content in cereal species, and have identified crucial gaps in our understanding of the dynamics of silicon uptake.
+It is unclear how silicon uptake rates are modified by the nature of the growth substrate, and there are almost certainly important questions to be answered in comparing nutrient-solution-based and soil-based silicon uptake.
+Our use of potting mix seemed to limit the silicon availability for our plants.
+Imposing silicon limitation on study plants through the use of potting mix may skew results for a wide variety of plant studies, as silicon is involved in all manner of stress and damage tolerance responses.
+More explicit consideration of how silicon limitation may influence the results of a given study, and integrating silicon into fertilizer mixes for plant science may help increase the translatability of potted plant experiments to real soil environments.
+The future of silicon-based crop technologies remains bright, but there are still important obstacles that must be overcome as we try to translate silicon ecology research into production methods.
 
@@
         \centering
-        \caption{ANOVA table for our linear mixed effects model analyzing the effect of defense induction and species identity on leaf silicon content.
-We generated the ANOVA table using the R package car, specifying a type III ANOVA.}
-        \label{Tab:params}
-        \begin{tabular}{lrrrr}
+        \caption{ANOVA table for our linear mixed effects model analysing the effect of defence induction and species identity on leaf silicon content.
+We generated the ANOVA table using the R package lmerTest, specifying a type III ANOVA.
+Degrees of freedom are Satterthwaite approximations.}
+        \label{Tab:si_params}
+        \begin{tabular}{lrrr}
                \hline
-                \textbf{Effect} & \textbf{F} & \textbf{Df} & \textbf{Df residual} & \textbf{\textit{P} value} \\
+                \textbf{Effect} & \textbf{Degrees of Freedom} & \textbf{F value} &    \textbf{P-value} \\ 
                 \hline   
-                (Intercept) & 32.4990 & 1 & 24.350 & 6.775e-06*** \\
-                Induction & 1.5671 & 2 & 186.230 & 0.21139 \\   
-                Species & 6.0830 & 3 & 10.308 & 0.01199* \\ 
-                Biomass & 16.5099 & 1 & 190.074 & 7.076e-05*** \\
-                Induction*Species & 2.2514 & 6 & 186.352 & 0.04027* \\
+                Induction         &  2, 186.028 &  4.6235 &   0.01097 \\  
+                Species           &  3, 7.908 &  6.5119 &   0.01568 \\  
+                Biomass            &  1, 190.778 & 16.3709 & $<0.0001$ \\
+                Induction $\times$ Species  &    6, 186.051 &  2.2494 &   0.04045 \\  
                 \hline
@@
 
-\mycomment{
-\begin{center}
-\begin{longtable}{lrrrrr}
-        \caption{emmeans results of pairwise comparisons between groups.
-p-values are tested using the  adjustment, against the  multivariate normal distribution, a less conservative approach than typical bonferroni corrections.}
-        \label{Tab:emm_params}
-            \\ \hline
-            \textbf{Contrast} & \textbf{Estimate} & \textbf{SE} & \textbf{df} & \textbf{\textit{t} ratio} & \textbf{\textit{p} value} 
-            \\ \hline
-            \endfirsthead
-            \multicolumn{3}{c}{{\bfseries \tablename\ \thetable{} -- continued from previous page}}
-            \\ \hline
-            \textbf{Contrast} & \textbf{Estimate} & \textbf{SE} & \textbf{df} & \textbf{\textit{t} ratio} & \textbf{\textit{p} value} 
-            \\ \hline
-            \endhead
+\begin{table}[ht]
+        \centering
+        \caption{ANOVA table for our linear mixed effects model analyzing the effect of defence induction and species identity on leaf phenolic content.
+We generated the ANOVA table using the R package lmerTest, specifying a type III ANOVA.
+Degrees of freedom are Satterthwaite approximations.}
+        \label{Tab:phe_params}
+        \begin{tabular}{lrrr}
                \hline
-            \endfoot
+                \textbf{Effect} & \textbf{Degrees of Freedom} & \textbf{F value} &    \textbf{P-value} \\ 
                 \hline   
+                Induction         &     2, 180.677 & 0.6926 & 0.501608 \\
+                Species           &     3,   9.577 & 9.2329 & 0.003518 \\
+                Biomass          &     1, 151.907 & 0.0760 & 0.783174 \\
+                Induction $times$ Species &     6, 180.798 & 2.7535 & 0.013850 \\
                 \hline
-            \endlastfoot
-
-            Control Barley - Insect Barley & -373.9 & 514 & 186.37 & -0.727 & 0.9997 \\
-            Control Barley - MeJA Barley & -793.4 & 449 & 186.05 & -1.769 & 0.7656 \\
-            Control Barley - Control Oats & 723.0 &1126 & 10.08 & 0.642 & 0.9997 \\
-            Control Barley - Insect Oats & 678.6 &1169 & 11.68 & 0.581 & 0.9999 \\
-            Control Barley - MeJA Oats & 215.6 &1127 & 10.12 & 0.191 & 1.0000 \\
-            Control Barley - Control Triticale &-2321.0 &1155 & 11.08 & -2.009 & 0.6139 \\
-            Control Barley - Insect Triticale &-2516.2 &1185 & 12.19 & -2.124 & 0.5467 \\
-            Control Barley - MeJA Triticale &-2813.2 &1151 & 10.94 & -2.445 & 0.3819 \\
-            Control Barley - Control Wheat &-3485.7 &1136 & 10.43 & -3.068 & 0.1735 \\
-            Control Barley - Insect Wheat &-1638.1 &1149 & 10.92 & -1.425 & 0.9054 \\
-            Control Barley - MeJA Wheat &-3044.7 &1130 & 10.20 & -2.695 & 0.2854 \\
-            Insect Barley - MeJA Barley & -419.5 & 508 & 186.31 & -0.825 & 0.9992 \\
-            Insect Barley - Control Oats & 1096.9 &1148 & 10.86 & 0.956 & 0.9928 \\
-            Insect Barley - Insect Oats & 1052.5 &1190 & 12.55 & 0.884 & 0.9963 \\
-            Insect Barley - MeJA Oats & 589.5 &1149 & 10.91 & 0.513 & 1.0000 \\
-            Insect Barley - Control Triticale &-1947.0 &1175 & 11.85 & -1.657 & 0.8083 \\
-            Insect Barley - Insect Triticale &-2142.3 &1202 & 12.94 & -1.782 & 0.7439 \\
-            Insect Barley - MeJA Triticale &-2439.2 &1169 & 11.68 & -2.086 & 0.5702 \\
-            Insect Barley - Control Wheat &-3111.8 &1156 & 11.16 & -2.693 & 0.2775 \\
-            Insect Barley - Insect Wheat &-1264.1 &1169 & 11.68 & -1.081 & 0.9824 \\
-            Insect Barley - MeJA Wheat &-2670.8 &1150 & 10.97 & -2.322 & 0.4421 \\
-            MeJA Barley - Control Oats & 1516.4 &1120 & 9.87 & 1.354 & 0.9262 \\
-            MeJA Barley - Insect Oats & 1472.0 &1164 & 11.49 & 1.265 & 0.9514 \\
-            MeJA Barley - MeJA Oats & 1009.0 &1122 & 9.92 & 0.900 & 0.9951 \\
-            MeJA Barley - Control Triticale &-1527.6 &1148 & 10.83 & -1.330 & 0.9350 \\
-            MeJA Barley - Insect Triticale &-1722.8 &1177 & 11.88 & -1.464 & 0.8932 \\
-            MeJA Barley - MeJA Triticale &-2019.8 &1143 & 10.66 & -1.767 & 0.7490 \\
-            MeJA Barley - Control Wheat &-2692.4 &1129 & 10.17 & -2.385 & 0.4163 \\
-            MeJA Barley - Insect Wheat & -844.7 &1142 & 10.66 & -0.739 & 0.9990 \\
-            MeJA Barley - MeJA Wheat &-2251.3 &1123 & 9.97 & -2.004 & 0.6186 \\
-            Control Oats - Insect Oats & -44.4 & 552 & 186.14 & -0.080 & 1.0000 \\
-            Control Oats - MeJA Oats & -507.4 & 448 & 186.03 & -1.133 & 0.9870 \\
-            Control Oats - Control Triticale &-3043.9 &1136 & 10.40 & -2.679 & 0.2909 \\
-            Control Oats - Insect Triticale &-3239.2 &1158 & 11.17 & -2.798 & 0.2414 \\
-            Control Oats - MeJA Triticale &-3536.1 &1126 & 10.08 & -3.140 & 0.1563 \\
-            Control Oats - Control Wheat &-4208.7 &1114 & 9.65 & -3.779 & 0.0707 \\
-            Control Oats - Insect Wheat &-2361.1 &1129 & 10.16 & -2.092 & 0.5698 \\
-            Control Oats - MeJA Wheat &-3767.7 &1112 & 9.60 & -3.387 & 0.1198 \\
-            Insect Oats - MeJA Oats & -463.0 & 557 & 186.23 & -0.831 & 0.9991 \\
-            Insect Oats - Control Triticale &-2999.5 &1185 & 12.26 & -2.532 & 0.3359 \\
-            Insect Oats - Insect Triticale &-3194.8 &1207 & 13.17 & -2.646 & 0.2826 \\
-            Insect Oats - MeJA Triticale &-3491.7 &1176 & 11.97 & -2.968 & 0.1818 \\
-            Insect Oats - Control Wheat &-4164.3 &1164 & 11.49 & -3.578 & 0.0764  \\
-            Insect Oats - Insect Wheat &-2316.6 &1178 & 12.03 & -1.967 & 0.6373 \\
-            Insect Oats - MeJA Wheat &-3723.3 &1161 & 11.39 & -3.206 & 0.1339 \\
-            MeJA Oats - Control Triticale &-2536.5 &1139 & 10.49 & -2.228 & 0.4960 \\
-            MeJA Oats - Insect Triticale &-2731.8 &1161 & 11.28 & -2.354 & 0.4264 \\
-            MeJA Oats - MeJA Triticale &-3028.7 &1129 & 10.17 & -2.683 & 0.2894 \\
-            MeJA Oats - Control Wheat &-3701.3 &1116 & 9.74 & -3.315 & 0.1322 \\
-            MeJA Oats - Insect Wheat &-1853.6 &1131 & 10.26 & -1.639 & 0.8139 \\
-            MeJA Oats - MeJA Wheat &-3260.3 &1115 & 9.69 & -2.924 & 0.2210 \\
-            Control Triticale - Insect Triticale & -195.3 & 559 & 186.25 & -0.350 & 1.0000 \\
-            Control Triticale - MeJA Triticale & -492.2 & 512 & 186.60 & -0.961 & 0.9967 \\
-            Control Triticale - Control Wheat &-1164.8 &1136 & 10.39 & -1.026 & 0.9872 \\
-            Control Triticale - Insect Wheat & 682.9 &1151 & 10.93 & 0.593 & 0.9999 \\
-            Control Triticale - MeJA Wheat & -723.8 &1136 & 10.38 & -0.637 & 0.9997 \\
-            Insect Triticale - MeJA Triticale & -296.9 & 548 & 187.32 & -0.542 & 1.0000 \\
-            Insect Triticale - Control Wheat & -969.5 &1154 & 11.05 & -0.840 & 0.9974 \\
-            Insect Triticale - Insect Wheat & 878.1 &1170 & 11.62 & 0.751 & 0.9990 \\
-            Insect Triticale - MeJA Wheat & -528.5 &1156 & 11.11 & -0.457 & 1.0000 \\
-            MeJA Triticale - Control Wheat & -672.6 &1124 & 10.00 & -0.598 & 0.9999 \\
-            MeJA Triticale - Insect Wheat & 1175.1 &1139 & 10.54 & 1.031 & 0.9866 \\
-            MeJA Triticale - MeJA Wheat & -231.5 &1125 & 10.03 & -0.206 & 1.0000 \\
-            Control Wheat - Insect Wheat & 1847.7 & 479 & 186.44 & 3.855 & 0.0057 \\
-            Control Wheat - MeJA Wheat & 441.0 & 443 & 186.04 & 0.995 & 0.9956 \\
-            Insect Wheat - MeJA Wheat &-1406.6 & 480 & 186.45 & -2.931 & 0.0982 \\
-\end{longtable}
-\end{center}
-}
+        \end{tabular}
+\end{table}
 
@@
 \begin{figure}[h]
+        \centering
+        \begin{subfigure}[b]{0.45\textwidth}
+                \centering
                 \includegraphics[width = \textwidth]{images/spp_si_content.png}
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}[b]{0.45\textwidth}
+                \centering
+                \includegraphics[width = \textwidth]{images/spp_phenolic_content.png}
+        \end{subfigure}
+
         \centering
@@
 \begin{figure}[h]
-        \includegraphics[width = \textwidth]{images/induction_plot.png}
         \centering
-        \caption{Boxplots showing leaf silicon content for four cereal species across three defense induction treatments.
-Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
-Leaves were sampled 18 hours after treatment, and were analyzed using XRF.}
+        \begin{subfigure}[b]{0.45\textwidth}
+                \centering
+                \includegraphics[width = \textwidth]{images/induction_plot_letters.png}
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}[b]{0.45\textwidth}
+                \centering
+                \includegraphics[width = \textwidth]{images/phenolic_induction_plot_letters.png}
+        \end{subfigure}
+        \caption{The response of leaf silicon content (left) and leaf phenolic content (right) for four cereal species across three defence induction treatments.
+We treated plants either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
+Leaves were sampled 18 hours after treatment, and were analysed for silicon using XRF or for phenolic content using the Fast Blue BB method.
+Dots and error bars are the mean $\pm$ 1 Standard Error.}
         \label{Fig:induction}
-\end{figure}
-
-\begin{figure}[h]
-        \includegraphics[width = \textwidth]{images/absorbance_boxplots.png}
-        \centering
-        \caption{Boxplots showing leaf phenolic content for four cereal species across three defense induction treatments.
-Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
-Leaves were sampled 18 hours after treatment, and were analyzed using the Fast Blue BB method.}
-        \label{Fig:absorbance_boxplots}
 \end{figure}
@@
         \centering
-        \caption{Scatter plot of the relationship between leaf silicon and leaf phenolic concentration}
+        \caption{The relationship between baseline (uninduced) silicon and phenolic content of leaves across four species of cereals.
+We measured leaf silicon content using XRF.
+We measured leaf phenolic content using the Fast Blue BB method.}
         \label{Fig:phe_si_scatter}
@@
 
-With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production.
-Over the past thirty years, increasing research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
+Over the past thirty years, plant silicon has emerged as an exciting possible tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
 Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}).
@@
 The soluble form of silicon, silicic acid (SiOH4) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM (\cite{epstein_anomaly_1994}).
-Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilization of silicon in decaying plant material (\cite{de_tombeur_silicon_2021-1}).
-Weathering of silicates releases a host of plant nutrients including aluminium, silicon, iron, and phophorus (\cite{de_tombeur_silicon_2021-1}).
-Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms.
-Plant roots can release carboxylates and phytosiderophores to weather phosphorus and silicon out of soil minerals.
-Along with Si and P mobilization, Mn is often released, and taken up by plants roots.
-Previous research has used leaf Mn content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far we are unaware of any studies looking for quantitative variation among genotypes of leaf Mn.
-If we could identify regions of the plant genome associated with variation in root weathering activity, we may be able to target this trait in breeding programs that improve nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
+Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilisation of silicon in decaying plant material (\cite{de_tombeur_silicon_2021-1}).
+Weathering of silicates releases a host of plant nutrients including aluminium, silicon, iron, and phosphorus (\cite{de_tombeur_silicon_2021-1}).
+Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms (\cite{de_tombeur_silicon_2021}).
+Plant roots can release carboxylates and phytosiderophores to weather phosphorus and silicon out of soil minerals (\cite{de_tombeur_silicon_2021-1}).
+Along with Si and P mobilisation, manganese is often released, and taken up by plants roots.
+Previous research has used leaf manganese content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far we are unaware of any studies looking for quantitative variation among genotypes of leaf manganese.
+If we could identify regions of the plant genome associated with variation in leaf manganese content, we could follow up with more targetted studies looking for genes that influence root carboxylate release.
+This trait could eventually be incorporated into breeding programs, improving nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
 
@@
 XRF works by using low-power x-rays to excite elements in the sample, and measures the resulting emitted light.
-One of the most exciting features of XRF is the fact that it can analyse multiple elements at once, allowing for broad characterization of the sample for most elements heavier than aluminum.
+One of the most exciting features of XRF is the fact that it can analyse multiple elements at once, allowing for broad characterisation of the sample for most elements heavier than aluminum.
 Though XRF is an established technique to measure plant silicon, its may also be used to measure other metals of interest, including manganese.
 In this study we use XRF to quantify variation in silicon and manganese content among a diversity panel of a wild ancestor of bread wheat, \textit{Aegilops tauschii}.
-This panel has publicly available sequence data, allowing us to perform a genome-wide association sutdy to link silicon and manganese variation to genotypic variation, laying the groundwork for future, more targetted, explorations of the genome to identify genetic controls over these traits, and hopefully develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
+This panel has publicly available sequence data \cite{gaurav_population_2022}, allowing us to perform a genome-wide association study to link silicon and manganese variation to genotypic variation, laying the groundwork for future, more targetted explorations of the genome.
+Identifying genetic controls over these traits will hopefully help develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
 
@@
 
-For this experiment, we used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites.
-Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment.
+For this experiment, we used the L2 panel of Aegilops tauschii from \textcite{gaurav_population_2022}.
+This panel is a 151 genotype subset of the full \textit{Aegilops tauschii} panel, selected for high diversity in nucleotide-binding leucine rich repeats receptor (NLR) genes.
+We grew the L2 panel at three different sites; two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalised seedlings in growth chambers prior to transplanting into pots.
 For full site details see Supplementary Table S1.
@@
 For the glasshouse environment, we started seedlings in growth chambers in January 2022.
-After 12 weeks, we moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks.
+After 12 weeks, we moved the seedlings to vernalisation chambers (4ºC, 8:16h light:dark) for eight weeks.
 We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with 0.17g of silicic acid (Tixosil 68B, Solvay).
@@
 To minimize cross-contamination between samples, we cleaned the pellet press and XRF device after each sample.
-We calibrated our measurments against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).
+We calibrated our measurements against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).
 
@@
 To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$).
-We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:
+We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) packages in R 4.2.2 (\cite{r_core_team_2022}) to regress leaf manganese content on leaf silicon content.
+We used a random effect of Genotype to control for genotype-specific differences in silicon content.
+We used the R package \verb|MuMIn| (\cite{barton_2023_mumin} to estimate the $R^2$ value of our mixed model.
 
-$Manganese \sim Silicon + (1|Genotype)$
-
-To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
+To perform the GWA analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
 For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
-As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
+As per \textcite{gaurav_population_2022}, to reduce the computational intensity of our analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
 For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers.
 We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.
-For peaks that were significant in at least two environments, we crossreferenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.
+For peaks that were significant in at least two environments, we cross-referenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.
 
@@
 
- Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%.
+ Silicon content in \textit{Aegilops tauschii} ranged from 0.784\% to 11.473\%.
 There were notable differences in silicon content based on the growing environment.
 The glasshouse plants averaged $1.450\% \pm 0.032 (SE)$, while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
-Manganese content varied from 36.5 ppm to 1296.5 ppm, and averaged $225.1 \pm 25.45 ppm (SE)$ in the glasshouse environment, and $ 71.05 \pm 2.571 ppm (SE)$ and $ 224.0 \pm 11.01 (SE)$ for the two outdoor growing environments.
-The results summarized in our rank plots (Figure \ref{Fig:rank_plots}) show that rank is variable for both silicon and manganese.
+Manganese content varied from 36.5 ppm to 1296.5 ppm, and averaged $225.1 \pm 25.45$ ppm (SE) in the glasshouse environment, and $ 71.05 \pm 2.571 ppm (SE)$ and $ 224.0 \pm 11.01 (SE)$ for the two outdoor growing environments.
+The results summarized in our rank plots (Figure \ref{Fig:rank_plots}) show that there are weak g.
 Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}).
-Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}).
+Overall, our analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}).
 For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance.
-However, when comparing accross environments, we found no regions that consistently exceeded the threshold significance.
+However, when comparing across environments, we found no regions that consistently exceeded the threshold significance.
 We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil, though absent in the glasshouse environment (Table \ref{tab:peaks_and_genes}).
+Of these genes, three had associated genes in NCBI.
+Two of the genes were uncharacterised non-coding RNA genes.
+The third gene codes for CASP-like protein 5B3.
 
@@
 Our rank plots revealed that relative content of both silicon and manganese is not stable across environments.
-This suggests strong genotype by environment interactions, likely reduced our ability to detect genetic contribution to the silicon and manganese phenotype.
-Interestingly, we detected a number of peaks for leaf managese content, though many of these were only found in a single environment.
+This suggests strong genotype by environment interactions, which likely reduced our ability to detect genetic contributions to the silicon and manganese phenotype.
+Interestingly, we detected a number of peaks for leaf manganese content, though many of these were only found in a single environment.
 Though we detected no peaks present in all three growing environments, we detected 11 regions that were significant in both outdoor environments.
@@
 In the glasshouse environment, we grew plants in standard potting mix, which is high in organic matter with a relatively low inorganic component.
-In this soil context, root exudation is unlikely to yeild appreciable nutrient liberation.
+In this soil context, root exudation is unlikely to yield appreciable nutrient liberation.
 The high manganese content found in the glasshouse plants is most likely due to the watering regime they experienced.
@@
 
-Our observations of a correlation between silicon and manganese content lend support to the idea of root-exudate based silicon aquisition (\cite{de_tombeur_silicon_2021-1}).
-It is possible that this positive correlation is due to a nutrient effect more generally, however further regressions show that silicon only shows positive correlations with iron and manganese, both of which are mobilized by root exudation (Supplementary Fig [Number]) (\cite{de_tombeur_silicon_2021-1}).
-Unfortunately, the two genes we did detect as associated with manganese content are both non-coding RNA, reducing the likelihood that they are involved in root exudation or manganese uptake.
+Our observations of a correlation between silicon and manganese content could be evidence of root-exudate driven silicon uptake.
+It is possible that this positive correlation is due to a nutrient effect more generally, where manganese and silicon covary in availability in the soil.
+However, further regressions show that silicon only shows positive correlations with iron and manganese, both of which are mobilized by root exudation (Supplementary Fig [Number]) (\cite{de_tombeur_silicon_2021-1}).
+Unfortunately, only one MTA is located in a coding-region of the genome.
+LOC109751197 codes for a CASP-like protein, involved in the casparian strip.
+The presence of a SNP in this region is negatively correlated with manganese content.
+CASP-like proteins are involved in cell membrane modification, and can form transmembrane scaffolds (\cite{roppolo_functional_2014}).
+This leads to a puzzling interpretation, where defects in a gene involved membrane modification (likely reducing permeability) decrease manganese uptake.
+The uncharacterized nature of the gene in question makes more indepth interpretation of the results difficult, but provides a target for future investigations into genomic controls over manganese uptake.
+
 Silicon interacts with manganese in the leaf, binding manganese to the cell wall and reducing manganese concentrations in the cell symplast (\cite{rogalla_role_2002}).
-Root exudation is most important under limited nutrients, and silicon-exudate dynamics have been extensively studied in well-weathered soils (\cite{lambers_plant_2008,de_tombeur_shift_2021,de_tombeur_silicon_2021-1}).
-In agricultural soils, it is unlikely that plants would be exposed to such nutrient poor conditions, and so the specific dynamics of root exudation and silicon mobilization deserve further investigation.
-Root exudation also results in increased soil carbon content, and there is interest in implementing cropping and fertilization systems with an explicit focus on improving soil carbon through root exudation (\cite{cornelis_soil_2022}).
+Root exudation is most important under limited nutrients, and silicon-exudate dynamics have been best characterized in well-weathered soils (\cite{lambers_plant_2008,de_tombeur_shift_2021,de_tombeur_silicon_2021-1}).
+In agricultural soils, it is unlikely that plants would be exposed to such nutrient poor conditions, and so the specific dynamics of root exudation and silicon mobilisation deserve further investigation.
+Root exudation also results in increased soil carbon content, and there is interest in implementing cropping and fertilisation systems with an explicit focus on improving soil carbon through root exudation (\cite{cornelis_soil_2022}).
 In this context, understanding how increased root exudation may impact both silicon and metal uptake in plants is crucial, and future studies should pursue similar genetic approaches under limited phosphorus regimes to try and detect genes associated with root exudation activity.
 
-Depsite our inability to detect QTLs that influence silicon or manganese content, this field of study still holds immense promise.
-We implemented a study design that we hoped would allow generalizations about the genetic regions we hoped to find.
+Depsite our inability to detect QTLs that reliably influence silicon or manganese content, this field of study still holds immense promise.
+We implemented a study design that we hoped would allow generalisations about the genetic regions we hoped to find.
 However, the use of multiple environments may have obscured QTLs that we may have otherwise detected in a single environment.
-The use of the diversity panel allowed for a rapid and low-cost first search for loci asscociated with silicon and manganese content, but may have also limited our power to detect loci.
+The use of the diversity panel allowed for a rapid and low-cost initial search for loci asscociated with silicon and manganese content, but may have also limited our power to detect loci.
 The panel we used was a subset of the larger \textit{Aegilops tauschii} panel, and ours was selected specifically for diversity in disease-resistance genes.
@@
 Alternatively, future researchers could attempt to create a mapping population.
-By using the silicon and manganese content rankings generated by this study, one could create a cross of a low and high genotype and use a traditional qtl mapping approach with the recombinant population.
+By using the silicon and manganese content rankings generated by this study, one could create a cross of a low and high genotype and use a traditional QTL mapping approach with the recombinant population.
 This approach might allow for more sensitivity in the analysis by reducing the overall genetic variation between genotypes in population, highlighting the genetically driven variation in silicon content.
@@
         \centering
-        \begin{subfigure}[b]{\textwidth}
-                \centering
-                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/tf_si_plot.png}
-                \label{Fig:tf_si_peak}
-        \end{subfigure}
-        \begin{subfigure}[b]{\textwidth}
-                \centering
-                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/sc_si_plot.png}
-                \label{Fig:sc_si_peak}
-        \end{subfigure}
-        \begin{subfigure}[b]{\textwidth}
-                \centering
-                \includegraphics[scale=0.3]{images/gwas_plots/svgtopng/gh_si_plot.png}
-                \label{Fig:gh_si_peak}
-        \end{subfigure}
+        \includegraphics[width=0.8\textwidth]{images/gwas_plots/svgtopng/si_manhattan_plots.png}
         \caption{Manhattan plots of the results of our GWAS analysis for leaf silicon content in \textit{Aegilops tauschii}.
 Red dots indicate a negative correlation, while blue dots denote a positive correlation.
-The top two plots show the results for the outdoor growing environments, and the bottom plot is the glasshouse environment.
+The top plot show the results for glasshouse environment, while the bottom two plots are the outdoor environments.
 The dotted line represents the bonferroni-corrected significance threshold.}
@@
 
+\begin{figure}[h]
+        \centering
+        \includegraphics[width = 0.8\textwidth]{images/gwas_plots/svgtopng/mn_manhattan_plot.png}
+        \label{Fig:mn_peak_plot}
+        \caption{Manhattan plots of the results of our GWAS analysis for leaf manganese content in \textit{Aegilops tauschii}.
+Red dots indicate a negative correlation, while blue dots denote a positive correlation.
+The top plot show the results for glasshouse environment, while the bottom two plots are the outdoor environments.
+The dotted line represents the bonferroni-corrected significance threshold.}
+\end{figure}
+
+\mycomment{
 \begin{figure}[h]
@@
 The dotted line represents the bonferroni-corrected significance threshold.}
-        \label{Fig:mn_peak_plot}
+        \label{Fig:mn_peak_plot_old}
 \end{figure}
+}
 
@@
 
+\chapter{Conclusion}
+\section{Research Outcomes}
+
+My aim with this thesis was to advance our understanding of how silicon-based plant traits may be applied to agriculture.
+To this end, I identified two separate problems that currently limit uptake and dissemination of silicon research into the broader agricultural context: 1) limited applicability of fundamental reserach into inducded silicon uptake to crop plants, and 2) a limited understanding of the genetic basis of variation in silicon content among genotypes of a common species.
+Based on the exciting finidngs of rapid silicon accumulation in \textit{Brachypodion distachyum}, I attemped to replicate this pattern of silicon uptake in cereal crops, in the hopes that this would spur further research into silicon-based crop defenses.
+In the growing conditions I used, I did not observe rapid silicon accumulation, but did observe variation in silicon and phenolic compound levels among the four crop species I tested.
+We also documented the significant variation among common cultivars of these cereal species.
+
+For my GWA study, we observed significant variation in silicon and manganese content between sites.
+Though we failed to find robsut genotype-phenotype links for either silicon or manganese, we did identify one gene that is associated with variation in leaf manganese content.
+Additionally, by working with a previously developed diversity panel, we hope that our findings will be easily integrated into future work on this topic.
+
+\section{Limitations and Future Directions}
+
+A number of limitations with my study design and experimental methods should be noted.
+Though I was ultimately unsucessful in my attempts to induce rapid silicification, I identified a number of factors that could have led to the discrepancies between previous research and my own.
+We utilized a single herbivore species in our treatments.
+Plants modulate their defensive responses based on the identify of the attacker, and it is likely that using different species or multiple species would yeild different outcomes.
+Some of the insects in my treatments did not initiate feeding, despite our best efforts to maximise the likelihood of herbivory, and this reduced our sample size for the experiment.
+Finally, it is likely that the growth media we used for the study imposed some silicon limitation on our plants.
+This limitation may have prevented rapid uptake of silicon, as there might just not have been enough in the soil to meet the demands of the plant.
+Future reserach should explicitly investigate the silicon availability of commercially available plant growth media, as nearly all plants can respond positively to silicon.
+Additionally, we used a single time point to test for rapid silicon accumulation.
+It is possible that we pre-empted silicon uptake in our plants, and had we waited longer, we would have seen a clear signal of silicon uptake in response to our treatments.
+Future studies would provide immense utility to the field if they were to investigate silicon uptake at regular intervals between 6 and 48 hours post induction, as we still lack replicated data on this time interval.
+
+For the GWA study, we may have limited our ability to detect important QTLs through the use of multiple growing environments.
+Though multiple growing environments are often used to prove the durability of a given QTL through GxE interations, the highly exploratory nature of our work would have benefitted from a deeper and narrower experimental design, where we had more replicates in a single environment.
+We are left with few markers to investigate, but future work could take the genotypes we identified as having high and low silicon, and develop a mapping population using experimental crosses.
+The increased level of control this design would bring might allow for the discovery of QTLs associated with silicon content.
+Similar studies in wheat and rice have yeilded promising results, and there is no reason to believe that \textit{Aegilops tauschii} would not have similar genes that influence silicon uptake.
+
 \end{document}

### 2023-06-09 src/R_linear_model.R
not sure

@@ -20,9 +20,21 @@ si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
 si_filtered$Genotype <- as.factor(si_filtered$Genotype)
 si_filtered$Induction <- as.factor(si_filtered$Induction)
 si_filtered$Species <- as.factor(si_filtered$Species)
-si_mod <- lmerTest::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Species/Genotype), si_filtered)
+si_mod <- lmerTest::lmer(log(Si_ppm/10000) ~ Induction * Species + mass_g + (1|Genotype), si_filtered)
 si_mod <- lme4::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Genotype), si_filtered)
-anova(si_mod, type = 2)
+anova(si_mod, type = 3)
+e1 <- emmeans(si_mod, ~ Species * Induction)
+e2 <- contrast(e1, interaction = c("pairwise", "pairwise"))
+my.aes = list(point = list(size = 4))
+e3 <- pwpp(e1, aes = my.aes) + ggplot2::theme_bw( base_size =12) + ggplot2::labs(y = "Contrast")
+e4 <- pwpm(e1)
+pairs(e1)
+
+model_means_cld <- CLD(object = e1,
+                       adjust = "Tukey",
+                       Letters = letters,
+                       alpha = 0.05)
+ggplot2::ggsave("/home/isaac/Desktop/test_fig.png", width = 30, height = 20, units = "cm", dpi = 600)
 summary(aov(Si_ppm/10000 ~ Induction * Species + mass_g, si_filtered))
 summary(si_mod)
 lmerTest::anova(si_mod)
@@ -81,11 +93,22 @@ abdf_filtered  <- absorbancedf[absorbancedf$isDamaged != "Undamaged",]
 abdf_filtered$Genotype <- as.factor(abdf_filtered$Genotype)
 abdf_filtered$Species <- as.factor(abdf_filtered$Species)
 abdf_filtered$Induction <- as.factor(abdf_filtered$Induction)
-ab_mod <- lme4::lmer(mcabsorbance ~ Induction * Species + mass_g + (1|Genotype), abdf_filtered)
-summary(ab_mod)
-Anova(ab_mod)
+ab_mod <- lmer(log(mcabsorbance) ~ Induction * Species + mass_g + (1|Genotype), abdf_filtered)
+anova(ab_mod)
[…]

### 2023-06-09 src/cereal_induction.jl

@@ -1,3 +1,4 @@
+begin
 using CSV
 using MLDataUtils: rescale!
 using Plots
@@ -10,7 +11,16 @@ using DataFrames
 using Measurements
 using Statistics
 using ColorSchemes
+end
 
+
+
+function standarderror(col) 
+    sd = std(col)
+    denom = sqrt(length(col)-1)
+    se = sd/denom
+    return se
+end
 ### Combine 2019-2021 insured acreages of grain crops
 ### Data sourced from the Canadian Grain Commission https://www.grainscanada.gc.ca/en/grain-research/statistics/varieties-by-acreage/
 
@@ -162,7 +172,7 @@ summ_stats_geno.Si_ppm_se = summ_stats_geno.Si_ppm_std ./ sqrt.(summ_stats_geno.
 gsi_spp = groupby(analysis_data_noinduction, [:OfficialName])
 spp_si_contents_noind = combine(gsi_spp, [:Si_ppm, :Species] => ((si, sp) -> (sppMEANsi = mean(si), sppSEsi = standarderror(si), spp = first(sp))) => AsTable)
 spp_si_contents_noind.GenotypeID = 1:12
-sort!(spp_si_contents_noind, [:spp, :sppMEANsi])
+sort!(spp_si_contents_noind, [:spp, :OfficialName])
 
 spp_palette = palette([
     :palegreen, 
@@ -234,12 +244,14 @@ markersize = 8,
 #legend = spp_si_contents_noind.Genotype
 labels = spp_si_contents_noind.spp[10])
 plot!(xrotation =45)
-plot!(xlabel = "Cultivar (Species)", ylabel = "Silicon Leaf Content (%)")
-end
+plot!(xlabel = "Cultivar", ylabel = "Leaf Silicon Content (%)")
 plot!(size = (600,600), dpi = 600)
 plot!(xtickfontsize =10)
+plot!(xlabelfontsize = 14)
+plot!(ylabelfontsize = 14)
+plot!(legendfontsize = 12)
 plot!(grid=false)
-
+end
 savefig(spp_mean_si_plot, "./manuscript/images/spp_si_content.png")
 
 
@@ -282,6 +294,55 @@ png(gboxplot, "manuscript/images/induction_plot")
 test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = repeat(["D", "E", "F"], inner = 9, outer = 1), yvals = rand(Normal(0,1), 27))
[…]

### 2023-06-13 manuscript/main.tex
better graphs and final data

@@
 We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction}, Table \ref{Tab:phe_params}).
-Oats had the highest average leaf phenolic content [$val \pm err$], while barley had the lowest average phenolic content [$val \pm err$].
+Oats had the highest average leaf phenolic content [$31.9 \pm 3.66 mg/kg$], while barley had the lowest average phenolic content [$12.9 \pm 1.82 mg/g$].
 We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).
@@
 In our study, oats had the highest phenolic concentration, and the lowest silicon content.
-The relative difference in the phenolic and silicon content of oats and wheat in our study [was similar] to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study.
+The relative difference in the phenolic and silicon content of oats and wheat in our study was larger than largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study.
 Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (eg.
@@
         \centering
-        \begin{subfigure}[b]{0.45\textwidth}
+        \begin{subfigure}[b]{0.49\textwidth}
                 \centering
@@
         \end{subfigure}
-        \hfill
-        \begin{subfigure}[b]{0.45\textwidth}
+        \begin{subfigure}[b]{0.49\textwidth}
                 \centering
@@
         \centering
-        \caption{Baseline (uninduced) silicon content (\% dry matter)in the cereal cultivars used in this study. Cultivar species is indicated by point color. Error bars are 1 Standard Error.}
+        \caption{Baseline (uninduced) silicon content (\% dry matter) and phenolic content (mg/kg) in the cereal cultivars used in this study.
+Cultivar species is indicated by point color.
+Error bars are 1 Standard Error.}
         \label{Fig:baseline_si}
@@
 \subsection{Plant harvest and sample preparation}
-
-When the plants had reached maturity, we harvested the entire above-ground portion of each plant.
-For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022.
-We placed harvested material in labelled paper bags, and dried it a 60ºC drying oven for 48 hours.
-To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers.
-Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation.
-We picked leaves until approximately 200mg of dry leaf was collected.
-Some plants did not yield enough leaf tissue to meet to 200mg threshold.
-We pooled leaf material from within sites.
-For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample.
-Overall, we were left with 359 samples across three sites, making approximately 115 useable genotypes per site.
-We packed dried leaf samples in 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz.
-We stored the resulting leaf powder sealed until XRF analysis.
 
@@
 Red dots indicate a negative correlation, while blue dots denote a positive correlation.
-The top plot show the results for glasshouse environment, while the bottom two plots are the outdoor environments.
+The top plot shows the results for the glasshouse environment, while the bottom two plots are the outdoor environments.
 The dotted line represents the bonferroni-corrected significance threshold.}

### 2023-06-13 src/absorbance_calibration_model.jl

@@ -1,0 +1,9 @@
+using DataFrames
+using GLM
+using CSV
+
+calibdata = CSV.read("./data/real_data/absorbances_calibration.csv", DataFrame)
+
+fm = @formula (ppm ~ absorbance)
+myreg = lm(fm, calibdata)
+plot(calibdata.absorbance, calibdata.ppm, seriestype = :scatter)

### 2023-06-13 src/cereal_induction.jl

@@ -11,15 +11,14 @@ using DataFrames
 using Measurements
 using Statistics
 using ColorSchemes
-end
-
-
+using Measures
 
 function standarderror(col) 
     sd = std(col)
     denom = sqrt(length(col)-1)
     se = sd/denom
     return se
+end
 end
 ### Combine 2019-2021 insured acreages of grain crops
 ### Data sourced from the Canadian Grain Commission https://www.grainscanada.gc.ca/en/grain-research/statistics/varieties-by-acreage/
@@ -144,6 +143,7 @@ analysis_data = CSV.read("data/real_data/biomass_si_data.csv", DataFrame)
 # Some summary stats on the data
 
 #############
+begin
 my_names = unique(analysis_data.Genotype)
 official_names = [
     "CS Camden", 
@@ -160,7 +160,8 @@ official_names = [
     "Tyndal" ]
 
 cultivar_dict = Dict(my_names .=> official_names)
-
+end
+begin
 transform!(analysis_data, :Genotype => ByRow(x->cultivar_dict[x]) => :OfficialName)
 analysis_data_noinduction = filter(row -> row.Induction == "Control", analysis_data)
 analysis_data_noinduction.Si_ppm = analysis_data_noinduction.Si_ppm./10000
@@ -168,12 +169,13 @@ gdat_geno = groupby(analysis_data_noinduction, :Genotype)
 summ_stats_geno = combine(gdat_geno, :Si_ppm => mean, :Si_ppm => std, nrow)
 summ_stats_geno.Si_ppm_se = summ_stats_geno.Si_ppm_std ./ sqrt.(summ_stats_geno.nrow)
 #CSV.write("./data/real_data/summary_statistics_genotype.csv", summ_stats_geno)
-
+end
+begin
 gsi_spp = groupby(analysis_data_noinduction, [:OfficialName])
 spp_si_contents_noind = combine(gsi_spp, [:Si_ppm, :Species] => ((si, sp) -> (sppMEANsi = mean(si), sppSEsi = standarderror(si), spp = first(sp))) => AsTable)
 spp_si_contents_noind.GenotypeID = 1:12
 sort!(spp_si_contents_noind, [:spp, :OfficialName])
-
+end
 spp_palette = palette([
     :palegreen, 
     :lime, 
@@ -187,7 +189,7 @@ spp_palette = palette([
     :orange, 
[…]

### 2023-06-13 src/phenolicdataprep.jl

@@ -2,6 +2,8 @@ using CSV
 using DataFrames
 using Statistics
 begin
+#I first cut out each table of the plate data, see the attached plate1_matrix.csv file to see the format
+#import each csv as a matrix, without the header
 p1m = CSV.File("data/real_data/plate1_matrix.csv", header = false) |> Tables.matrix
 p2m = CSV.File("data/real_data/plate2_matrix.csv", header = false) |> Tables.matrix
 p3m = CSV.File("data/real_data/plate3_matrix.csv", header = false) |> Tables.matrix
@@ -13,8 +15,9 @@ p8m = CSV.File("data/real_data/plate8_matrix.csv", header = false) |> Tables.mat
 p9m = CSV.File("data/real_data/plate9_matrix.csv", header = false) |> Tables.matrix
 p10m = CSV.File("data/real_data/plate10_matrix.csv", header = false) |> Tables.matrix
 
+#make a list of the matrices (in R it should be c(p1m, p2m, ...))
 matlist = [p1m, p2m, p3m, p4m, p5m, p6m, p7m, p8m, p9m, p10m]
-
+#These are matrices of which samples were in each well. See attached file
 p1k = CSV.File("data/real_data/plate1_key.csv", header = false) |> Tables.matrix
 p2k = CSV.File("data/real_data/plate2_key.csv", header = false) |> Tables.matrix 
 p3k = CSV.File("data/real_data/plate3_key.csv", header = false) |> Tables.matrix
@@ -25,28 +28,47 @@ p7k = CSV.File("data/real_data/plate7_key.csv", header = false) |> Tables.matrix
 p8k = CSV.File("data/real_data/plate8_key.csv", header = false) |> Tables.matrix
 p9k = CSV.File("data/real_data/plate9_key.csv", header = false) |> Tables.matrix
 p10k = CSV.File("data/real_data/plate10_key.csv", header = false) |> Tables.matrix
-
+#make another list
 keylist = [p1k, p2k, p3k, p4k, p5k, p6k, p7k, p8k, p9k, p10k]
 end
-function builddf(ml, kl)
+```
+This function compiles a dataframe from a list of absorbance matrices and a list of key matrices
+
+```
+
+
+function builddf(ml, kl) 
+    #make an empty list for keys
     keys = String[];
+    #make an empty list for the absorbance values
     vals = Float64[];
+    #error check to ensure an equal number of key and value matrices
     if length(ml) != length(kl)
         print("Unequal list lengths")
     else 
[…]

### 2023-08-23 manuscript/coverpage.tex
I think it is ready!

\thispagestyle{plain}
\setcounter{page}{2}
{
%\setlength{\parskip}{12pt}
%\begin{adjustwidth}{0pt}{0pt}
\vspace{1cm}

\noindent The following individuals certify that they have read, and recommend to the Faculty of Graduate and Postdoctoral Studies for acceptance, the thesis entitled:

\vspace{1cm}

\noindent \uline{Ecological and Big Genetic Drivers of Silicon Accumulation in Cereal Crops\hfill}
%\hfill\rule{\textwidth}{0.4pt}

\vspace{1cm}

\noindent submitted by
%\rule{\textwidth}{0.4pt}
\noindent \uline{Isaac Peetoom Heida\hfill}
\noindent in partial fulfilment of the requirements for

\vspace{1cm}
\noindent the degree of \uline{Master of Science\hfill}

\vspace{1cm}
\noindent in \uline{Plant Science\hfill}

\vspace{1cm}

\noindent \textbf{Examining Committee}:

%\vspace{1cm}

\noindent \uline{Juli Carrillo, Professor, Plant Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisor}

\vspace{1cm}

\noindent \uline{Gurcharn Singh Brar, Professor, Plant Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisory Committee Member}

\vspace{1cm}

\noindent \uline{Jean-Thomas Cornelis, Professor, Soil Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisory Committee Member}
%\end{adjustwidth}
}

### 2023-08-23 manuscript/main.tex
I think it is ready!

@@
-\listfiles
-\documentclass[12pt, letterpaper, ]{report}
+\documentclass[12pt, letterpaper]{report}
 \usepackage[style=authoryear, maxnames=2, maxbibnames=99, citestyle=apa]{biblatex}
@@
 \addbibresource{internet_stuff.bib}
-%\addbibresource{induction.bib}
-%\addbibresource{gwas_silicon.bib}
 \usepackage{caption}
@@
 \usepackage{setspace}
+\usepackage{ulem}
 \usepackage{url}
@@
 \usepackage{fancyhdr}
+\usepackage{chemformula}
+\usepackage[hidelinks]{hyperref}
 \pagestyle{fancy}
 \fancyhf{}
+\renewcommand{\headrulewidth}{0pt}
 \fancyfoot[R]{\thepage}
 \fancypagestyle{plain}{%
+\fancyhf{}%
     \renewcommand{\headrulewidth}{0pt}%
+    \fancyfoot[R]{\thepage}%
+}
+
+% Modify the fancy page style for chapters
+\fancypagestyle{fancy}{%
     \fancyhf{}%
@@
 \newcommand{\mycomment}[1]{}
-% Change chapter heading font size to 14pt
-\titleformat{\chapter}{\Large\bfseries}{\thechapter.}{20pt}{\Large}
 
-% Change section heading font size to 12pt
+\titleformat{\chapter}[block]{\Large\bfseries}{Chapter \thechapter.}{20pt}{\raggedright\Large}
+\titlespacing{\chapter}{0pt}{5pt}{5pt}
+% Adjust section spacing to prevent orphans
 \titleformat{\section}{\large\bfseries}{\thesection}{1em}{\large}
+\titlespacing*{\section}{0pt}{\baselineskip}{0.5\baselineskip}
 
-% Change subsection heading font size to 10pt
+% Adjust subsection spacing to prevent orphans
 \titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{\normalsize}
+\titlespacing*{\subsection}{0pt}{\baselineskip}{0.5\baselineskip}
 
-\linenumbers
 \graphicspath{./images/}
@@
         letterpaper,
-        top = 1.27cm,
-        bottom = 1.27cm,
+        top = 2.54cm,
+        bottom = 2.54cm,
         left = 2.54cm,
@@
 \pagenumbering{roman}
+\setcounter{page}{2}
 \input{title.tex}
+\input{coverpage.tex}
+% Rest of your document
 
-\tableofcontents
-\newpage
-\listoftables
-\newpage
-\listoffigures
 \chapter*{Abstract}
@@
 
-One of the most promising developments to arise over the past 30 years is the potential of silicon to improve crop plant performance.
-With benefits to multiple dimensions of crop performance, silicon may be a key tool to guard crop production against uncertain future growing conditions.
-Our ability to mobilize silicon-based crop improvement strategies depends on a thorough understanding of the proximate and ultimate causes of silicon accumulation, including both the ecological and genetic interactions that can trigger increased uptake.
-In this thesis, I extend recent advances in our understanding of silicon ecology in cereal crops, testing for the presence of rapid silicification in common Canadian crops, and use a  genome-wide association study to identify genetic markers associated with high silicon content.
-We failed to observe rapid silicification in the tested crops.
-This may could be a result of a soil environment incapable of supplying sufficient silicon resources for rapid accumulation, or could reflect the effect of herbivore identity on specific defensive outcomes.
-In our genetic analysis, we failed to find genetic markers with significant associations to silicon content, but did find a marker that had a significant correlation with manganese content.
-Future work in this field might develop a mapping population using crosses from high and low silicon genotypes in this study, yielding experimental designs with increased power and thus may find QTLs associated with silicon content.
-Applications silicon research into crop production techniques are still limited by crucial gaps in our understanding of how ecology and genetics control silicon uptake, but this work provides yet another stepping stone towards the broader adoption of silicon into agricultural systems.
+Over the past 30 years, the potential of silicon to improve crop plant performance has gained increasing recognition in the field of plant science.
+Silicon may be a key tool to guard crop production against uncertain future growing conditions by improving crop tolerance to abiotic and biotic stressors.
+In this thesis, I take recent advances in our understanding of silicon ecology and extend them into cereal crops, testing for the presence of rapid ($<24$ hour) silicification in common Canadian crops after defence induction events, including herbivory, and used a genome-wide association study to identify potential genetic markers associated with high silicon content.
+We found that overall, simulated herbivory, but not cricket herbivory, increased silicon leaf content.
+However, at the species level, we found no significant increases in silicon content in response to herbivory treatments.
+This may be due to a silicon-poor soil environment or could reflect the effect of herbivore identity on specific defensive outcomes.
+In our genetic analysis, we found no genetic markers with significant associations to silicon content but did find a marker that had a significant correlation with manganese content.
+Our limited results may reflect the polygenic nature of silicon content in plants, or strong environmental effects.
+Future studies could build upon our results by developing a mapping population using genotypes we identified as representing the extremes of silicon content.
+Applications of silicon research into crop production techniques are still limited by crucial gaps in our understanding of how ecology and genetics control silicon uptake, but this work provides yet another steppingstone towards the broader adoption of silicon into agricultural systems.
 
@@
 
-Silicon provides tremendous benefits to plant health, but is not yet widely utilized in agriculture despite it being highly abundance and naturally occurring in most soils.
-Silicon acts a defensive compound against insects, and also helps plants deal with stressful growing conditions, but is currently not widely applied in agriculture.
+Silicon provides tremendous benefits to plant health but is not yet widely utilized in agriculture despite it being highly abundant and naturally occurring in most soils.
+Silicon acts as a defensive compound against insects, and helps plants deal with stressful growing conditions, but is not currently widely applied in agriculture.
 Using cereal crops and a wild ancestor of wheat, I attempted to identify ecological and genetic factors that drive silicon content.
-We failed to find genetic regions that exerted control over silicon content, but observed large variations in silicon content between the growing sites.
-We did not observe rapid silicon uptake in our tested cereal crops, complicating our understanding of when rapid uptake is achieved by plant species.
-Continuing to build upon this research will help move silicon crop technologies closer to feasibility, and could provide an environmentally friendly strategy to reduce pesticide use while maintaining yields.
+We failed to find genetic regions that exerted control over silicon content but observed large variations in silicon content between the growing sites.
+We did not observe short-term silicon uptake in our tested cereal crops, complicating our understanding of when short-term silicon accumulation is achieved by plant species.
+Continuing to build upon this research will help move silicon crop technologies closer to feasibility and could provide an environmentally friendly strategy to reduce pesticide use while maintaining yields.
 
@@
 Daria Reshetniak, Paul Fisher, Lucas Friesen, Katie Pryer, Dr.
-Kinga Treder, Chelsea Gowten, Dennis Chiu, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
-I use the first person plural for chapters two and three as they are intended for submission to a peer-reviewed journal.
+Kinga Treder, Chelsea Gowton, Dennis Chiu, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
+I use the first-person plural for chapters two and three as they are intended for submission to a peer-reviewed journal.
+
+\newpage
+\tableofcontents
+\newpage
+\listoftables
+\newpage
+\listoffigures
 
@@
 Dr.
-Juli Carrillo has guided me to become a more well-rounded scientist and person, and has always believed in my abilities.
+Juli Carrillo has guided me to become a more well-rounded scientist and person and has always believed in my abilities.
 I thank Dr.
@@
 Dr.
-Gurcharn Singh Brar lent his time and lab to me to support the work, and stoked my passion for grasses from a new angle.
-The members of the PIEE lab were more than generous with their time, and their continued interest in my project and general convivatilty provided the encouragement I needed to keep going with my research.
+Gurcharn Singh Brar lent his time and lab to me to support the work and stoked my passion for grasses from a new angle.
+The members of the PIEE lab were more than generous with their time, and their continued interest in my project and general conviviality provided the encouragement I needed to keep going with my research.
 In some ways this work is the culmination of a lifetime of support from my family, as they encouraged my curiosity about the natural world and sought out opportunities for me when I was too young (or too dumb) to seek them out for myself.
@@
 
-\chapter*{Decication}
+\chapter*{}
 \addcontentsline{toc}{chapter}{Dedication}
 
-To my Auntie Margaret, for showing me where the wetland was all those years ago.
+\noindent{\Large\textit{To my Auntie Margaret, for showing me where the wetland was all those years ago.}}
 
@@
 
-Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass (\cite{holland_41_2014}).
-Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as the plant absorbs water (\cite{debona_silicons_2017}).
-Silicates occur in a variety of forms, and vary in their plant availability.
-Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available (\cite{fraysse_surface_2009}).
-As soils age, biogenic amorphous silicates, such as diatom testes and plant phytholiths, comprise an increasing share of the plant-available silicon (\cite{de_tombeur_plants_2020}).
-Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of their xylem into solid masses in the leaves and stems.
-Within the soil environment, silicon commonly exists in both crystalline (geologic) and amorphous (biogenic) forms (\cite{haynes_contemporary_2014}).
+Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (\ch{SiO2}), comprising about 60\% of the crust by mass (\cite{holland_41_2014}).
+Nearly all terrestrial plants grow in soils containing silicon, and thus absorb at least nominal amounts through passive transport as the plant absorbs water (though many plants have active transport) (\cite{debona_silicons_2017}).
+Silicates occur in a variety of forms and vary in their plant availability.
+Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of \ch{SiO2} are more available (\cite{fraysse_surface_2009}).
+Plants interact with silicon on a variety of levels, mobilizing it from soil aggregates, transporting it into and throughout their bodies, and finally precipitating it out of the xylem into solid masses in the leaves and stems.
+Within the soil environment, silicon commonly exists in both crystalline (geologic) and amorphous (biogenic) forms (Haynes 2014).
 Amorphous silicates can derive from previous plant material that has decayed in the soil, but also from marine and aquatic organisms such as diatoms.
-Globally, the silicon cycle involves silicates weathering out of terrestrial sediments, moving along water courses, and eventually being deposited in the sea, where it is incorporated into various plankton species, and eventually deposited in seafloor sediments.
-The continual exodus of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil (\cite{de_tombeur_plants_2020}).
+Young soils typically have high fractions of silicon-bearing materials, including clay minerals, though plant-available silicon peaks in middle-age soils (~120,000 years old) (\cite{de_tombeur_silicon_2020}).
+The continual leaching of silicon from terrestrial sediments over geologic timescales means as ecosystems age, plants become more and more central in the local silicon cycle, with much of the silicon in living plant tissue being recycled from previous plant material decaying in the soil (\cite{de_tombeur_plants_2020}).
 
@@
 
-One of the most important advances in plant-silicon research was understanding the mechanisms through which plants acquire and transport silicon to specific location within the plant (\cite{coskun_controversies_2019}).
-Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}).
-In highly weathered soils with low nutrient availabilities, plants take a more active role in liberating nutrients, including silicon, for uptake.
-Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients such as phosphorus and silicon from soil aggregates, increasing their availabilities for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
+One of the most important advances in plant-silicon research was understanding the mechanisms through which plants acquire and transport silicon to specific locations within the plant (\cite{coskun_controversies_2019}).
+Silicon’s most common form in soil solution is silicic acid (\ch{H_4SiO_4}), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}).
+In highly weathered soils with low nutrient availability, plants take a more active role in liberating nutrients, and other elements, including silicon, for uptake.
+Organic acids and chelating agents, exuded from plant roots, pry tightly bound nutrients and beneficial elements such as phosphorus and silicon from soil aggregates, increasing their availability for uptake into the root system (\cite{de_tombeur_silicon_2021-1}).
 Once dissolved in soil solution, silicon can be taken up by active or passive transport by the root system.
 While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families.
+
 Research in rice has identified four gene products that actively transport silicon into and through the plant body.
@@
 Throughout the plant kingdom, clades vary significantly in their silicon content and expression of silicon transporter proteins (\cite{ma_chapter_2001}).
-The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice (\cite{yamaji_transporter_2008}).
-Breeding for silicon content, improved exudate-based silicon mobilisation and silicon use-efficiency in crop plants may be crucial to improving crop performance under a changing climate (\cite{christian_breeding_2022,de_tombeur_silicon_2021}).
-However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant.
-Further investigations into how genotypic variation is reflected in the silicon content of plants can aid in the discovery of new genes involved in silicon accumulation, and may provide targets for silicon breeding programs.
+The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice (\cite{yamaji_transporter_2008,yamaji_orchestration_2015}).
+Breeding for silicon content, improved exudate-based silicon mobilisation and silicon use-efficiency in crop plants may be crucial to improving crop performance under a changing climate (\cite{de_tombeur_silicon_2021,christian_breeding_2022}).
+However, we still have a relatively poor understanding surrounding how genetics influence the silicon phenotype of a plant.
+Further investigations into how genotypic variation is reflected in the silicon content of plants can aid in the discovery of new genes involved in silicon accumulation and may provide targets for silicon breeding programs.
 
@@
 
-Plants deposit silicon in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}).
-Silicon deposits taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
-In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}).
+Plants deposit silicon in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021-1}).
+Silicon deposits show taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}).
+In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity (\cite{stromberg_functions_2016}).
 The use of silicon as a structural component represents a highly energetically efficient strategy, as silicon is ~10x cheaper on an energy unit basis to produce than lignin (\cite{stromberg_functions_2016}).
 Stem silicification has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation has been found to reduce the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}).
-In leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf (\cite{simpson_still_2017}).
-This overall toughness, and abundance of phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}).
-Leaf silicon concentrations are negatively correlated with relative growth rate and survival in a number of invertebrate herbivores (\cite{juma_influence_2015,massey_silica_2006,mir_silicon_2019}).
-Phytoliths act by wearing down the mandibles of insect herbivores, thus reducing feeding efficiency (\cite{waterman_short-term_2021-1,mir_silicon_2019}), as well as reducing the digestive efficiency of the herbivore gut (\cite{hunt_novel_2008}).
+In leaves, phytoliths are typically stouter, though they still increase the mechanical toughness of the leaf (\cite{simpson_still_2017}).
+This overall toughness, and abundance of phytoliths in leaves is thought to have evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}).
+
+Leaf silicon concentrations are negatively correlated with relative growth rate and survival in several invertebrate herbivores (\cite{massey_silica_2006,juma_influence_2015,mir_silicon_2019}).
+Phytoliths act by wearing down the mandibles of insect herbivores, thus reducing feeding efficiency (\cite{mir_silicon_2019,waterman_short-term_2021-1}), as well as reducing the digestive efficiency of the herbivore gut (\cite{hunt_novel_2008}).
 Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}).
@@
 In addition to the anti-herbivore benefits, silicon deposition seems to provide heightened disease resistance and stress tolerance.
-when supplemented with silicon, plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
-Silicon may create apoplastic barriers that seal and toughen the plant tissue, which might explain the range of plant traits that show a positive response to silicon (\cite{coskun_controversies_2019}).
-Under this apoplastic barrier hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
-Empirical evidence shows that silicon is effective at limiting the growth of plant fungal pathogens (\cite{fauteux_silicon_2005}).
+When supplemented with silicon, plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}).
+Empirical evidence also shows that silicon is effective at limiting the growth of some fungal plant pathogens (\cite{fauteux_silicon_2005}).
+These wide-ranging positive effects may be due to silicon-based apoplastic barriers that seal and toughen the plant tissue (\cite{coskun_controversies_2019}).
+Under this apoplastic barrier hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}).
 Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.
@@
 
-Much of today’s plant silicon work is indebted to the pioneering work of \textcite{jones_silica_1967}, who published a comprehensive study of silica in biotic systems, and the subsequent charting of silicon content across the plant kingdom by \textcite{takahashi_possibility_1990}).
-Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world (\cite{coskun_controversies_2019,hartley_ecology_2016,cooke_is_2011,christian_breeding_2022,de_tombeur_silicon_2021}).
-Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}).
+Much of today’s plant silicon work is indebted to the pioneering work of \textcite{jones_silica_1967}, who published a comprehensive study of silica in biotic systems, and the subsequent charting of silicon content across the plant kingdom by \textcite{takahashi_possibility_1990}.
+Epstein’s \citedate{epstein_silicon_1999} seminal paper provided a comprehensive review of the state of knowledge in plant silicon and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world (\cite{cooke_is_2011,hartley_ecology_2016,coskun_controversies_2019,de_tombeur_silicon_2021,christian_breeding_2022}).
+Silicon has been best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}).
 The domesticated grass species rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}).
-Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}).
-The critical depletion of silicon in these production systems has spurred much research activity targeting the biology of silicon, particularly in rice (\cite{deren_variable_1992-1,dai_genetic_2005,rodrigues_silicon_2004,ye_priming_2013}).
+Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to be grown in soils naturally low in silicon, and rapidly deplete the remaining silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}).
+The depletion of silicon in these production systems has spurred much research activity targeting the biology of silicon, particularly in rice (\cite{deren_variable_1992-1,dai_genetic_2005,rodrigues_silicon_2004,ye_priming_2013}).
 In most temperate soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}).
-Thus, silicon research has been slower to develop in dryland temperate cereal crops, (but see \cite{rains_active_2006,neu_silicon_2017,ahmad_silicon_2016}).
+Thus, silicon research has been slower to develop in dryland temperate cereal crops, (but see \cite{rains_active_2006,ahmad_silicon_2016,neu_silicon_2017}).
 Great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils.
-Our ability to integrate silicon as a tool for production improvement in crop production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which crops utilize silicon in pest-protection.
+Our ability to integrate silicon as a tool for improving crop production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which crops utilize silicon in pest-protection.
 
-\printbibliography
+%\printbibliography
 
-\chapter{Chapter 1: Testing rapid silicon accumulation in four cereal crops under real and artificial herbivory}
+\chapter{Testing rapid silicon accumulation in four cereal crops under real and artificial herbivory}
 
@@
 To counter acute damage from herbivores, plants have developed a host of defensive strategies, ranging from changes to the body plan to the development of novel compounds to poison organisms that try to eat the plant (\cite{agrawal_plant_2006}).
-Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation, to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
+Due to the vastly different nature and ontogeny of various defensive strategies in plants, plant defences operate across a range of intensities and time scales, from short-term temporary activation to long-lasting changes in the morphology of the plant (\cite{agrawal_plant_2006, karban_induced_1989}).
 In most scenarios, plants induce defences in response to an external cue, and build in intensity over time, with defensive hormone signals peaking approximately five hours after the initial induction event (\cite{schmelz_quantitative_2003}).
 Despite this rapid hormonal response, actual defensive phenotypes are slower to emerge, due to the higher costs of synthesizing defensive compounds, or developing new tissues like thorns and spines (\cite{karban_induced_1989}).
-Many defensive responses are also context dependent, where the identity of the damaging actor, the severity of damage, and a host of other factors interact to determine the final defensive response (\cite{waterman_simulated_2019}).
+Many defensive responses are also context dependent, where the identity of the damaging actor (e.g.
+species of insect), the severity of damage, and a host of other factors interact to determine the final defensive response (\cite{waterman_simulated_2019}).
 The most effective defensive strategies should be those that can either prevent herbivory outright, or rapidly respond to limit damage.
 These same strategies are also the most promising for crop production, where pest damage represents both an economic and food security cost.
-Integrating better natural plant defences into crop production systems may be key to reducing the environmental impact of agriculture, but hinges upon a thorough understanding of plant defensive physiology.
+Integrating better natural plant defences into crop production systems could reduce some of the environmental impact of agriculture, but hinges upon a thorough understanding of plant defensive physiology.
 
@@
 Soluble silicon taken up from the soil is deposited predominantly in the leaf epidermis, where it forms solid granules that increase the toughness of the tissue, reducing herbivore digestive efficiency (\cite{cooke_is_2011}).
-Plant silicon is expressed latently, but also increases in response to herbivory (\cite{takahashi_possibility_1990, }).
-Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), and recent evidence points to silicon accumulation as being a relatively rapid response, even outpacing the production of phenolic compounds, a group of defensive compounds (\cite{waterman_short-term_2021}).
-This rapid action makes silicon accumulation a promising trait for future crop development.
-Despite the novel results, this pattern has so far been observed in just one species, and only under simulated herbivory through the application of methyl-jasmonate.
-Though a useful tool for herbivory research, methyl-jasmonate application fails to reproduce a complete herbivory signal for the plant, thus observed changes to plant defence may not be representative of a true herbivory scenario (\cite{strauss_direct_2002}).
-Testing for this rapid silicon accumulation across a variety of grain crops, and under both simulated (methyl-jasmonate) and real herbivory is a crucial first step towards integrating rapid silicification into our understanding of plant defence and crop protection.
+Nearly all plants accumulate silicon over their lifespans, but there is also evidence of more acute silicon accumulation in response to herbivory (\cite{takahashi_possibility_1990}).
+
+Multiple studies have demonstrated lasting elevated silicon in response to real and simulated herbivory (\cite{massey_are_2008,hartley_ecology_2016}), with much of the silicon accumulation occurring in the days after the herbivory cue (\cite{ye_priming_2013}).
+Recent evidence points to silicon accumulation as being a relatively rapid response, increasing by up to 30\% within six hours after defensive cues, outpacing increases in phenolic compound-based defences (\cite{waterman_short-term_2021}).
+This rapid action makes herbivory-induced silicon accumulation a promising trait for future crop development.
+Despite the novel results, this pattern has so far been observed in just one species: \textit{Brachypodium distachyion}.
+Testing for this rapid herbivory-induced silicon accumulation across a variety of grain crops, and under both simulated (methyl-jasmonate) and real herbivory is a crucial first step towards integrating rapid silicification into our understanding of plant defence and crop protection.
 
 Plant silicon research has mostly focused on members of the grass family (Poaceae) due to their exceptional silicon content within the plant kingdom, as well as the economic importance of domesticated grass species (\cite{reynolds_silicon_2016}).
-Domesticated crops differ significantly from their wild relatives, due to effects of strong selective pressure imposed by humans (\cite{chen_crop_2015}).
-Most domesticated crops show much lower genetic diversity than their wild ancestors (\cite{hafeez_creation_2021, smith_domestication_2019}).
-Initial selection for a few individuals with favourable traits creates a genetic bottleneck, and the majority of allelic diversity is lost.
-Subsequent selection by humans for agronomically relevant traits can result in concurrent losses of adaptations to natural environments, as the traits that maximize human value (eg.
-yield, ease of harvest) can come at the cost of ecologically relevant traits such as defence (\cite{whitehead_domestication_2017, chen_crop_2015}).
+Much of the research on silicon-herbivore interactions used wild species, yet domesticated crops differ significantly from their wild relatives, due to effects of strong selective pressure imposed by humans (\cite{chen_crop_2015}).
+Most domesticated crops show much lower genetic diversity than their wild ancestors (\cite{smith_domestication_2019,hafeez_creation_2021}).
+Initial selection for a few individuals with favourable traits creates a genetic bottleneck, and most allelic diversity is lost.
+Subsequent selection by humans for agronomically relevant traits can result in concurrent losses of adaptations to natural environments, as the traits that maximize human value (e.g., yield, ease of harvest) can come at the cost of ecologically relevant traits such as defence (\cite{chen_crop_2015,whitehead_domestication_2017}).
 Indeed, in the context of silicon, we can detect clear signals of domestication across the Poaceae family, where wild ancestors consistently have higher baseline silicon content than their domesticated descendants (\cite{simpson_still_2017}).
-Due to the effects of selection on plant defence it becomes crucial to test new developments in the silicon-defence literature in modern crop species, both to validate their utility towards agricultural production, and to gather further observations on the dynamics of silicon-based defences in the first hours after herbivory.
+Novel findings and developments in the silicon-herbivore literature need to be tested with modern crop species, both to validate their utility towards agricultural production, and to gather further observations on the dynamics of silicon-based defences in a broader set of species.
 
-In this study, we tested four globally important cereal crop species for rapid silicon accumulation under artificial and real herbivory.
-In a glasshouse environment, we grew bread wheat (\textit{Triticum aestivum}), oats (\textit{Avena sativa}), barley (\textit{Hordeum vulgare}) and Triticale ($\times$ \textit{Triticosecale}), and tested the following hypotheses:
+In this study, we tested cultivars of four globally important cereal crop species for rapid herbivory-induced silicon accumulation under artificial and real herbivory.
+In a glasshouse environment, we grew bread wheat (Triticum aestivum), oats (Avena sativa), barley (Hordeum vulgare) and Triticale (x Triticosecale), and tested the following hypotheses:
 \begin{enumerate}
@@
         \item Due to disparate taxonomic origins, the tested species would vary in the strength of their silicon accumulation response to herbivory. 
-        \item Due to the different cues involved when comparing true herbivory damage and methyl-jasmonate induced defensive induction, the tested species would show different patterns of short-term silicon accumulation in response to cricket (\textit{Acheta domesticus}) herbivory and methyl-jasmonate application.
+        \item Due to the different cues involved when comparing true herbivory damage and methyl-jasmonate induced defensive induction, the tested species would show different patterns of short-term silicon accumulation in response to cricket (Acheta domesticus) herbivory and methyl-jasmonate application.
 \end{enumerate}
-This study is a thematic replication of Waterman et al.’s \citedate{waterman_short-term_2021} paper, which found that silicon accumulated within six hours after methyl-jasmonate application, and was faster to accumulate than phenolic compounds.
+This study is a thematic replication of Waterman et al.’s 2021 paper, which found \textit{Bracypodium distachyon} accumulated silicon within six hours after methyl-jasmonate application and that silicon was faster to accumulate than phenolic compounds.
 We attempt to extend these findings to commercially important grain crops.
-The findings of this study will refine our understanding of the prevalence of rapid silicification in the Poaceae, and will help to inform the value of potential applications of silicon-based defences into grain crops.
+The findings of this study will refine our understanding of the prevalence of rapid silicification in the Poaceae and will help to inform the value of potential applications of silicon-based defences into grain crops.
 
@@
 
-To test the prevalence of rapid silicon accumulation in Canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley (\ref{Fig:exp_design}).
-We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
+To test the prevalence of rapid silicon accumulation in Canadian cereal crops, we selected three cultivars each of oats, bread wheat, triticale, and barley (Figure \ref{Fig:exp_design}).
+We selected cultivars based on minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations.
 At the start of the experiment, we germinated seeds in germination trays filled with moist sand.
-After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon / kg of soil).
+After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro Sunshine Mix \#4 amended with 0.17 g of solid silicic acid (a rate of 2.4 g silicon / kg of soil).
 Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to reduce silicon limitation to the plants.
@@
 We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or cricket herbivory.
-We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al.
-2021b), while crickets housed in water-pik tubes provided true herbivory (\ref{Fig:exp_design}).
+We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (\cite{waterman_short-term_2021}), while crickets housed in water-pik tubes provided true herbivory (Figure \ref{Fig:exp_design}).
 Prior to introduction to the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial.
@@
 
-We harvested three fully expanded upper leaves from the plants 18 hours after treatment application, and split the leaves in half along the midvein.
-We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2 mL microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
+We harvested three fully expanded upper leaves from the plants 18 hours after treatment application and split the leaves in half along the midvein.
+We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2 mL microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tissuelyser (60 seconds at 30 Hz) in preparation for silicon content analysis.
 We placed the other half of the tissue into a microcentrifuge tube, flash froze it in liquid nitrogen, and subsequently freeze dried it.
@@
 To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}).
-We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13 mm die to create a pellet.
+We pressed leaf powder in a hydraulic press at 300 bar, using a 13 mm die to create a pellet.
 We then placed the pellet in an Olympus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon.
@@
 
-We analysed phenolics as a way to validate the effectiveness of our methyl-jasmonate applications on defence induction.
-To measure the response of phenolics to our treatments, we used the Fast Blue BB method (\cite{pico2020systematic}).
-To prepare our samples, we took 0.075 g of freeze-dried leaf tissue, and ground it to a fine powder in a tissuelyser using three 3.2 mm chrome steel beads at 30 Hz for 60 seconds.
-To the leaf powder, we added 1 mL of 1\% formic acid in 50\% methanol, sonicated for 20 minutes, centrifuged at 15000 G for 10 minutes, then pipetted out the supernatant into a new 2 mL  microcentrifuge tube.
-We then repeated these steps, implementing a double extraction to minimize the amount of phenolics left in the sample.
-To the resulting extract, we added 100 $\mu$L of Fast Blue BB solution, vortexed for one minute, added 100 $\mu$L of 5\% NaOH, vortexed for 10 seconds, and allowed the color to develop in dark conditions for one hour.
-We used a 96-well plate to read the absorbance of 420 nm light.
-We compared the readings to a standard curve created using gallic acid.
+We analysed phenolics to double check the effectiveness of our methyl-jasmonate applications on defence induction and to examine the relationship between phenolics and silicon.
+To measure the response of phenolics to our treatments, we used the Fast Blue BB method (\cite{pico2020systematic}), scaled to use 0.075g of freeze-dried leaf tissue.
 
@@
 Despite the starvation, some crickets did not initiate feeding during the exposure period ($\sim38\%$).
-We filtered out plants assigned to the insect induction treatment that received no damage, to ensure that they would not confound the model (n = 33).
+We filtered out plants assigned to the insect induction treatment that received no damage, to ensure that they would not confound the model ($n = 33$).
 Prior to running our full model, we first tested for an effect of plant biomass on silicon content, as defence levels can be influenced by plant size (\cite{carmona_plant_2011}).
@@
 We implemented these models using the following R model formulae:
-
-\[Silicon \sim Species * Induction + Biomass + (1|Genotype)\]
-\[Phenolics \sim Species * Induction + Biomass + (1|Genotype)\]
-
-We assessed the models using \verb|lmerTest| (\cite{kuznetsova_2017_lmerTest}) with a type III ANOVA.
-
+\[Silicon ~ Species * Induction + Biomass + (1|Cultivar)\]
+\[Phenolics ~ Species * Induction + Biomass + (1|Cultivar)\]
+We assessed the models using lmerTest (\cite{kuznetsova_2017_lmerTest}) with a type III ANOVA. 
 Though not directly related to our hypotheses, we conducted additional post-hoc tests investigating trade-offs between phenolic and silicon content (\cite{simpson_still_2017,waterman_short-term_2021}).
@@
 
-\[Phenolics \sim Silicon + Species + Biomass + (1|Genotype)\]
+\[Phenolics ~ Silicon + Species + Biomass + (1|Cultivar)\]
 
@@
 
-We compiled the final dataset using \verb|DataFrames.jl| (\cite{bogumil_kaminski_2023_7632427}) in Julia 1.8.5 (\cite{bezanson2017julia}).
-We implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}).
-We tested the mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}) and \verb|pbkrtest| (\cite{halekoh_pbkrtest_2014}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}).
-We generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).
+We compiled the final dataset using DataFrames.jl (\cite{bogumil_kaminski_2023_7632427})Kamiński et al.
+2022) in Julia 1.9.0 (\cite{bezanson2017julia}).
+We tested the mixed effects models in R 4.2.2 (\cite{r_core_team_2022}) using lmerTest (\cite{kuznetsova_2017_lmerTest}) and performed a post-hoc Tukey test using emmeans (\cite{lenth_2023_emmeans}).
+We generated graphics in Julia using Plots.jl (\cite{tom_breloff_2023_7736124}).
 
 \section{Results}
-Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}).
-We found strong support for a species effect on silicon content ($p=0.012, F = 6.08, df=3,10.3$).
-Oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}).
-Counter to our predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species.
-Despite a small p-value for an induction treatment effect ($p = 0.011, F = 4.62, df =2, 186$), the only significant pairwise differences between treatments were in wheat, where silicon content was lowest in the insect treatment group (Figure \ref{Fig:induction}).
-We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,186$), likely driven by the stark differences between wheat and the other species' responses to the induction treatments (Figure \ref{Fig:induction}).
-We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction}, Table \ref{Tab:phe_params}).
-Oats had the highest average leaf phenolic content [$31.9 \pm 3.66 mg/kg$], while barley had the lowest average phenolic content [$12.9 \pm 1.82 mg/g$].
+
+Among cultivars, average baseline (control treatment) silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}).
+Species differed in silicon content ($species: p = 0.012, F = 6.08, df = 3,10.3$) with the lowest average silicon content at $0.34 \pm 0.02\% (mean \pm SE)$ in oats, and  the highest average silicon content at $0.76 \pm 0.05\% (mean \pm SE)$ in wheat (Figure \ref{Fig:baseline_si}).
+Silicon content varied by induction treatment ($induction: p = 0.011, F = 4.62, df = 2,186$, Table \ref{Tab:si_params}) with higher silicon content in plants exposed to methyl-jasmonate, while insect and control plants had similar silicon content (Figure \ref{Fig:induction_across_spp}).
+Within species, we did not see induced silicon accumulation for either methyl-jasmonate application nor insect feeding (Figure \ref{Fig:induction_by_spp}), however wheat had significantly lower silicon with cricket feeding, leading to a significant interaction between species and induction ($interaction: p=0.040, F = 2.52, df = 6,186$).
+
+We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction_by_spp}, Table \ref{Tab:phe_params}).
+Oats had the highest average leaf phenolic content $(31.9 \pm 3.66 mg/g) (mean \pm SE)$, while barley had the lowest average phenolic content $(12.9 \pm 1.82 mg/g) (mean \pm SE)$.
 We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).
@@
 
-Recent research in inducible silicon plant defences has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
+Recent research in inducible silicon plant defences has focused on the short-term ($<24$ hours) dynamics of herbivory-triggered silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}).
 The promising results of this work have been highlighted for their potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures.
-In this study, we attempted to demonstrate rapid silicification in four cereal crops.
-We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.
+In this study, we assessed rapid silicification in four cereal crops.
+We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore feeding within any of the species tested.
+Despite our model showing significant effects of both treatment and species on the silicon content, comparing across treatments we see only a slight trend towards increased silicon content, with methyl-jasmonate treated plants displaying the highest silicon content, while control and insect-treated plants had lower silicon content.
+Surprisingly, in wheat the insect treatment resulted in the lowest silicon content.
+Silicon deposition is a permanent process; plants cannot remobilize silicon once it has been deposited.
+It is virtually impossible that cricket feeding on wheat led to a plant response of decreasing silicon.
+However, this pattern may be a signal of silicon-content based feeding preferences, which we discuss more in depth farther below.
 
-This study differed from previous studies demonstrating rapid silicon uptake in a number of ways.
-The two previous studies (\cite{waterman_short-term_2021-1,waterman_short-term_2021}) both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
+This study differed from previous studies demonstrating rapid herbivory-induced silicon uptake in several ways.
+The two previous studies (\cite{waterman_short-term_2021,waterman_short-term_2021-1}) both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability.
 In our study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}.
-In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution.
-The maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that.
-In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-available silicon) (\cite{de_tombeur_plants_2020}), so as to avoid soil conditions with low silicon presence.
+In natural soil environments, most plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution (\cite{de_tombeur_silicon_2021-1}).
+The maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much lower.
+Phytoliths (plant-derived silicates) are a major source of silicon on soils, and for our study, we applied silicic acid to a level in excess of the average availability of phytoliths in global soils (4 g phytoliths/kg of soil) (\cite{de_tombeur_plants_2020}), to avoid soil conditions with low silicon presence.
 Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.
-Furthermore, the observed silicon content of wheat and barley is lower than what is typically found for these species when grown in natural soils (\cite{simpson_still_2017,parr_phytolith_2011}.
+Furthermore, the observed silicon content of wheat and barley is lower than what is typically found for these species when grown in natural soils (\cite{parr_phytolith_2011,simpson_still_2017}).
 The substrate used in this study is ubiquitous in plant science, particularly for plants grown in pots.
-The lack of silicon availability in this substrate, even with silicon supplementation is notable, and deserves serious thought.
+It is notable that our plants failed to reach silicon contents commonly observed in field trials, particularly if this is a result of silicon limitation from the substrate (however, we did not measure the silicon or nutrient content of the soil substrate).
+Further testing on our soil substrate would help to reveal how silicon soil availability may have influenced our observations.
 Silicon plays a role in a myriad of plant physiological responses (\cite{coskun_controversies_2019}), often mediating/reducing the negative effects of stressors.
-Silicon limitation is assumed to be infrequent in global soils (\cite{guntzer_benefits_2012}), yet our data suggest that plants grown in silicon fail to reach typical levels of silicon content.
-We can only speculate how silicon deficiencies may skew observations in potted plant experiments, but we think it would be prudent for researchers to consider adding soluble silicon to their nutrient mixes to emulate realistic silicon availabilities.
-Nearly all cereal production worldwide takes place in soils and we need further work on the temporal aspects of silicon uptake in a range of soil conditions to broaden our understanding of rapid silicon uptake.
-Though rapid silicification retains promise for improved crop protection, our work highlights the need for future research and replication to fully characterize the nature of silicon induction.
+Silicon limitation is assumed to be infrequent in global agricultural soils (\cite{guntzer_benefits_2012,haynes_contemporary_2014}), yet our data suggest that plants grown in potting mix fail to reach typical levels of silicon content.
+We can only speculate how silicon deficiencies may skew observations in potted plant experiments, but we think it would be prudent for researchers to explicitly measure the soluble silicon content of the growth media when conducting experiments focussed on physiological traits that have known responses to silicon supplementation.
+In agricultural settings, silicon fertilization is often achieved using silicate slag application (\cite{meena_case_2014}), though liquid silicon fertilizers are also coming onto the market (e.g.
+Orion Future Technologies).
+We need further work on the temporal aspects of silicon uptake in a range of soil conditions to broaden our understanding of rapid silicon uptake.
+Though rapid silicification still holds promise for improved crop protection, our work highlights the need for future research and replication to fully characterize the nature of silicon induction.
 
-We did not see a clear relationship between silicon and phenolic content in our plants, matching the observations of \textcite{waterman_short-term_2021}).
-Though they observed a negative relationship between silicon and phenolics in plants grown without supplemental silicon, this relationship disappeared when they added silicon to the growth media.
-As previously discussed, our study seems to have imparted moderated silicon limitation on the plants, and this may explain the hint of a negative trend observable in the scatterplot (Figure \ref{Fig:phe_si_scatter}).
-Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}.
-In their study, they tested 8 species groups of wild and domesticated cereal crops, including wheat and barley.
-Our results for wheat and barley silicon and phenolic content more or less follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content.
+We did not see a clear relationship between silicon and phenolic content in our plants, suggesting that in the conditions of our study, there were no defensive trade-offs between silicon and phenolics.
+\textcite{waterman_short-term_2021} observed a negative relationship between silicon and phenolics in plants grown without supplemental silicon, but this relationship disappeared when they added silicon to the growth media.
+In their study, they posited that silicon supplementation limits the conversion of pre-cursor compounds into phenolics, limiting phenolic accumulations under herbivory cues.
+This process could help to explain our lack of phenolic induction in response to our induction treatments.
+However, we only have one silicon treatment (supplemented), so are unable to know how phenolic content would have responded to our treatments in the absence of silicon.
+
+We observed marked variation in the silicon and phenolic content across species.
+Some of is likely simply due to different evolutionary and domestication histories selecting from different traits.
+In general, our observations on silicon and phenolic content variation among species complement those of \textcite{simpson_still_2017} .
+Our results for wheat and barley silicon and phenolics generally follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content.
 \textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species.
-Both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we can interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic content among studies and growing environments.
+Both silicon content and leaf phenolic content are dependent on local environmental conditions (\cite{schaller_silicon_2012,quigley_soil_2020}), but comparing relative differences across studies is still useful.
 In our study, oats had the highest phenolic concentration, and the lowest silicon content.
-The relative difference in the phenolic and silicon content of oats and wheat in our study was larger than largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study.
-Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (eg.
-domestication) and ecological (eg.
-growing environment) drivers that also determine the final defensive phenotype of cereal species.
-Understanding how these various factors constrain the range of defensive expression for these species will help bring focus to efforts to translate silicon ecology into the field of agriculture.
+The relative difference in the phenolic and silicon content of oats and wheat in our study was similar to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson et al.’s study.
+Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (e.g., domestication) and ecological (e.g., growing environment) drivers that also determine the final defensive phenotype of cereal species.
+Understanding how these various factors interact to determine defensive expression for these species will help bring focus to efforts to translate silicon ecology into the field of agriculture.
 
-This study was the first that we are aware of to use crickets as a treatment to induce silicon uptake.
-Other studies have used crickets to measure herbivore performance in response to other defensive phenotype manipulations (eg.
-eCO$_2$, caterpillar herbivory) (\cite{biru_contrasting_2022, ryalls_impacts_2017}).
-Plant defensive responses to herbivory are linked to the identity of the herbivore (\cite{afkhami_endophyte-mediated_2009, amsberry_effects_2006}), and it could be that house crickets elicit different silicon responses than more commonly used caterpillar herbivores.
-Other studies have observed silicon increases in response to locust (Orthoptera) damage (\cite{massey_herbivore_2007, garbuzov_interactive_2011}), so it is unlikely that crickets do not induce silicon uptake, but it may be that the temporal dynamics of such a response differ from caterpillar-triggered defensive induction.
-approximately one third of the applied crickets do not initiate feeding within the time frame of the study.
-In \textcite{massey_herbivore_2007}, crickets tended to avoid feeding on plants with higher silicon content.
+We did not observe an inducible effect of cricket feeding on plant silicon content.
+Other studies have used crickets to measure herbivore performance in response to other defensive phenotype manipulations (e.g., $eCO_2$, caterpillar herbivory) (\cite{ryalls_impacts_2017,biru_contrasting_2022}), however ours is the first study that we are aware of that used crickets as a treatment to induce silicon uptake.
+Plant defensive responses to herbivory are linked to the identity of the herbivore (\cite{amsberry_effects_2006,afkhami_endophyte-mediated_2009,carrillo2012induction,carrillo_loss_2014}), and it could be that house crickets elicit different silicon responses than other herbivores.
+Other studies have observed silicon increases in response to locust (Orthoptera) damage (\cite{massey_herbivore_2007,garbuzov_interactive_2011}),, so it is unlikely that crickets do not induce silicon uptake, but it may be that the temporal dynamics of such a response differ from caterpillar-triggered defensive induction.
+
+We found that approximately one third of the applied crickets did not initiate feeding within the time frame of the study.
+In \textcite{ryalls_impacts_2017}, crickets fed less on plants with higher silicon content.
 This same pattern may have been at play in our own study system, where plants that had relatively high silicon content escaped damage, while the lower silicon plants experienced herbivory.
@@
 This exclusion may in part explain our observations of the low silicon content in the wheat-insect treatment group.
-Unfortunately, our study lacks the replication (and thus statistical power) to confirm if there were significant differences between the silicon content in the damaged and undamaged plants within the insect treatments.
+Unfortunately, as our study was not designed to test whether initial plant silicon content determined feeding likelihood, we did not have enough statistical power to determine if there were significant differences between the silicon content in the damaged and undamaged plants within the insect treatments across species.
 A future study that quantified the effect of pre-herbivore-exposure silicon content on the likelihood or extent of herbivory would help to answer this question.
 
-Though we did not find support for rapid silicon uptake in response to herbivory, we have added to our knowledge of silicon content in cereal species, and have identified crucial gaps in our understanding of the dynamics of silicon uptake.
+Though we did not find support for rapid silicon accumulation in response to herbivory, we have added to our knowledge of silicon content in cereal species and have identified crucial gaps in our understanding of the dynamics of silicon uptake.
 It is unclear how silicon uptake rates are modified by the nature of the growth substrate, and there are almost certainly important questions to be answered in comparing nutrient-solution-based and soil-based silicon uptake.
-Our use of potting mix seemed to limit the silicon availability for our plants.
+Our use of potting mix seemed to limit the silicon availability for our plants, as our wheat and barley had much lower silicon contents than those of \textcite{simpson_still_2017}.
 Imposing silicon limitation on study plants through the use of potting mix may skew results for a wide variety of plant studies, as silicon is involved in all manner of stress and damage tolerance responses.
 More explicit consideration of how silicon limitation may influence the results of a given study, and integrating silicon into fertilizer mixes for plant science may help increase the translatability of potted plant experiments to real soil environments.
-The future of silicon-based crop technologies remains bright, but there are still important obstacles that must be overcome as we try to translate silicon ecology research into production methods.
+The future of silicon-based crop technologies remains bright, but there are still important obstacles that must be overcome as we try to translate ecological research on silicon into production methods.
 
@@
 
-\section{Data Availability}
+%\section{Data Availability}
 
-\printbibliography
+%\printbibliography
+\clearpage
 
@@
         \centering
-        \caption{ANOVA table for our linear mixed effects model analyzing the effect of defence induction and species identity on leaf phenolic content.
+        \caption{ANOVA table for our linear mixed effects model analysing the effect of defence induction and species identity on leaf phenolic content.
 We generated the ANOVA table using the R package lmerTest, specifying a type III ANOVA.
@@
                 Biomass          &     1, 151.907 & 0.0760 & 0.783174 \\
-                Induction $times$ Species &     6, 180.798 & 2.7535 & 0.013850 \\
+                Induction $\times$ Species &     6, 180.798 & 2.7535 & 0.013850 \\
                 \hline
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
         \centering
-        \begin{subfigure}[b]{0.49\textwidth}
+        \begin{subfigure}[b]{0.65\textwidth}
                 \centering
@@
         \end{subfigure}
-        \begin{subfigure}[b]{0.49\textwidth}
+
+        \begin{subfigure}[b]{0.65\textwidth}
                 \centering
@@
         \centering
-        \caption{Baseline (uninduced) silicon content (\% dry matter) and phenolic content (mg/kg) in the cereal cultivars used in this study.
+        \caption{Mean baseline (uninduced) silicon content (\% dry matter) and phenolic content (mg/kg) in the cereal cultivars used in this study.
 Cultivar species is indicated by point color.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
+        \includegraphics[width = \textwidth]{images/induction_only_bytrt_annotated.png}
         \centering
-        \begin{subfigure}[b]{0.49\textwidth}
+        \caption{Average silicon content of cereal crops in response to three treatment types.
+We treated plants either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
+We sampled leaves 18 hours after treatment and analysed them for silicon using XRF.
+Letters indicate significant differences with a threshold of $\alpha = 0.05$.}
+        \label{Fig:induction_across_spp}
+\end{figure}
+
+\begin{figure}[ht]
+        \centering
+        \begin{subfigure}[b]{0.7\textwidth}
                 \centering
@@
         \end{subfigure}
-        \begin{subfigure}[b]{0.49\textwidth}
+
+        \begin{subfigure}[b]{0.70\textwidth}
                 \centering
@@
 We treated plants either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}).
-Leaves were sampled 18 hours after treatment, and were analysed for silicon using XRF or for phenolic content using the Fast Blue BB method.
+We sampled leaves 18 hours after treatment, and analysed them for silicon using XRF or for phenolic content using the Fast Blue BB method.
 Dots and error bars are the mean $\pm$ 1 Standard Error.}
-        \label{Fig:induction}
+        \label{Fig:induction_by_spp}
 \end{figure}
 
-\begin{figure}[h]
+\begin{figure}[ht]
         \includegraphics[width = \textwidth]{images/phenolic_silicon_regression.png}
@@
 
-\chapter{Chapter 2: Genetic drivers of silicon accumulation in a wild ancestor of wheat}
+\chapter{Genetic drivers of silicon accumulation in a wild ancestor of wheat}
 
@@
 Over the past thirty years, plant silicon has emerged as an exciting possible tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}).
-Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}).
-Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content ($\sim0.6 - 10\%$ total dry weight) (\cite{reynolds_silicon_2016}).
-Silicon's high abundance in many soils, high expression in cereal crops, and incredible broad spectrum effects on plant vigor and stress tolerance have made it a tantalizing target for improvements in agricultural yield and sustainability.
+Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{rudel_agricultural_2009,awika_major_2011,faostat}).
+Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (~0.6 - 10\% total dry weight) (\cite{reynolds_silicon_2016}).
+Silicon's high abundance in many soils, high concentration in cereal crops, and incredible broad-spectrum effects on plant vigour and stress tolerance have made it a tantalizing target for improvements in agricultural yield and sustainability (\cite{christian_breeding_2022}).
 
-Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress.
+Silicon accumulation in plants underpins a variety of physiological and developmental strategies that plants use to cope with stress.
 Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.
-For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005, katz_silicon_2019}).
-On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}).
+For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005,katz_silicon_2019}).
+On the abiotic side, silicon supplementation improves plant tolerance of soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}).
 In comparing stressed plants grown in the absence or presence of silicon, silicon-exposed plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}).
-A current hypothesis explaining the broad-spectrum activity of silicon is presented in  \textcite{coskun_controversies_2019}.
-Their apoplastic barrier hypothesis suggests that silicon deposited in the apoplast of plant tissues modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance.
+\textcite{coskun_controversies_2019} hypothesized that silicon deposited in the apoplast of plant tissues modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance.
 Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts.
@@
 
-Plants gather silicon from the soil solution, using a suite of transporter proteins to pump it into their vascular systems and then transport it throughout the body (\cite{reynolds_silicon_2016}).
+Plants uptake silicon from the soil solution, using a suite of transporter proteins to pump it into their vascular systems and then transport it throughout the body (\cite{reynolds_silicon_2016}).
 Variation in the relative expression of these transporters, as well as differences in the development of the end points for silicon deposition (silica cells), may drive phenotypic variation among individuals.
 Additionally, individuals may vary in their ability to scavenge silicon from the soil.
-The soluble form of silicon, silicic acid (SiOH4) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM (\cite{epstein_anomaly_1994}).
-Soluble silicon in the soil is derived primarily from the weathering of silicate minerals, and secondarily from the remobilisation of silicon in decaying plant material (\cite{de_tombeur_silicon_2021-1}).
+The soluble form of silicon, silicic acid (\ch{H4SiO4}) has a maximum solubility in water of around 2 mM, though typical soil concentrations range from 0.1 mM to 0.6 mM (\cite{epstein_anomaly_1994}).
+Plants derive soluble silicon in the soil primarily from the weathering of silicate minerals, but also access silicon that dissolves out of decaying plant material (\cite{de_tombeur_silicon_2021-1}).
 Weathering of silicates releases a host of plant nutrients including aluminium, silicon, iron, and phosphorus (\cite{de_tombeur_silicon_2021-1}).
-Soil biota can drive weathering, using organic acids and other molecules to complex metal ions off of soil aggregates, making them available for uptake by organisms (\cite{de_tombeur_silicon_2021}).
-Plant roots can release carboxylates and phytosiderophores to weather phosphorus and silicon out of soil minerals (\cite{de_tombeur_silicon_2021-1}).
-Along with Si and P mobilisation, manganese is often released, and taken up by plants roots.
-Previous research has used leaf manganese content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), yet so far we are unaware of any studies looking for quantitative variation among genotypes of leaf manganese.
-If we could identify regions of the plant genome associated with variation in leaf manganese content, we could follow up with more targetted studies looking for genes that influence root carboxylate release.
-This trait could eventually be incorporated into breeding programs, improving nutrient use efficiency, ultimately easing our dependence on external inputs to agricultural fields.
 
-The use of x-ray fluorescence (XRF) to quantify plant silicon has greatly reduced the costs, danger, and processing time of for studies focussing on this topic (\cite{reidinger_rapid_2012}).
-XRF works by using low-power x-rays to excite elements in the sample, and measures the resulting emitted light.
+Harnessing root exudation in agricultural systems could help reduce issues of over-fertilization and may help to improve carbon sequestration in agricultural soils (\cite{cornelis_soil_2022,panchal_soil_2022}).
+Plant roots can release carboxylates and phytosiderophores to weather phosphorus and silicon out of soil minerals (\cite{de_tombeur_silicon_2021-1}), reducing the need for additional fertilizers.
+Soil biota, which are highly affected by root exudates, can drive weathering, using organic acids and other molecules to complex metal ions off soil aggregates, making them available for uptake by organisms (\cite{de_tombeur_silicon_2021}).
+Carboxylates (and phytosiderophores) react with aluminium and iron oxides, to which phosphorus and silicon are bound (\cite{lambers_leaf_2015}).
+The carboxylates take the place of these phosphates and silicates, chelating iron oxides and liberating the phosphates and silicates for uptake.
+In this process, the carboxylates also complex with other metal ions such as zinc, manganese, and copper, increasing their availability to the plant roots.
+
+Previous research has used leaf manganese content to proxy for the carboxylate releasing activity of plants (\cite{lambers_leaf_2015}), and there is considerable potential in harnessing root exudates to improve agricultural sustainability (\cite{cornelis_soil_2022}).
+Indeed, \textcite{lambers_leaf_2015} proposed using variation in leaf manganese content to screen crops for genotypes that have high root exudate activity, with a specific focus on improving phosphorus acquisition.
+However, carboxylates also liberate silicon, increasing its plant availability in the soil, so screening crop plants to find high carboxylate-releasing genotypes, could also bring benefits to breeding efforts targeting silicon acquisition.
+To date, this manganese screening approach has been implemented in chickpea (\cite{pang_carboxylate-releasing_2018,wen2021addition}), and in this study we explore genotypic variation in leaf manganese content in a wild ancestor of wheat.
+Identifying genes associated with variation in leaf manganese content could provide a starting point for further investigations into the genetic controls of carboxylate release and may ultimately allow for crop breeding targeting root exudation activity.
+
+Crop improvement has increasingly leveraged the power of diversity panels to explore genetic variation and identify new genes for integration into commercial germplasm.
+Diversity panels can be assembled of extant domesticated varieties (\cite{wingen_establishing_2014}) or of wild ancestors of current domesticated crop species (\cite{gaurav_population_2021}).
+To identify new genes, researchers must phenotype and sequence large numbers of individuals, which can be a costly process.
+However once sequence data are generated, they may be reused with other phenotypic data.
+The use of x-ray fluorescence (XRF) to quantify plant silicon has greatly reduced the costs, danger, and processing time for studies focusing on the chemical composition of plants (\cite{reidinger_rapid_2012}).
 One of the most exciting features of XRF is the fact that it can analyse multiple elements at once, allowing for broad characterisation of the sample for most elements heavier than aluminum.
-Though XRF is an established technique to measure plant silicon, its may also be used to measure other metals of interest, including manganese.
-In this study we use XRF to quantify variation in silicon and manganese content among a diversity panel of a wild ancestor of bread wheat, \textit{Aegilops tauschii}.
-This panel has publicly available sequence data \cite{gaurav_population_2022}, allowing us to perform a genome-wide association study to link silicon and manganese variation to genotypic variation, laying the groundwork for future, more targetted explorations of the genome.
-Identifying genetic controls over these traits will hopefully help develop breeding targets to improve plant performance and safeguard yeilds against a destabilizing climate.
+Though XRF is an established technique to measure plant silicon, it may also be used to measure other metals of interest, including manganese, decreasing the costs associated with phenotyping.
+
+In this study we use XRF to quantify variation in silicon and manganese content among a diversity panel of a wild ancestor of bread wheat, Aegilops tauschii.
+We pair these phenotypic data with publicly available sequence data (\cite{gaurav_population_2021}) and use a genome-wide association (GWAS) to try to link silicon and manganese variation with genotypic variation.
+This study is exploratory in nature, and lays the groundwork for future, more targeted explorations of the genome.
+Identifying genetic controls over these traits will hopefully help develop breeding targets to improve plant performance and safeguard yields against a destabilizing climate.
 
@@
 
-For this experiment, we used the L2 panel of Aegilops tauschii from \textcite{gaurav_population_2022}.
-This panel is a 151 genotype subset of the full \textit{Aegilops tauschii} panel, selected for high diversity in nucleotide-binding leucine rich repeats receptor (NLR) genes.
-We grew the L2 panel at three different sites; two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalised seedlings in growth chambers prior to transplanting into pots.
-For full site details see Supplementary Table S1.
-Using 151 accessions, we started trays of seedlings in glasshouse or growth chamber environments.
-At approximately eight weeks after germination, seedlings were transplanted to their field sites.
+For this experiment, we used the L2 panel of Aegilops tauschii from \textcite{gaurav_population_2021}.
+This panel is a 151-genotype subset of the full Aegilops tauschii panel, selected for high diversity in nucleotide-binding leucine rich repeats receptor (NLR) genes.
+In order to detect both genotype and genotype x environment effects, and as is common in genetic studies, we utilized multiple common gardens for our experiment.
+We grew the L2 panel in two field sites on the University of British Columbia campus, with planting occurring in the fall, and in a glasshouse, with planting occurring in the spring.
+Though geographically close, the two outdoor sites differed noticeably.
+One site, Totem Field is managed as an agricultural research station, with regular cover-cropping, irrigation, and fertilization.
+The other outdoor site is at the south campus research area, and is a raised berm on a roadside, that was built up approximately 30 years ago.
+This was essentially a waste area, and the only management it received prior to the start of the study was intermittent mowing.
+The soil had significant grit and gravel fractions compared to the much finer soils at the agricultural site, and we assumed that the differences in management history would present different soil chemistries and biologies for the plants.
+In contrast, we used potting soil in the glasshouse environment.
+In selecting these three environments (Totem Field, South Campus, Glasshouse) we hoped to present the plants with three distinct growing conditions.
+Under this design, we can discriminate between genotypic- and genotype by environment interactive-effects on the phenotype.
+This can reduce the type I error rate and improves the applicability of identified genetic markers in deployment to crop breeding.
+
+Prior to experimental planting, we started trays of seedlings in glasshouse (for outdoor transplants) or growth chamber (for glasshouse transplants) environments.
 For each environment, we started four replicates of each accession.
-We planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements.
-Each outdoor block was a 16 m\textsuperscript{2} square, with plants arranged $\sim$35 cm apart.
-Shortly after transplanting to the field sites, we applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets.
-Field transplantation took place on the 15th of October 2021 and the 16th of December 2021.
-For the glasshouse environment, we started seedlings in growth chambers in January 2022.
-After 12 weeks, we moved the seedlings to vernalisation chambers (4ºC, 8:16h light:dark) for eight weeks.
-We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with 0.17g of silicic acid (Tixosil 68B, Solvay).
-Pots were arranged using the same randomized block design, adapted to fit on two flood tables.
-To ensure a comparable life stage across environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.
+After approximately eight weeks, we began vernalization of the plants.
+Aegilops tauschii requires a vernalization period to reach maturity.
+For the outdoor gardens, vernalization occurred naturally over the winter season, and thus we transplanted them to their final garden locations.
+We transplanted into Totem Field on the 15th of October 2021, and into the South Campus garden on the 16th of December 2021.
+In each garden, we planted according to a randomized block design, to minimize the effects of soil heterogeneity and plant location on our phenotype measurements.
+Each garden consisted of four blocks, each containing a single replicate of the genotypes.
+Each outdoor block was a $16 m^2$  square, with plants spaced on a 35 cm grid.
+Shortly after transplanting to the field sites, we applied a dose of liquid fertilizer to improve transplant survival, as well as broadcast slow-release fertilizer pellets to the blocks.
+
+For the glasshouse environment, we used growth chambers for vernalization, keeping plants at 4\degree C for eight weeks under a 8:16h light:dark cycle.
+We then transplanted them into 10cm pots filled with SunGro Sunshine Mix \#4 amended with 0.17g of solid silicic acid.
+Approximately eight weeks after germination, we transplanted seedlings to their garden locations.
+Pots were arranged using the same randomized block design as the outdoor gardens, adapted to fit on two flood tables.
+To ensure a comparable life stage across environments at time of harvest, these plants grew for three months (mid June--mid September 2022), until they had mature flower heads.
 
 \subsection{Plant harvest and sample preparation}
+
+When the plants reached maturity, we harvested the entire above-ground portion of each plant.
+For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022.
+We placed harvested material in labelled paper bags and dried it in a 60\degree C drying oven for 48 hours.
+To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers.
+Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation.
+We picked leaves until we collected 200 mg of dry leaf material.
+Some plants did not yield enough leaf tissue to meet the 200 mg threshold and thus we decided to pool leaf material by genotype within sites, excluding genotypes with fewer than three individuals at each site.
+For each of the remaining genotypes, we grouped samples by environment, took a 100 mg subsample of the harvested leaf material and combined these subsamples into a new sample.
+Overall, we were left with 359 samples across three sites, making approximately 115 usable genotypes per site.
+We packed dried leaf samples in 2 mL microcentrifuge with three 3.2 mm chrome steel grinding pellets and ground them in a tissuelyser ball mill for 60 seconds at 30 Hz.
+We sealed and stored the resulting leaf powder at room temperatures until XRF analysis.
 
@@
 In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand.
-For beam 1 (manganese), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time.
+For beam 1 (manganese), we used a 20 second read time, while for beam 2 (silicon), we used a 45 second read time.
 Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy.
@@
 We calibrated our measurements against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).
+We were unable to complete a full calibration for manganese in the scope of this thesis so results using the manganese data are preliminary.
+A full calibration of the manganese data will be completed as part of the publication process of this chapter.
 
@@
 
-To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{Fig:rank_plots}).
+To compare the average silicon and manganese content of our plants between the different gardens, we calculated the mean and standard error for each element at each garden and plotted.
+To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted.
 To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content.
@@
 We observed right-skew for both elements and applied a log-transformation.
-To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$).
-We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) packages in R 4.2.2 (\cite{r_core_team_2022}) to regress leaf manganese content on leaf silicon content.
+To control for inter-site-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X-\mu}{\sigma}$).
+We then used the lmerTest and MuMIn packages in R 4.2.2 to regress leaf manganese content on leaf silicon content (\cite{barton_2023_mumin,kuznetsova_2017_lmerTest,r_core_team_2022}.
 We used a random effect of Genotype to control for genotype-specific differences in silicon content.
-We used the R package \verb|MuMIn| (\cite{barton_2023_mumin} to estimate the $R^2$ value of our mixed model.
+We used the R package MuMIn to estimate the R2 value of our mixed model.
 
-To perform the GWA analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}.
-For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}.
+To perform the GWA analysis, we followed the methodology and code published in \textcite{gaurav_population_2021}.
+For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2021}.
 For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript.
-As per \textcite{gaurav_population_2022}, to reduce the computational intensity of our analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative.
-For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers.
-We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.
-For peaks that were significant in at least two environments, we cross-referenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.
+To reduce the computational intensity of our analysis, we prefiltered the total k-mer matrix to remove k-mers with a low chance of being informative (e.g., found in just a few accessions).
+For each environment, we ran the GWA, filtered k-mers with an association score of $<6$, and plotted the remaining k-mers.
+We calculated our Bonferroni correction threshold using the script provided by \textcite{gaurav_population_2021}.
+For marker-trait associations that were significant in at least two environments, we cross-referenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.
 
@@
 Silicon content in \textit{Aegilops tauschii} ranged from 0.784\% to 11.473\%.
-There were notable differences in silicon content based on the growing environment.
-The glasshouse plants averaged $1.450\% \pm 0.032 (SE)$, while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132.
-Manganese content varied from 36.5 ppm to 1296.5 ppm, and averaged $225.1 \pm 25.45$ ppm (SE) in the glasshouse environment, and $ 71.05 \pm 2.571 ppm (SE)$ and $ 224.0 \pm 11.01 (SE)$ for the two outdoor growing environments.
-The results summarized in our rank plots (Figure \ref{Fig:rank_plots}) show that there are weak g.
-Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}).
-Overall, our analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}).
-For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance.
+There were notable differences in silicon content based on the growing environment (Figure \ref{Fig:environment_elements}).
+The glasshouse plants averaged $1.450\% \pm 0.032$ (SE), while Totem Field averaged $4.935\% \pm 0.108$ and the South Campus site averaged $6.471\% \pm 0.132$.
+Manganese content varied from 36.5 ppm to 1296.5 ppm and averaged $225.1 \pm 25.45$ ppm (SE) in the glasshouse environment, while Totem Field averaged $71.05 \pm 2.571$ ppm (SE) and the South Campus site averaged $224.0 \pm 11.01$ (SE) (Figure \ref{Fig:environment_elements}).
+Our rank plots display the variation in rank silicon and manganese content averaged across the three sites (Figure \ref{Fig:rank_plots}).
+Accession \textit{BW\_01107} had the lowest rank for silicon, while \textit{BW\_01004} had the highest rank.
+Accession \textit{BW\_01138} had the lowest rank for manganese, while \textit{BW\_01091} had the highest rank.
+Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051$, $z = 5.46$, $p < 0.0001$, $R^2_{\text{marginal}} = 0.05$, $R^2_{\text{conditional}} = 0.35$) (Figure \ref{Fig:mn_si_regression}).
+Overall, our analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across all environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}).
+For manganese, we detected several significant marker-trait associations in each environment that exceeded the threshold for significance.
 However, when comparing across environments, we found no regions that consistently exceeded the threshold significance.
-We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil, though absent in the glasshouse environment (Table \ref{tab:peaks_and_genes}).
+We did, however, find 11 regions exceeding the significance threshold that were present in both of our outdoor growing environments, though absent in the glasshouse environment (Table \ref{Tab:peaks_and_genes}).
 Of these genes, three had associated genes in NCBI.
-Two of the genes were uncharacterised non-coding RNA genes.
-The third gene codes for CASP-like protein 5B3.
+Two of the genes were uncharacteristic non-coding RNA genes.
+The third gene codes for CASP-like protein 5B3 and has a negative association with manganese content.
 
@@
 
-We observed major differences in silicon and manganese content between our growth environments, and did not find consistent correlations between genetic variation and variation in silicon content.
-Our rank plots revealed that relative content of both silicon and manganese is not stable across environments.
-This suggests strong genotype by environment interactions, which likely reduced our ability to detect genetic contributions to the silicon and manganese phenotype.
-Interestingly, we detected a number of peaks for leaf manganese content, though many of these were only found in a single environment.
-Though we detected no peaks present in all three growing environments, we detected 11 regions that were significant in both outdoor environments.
-Compared to the glasshouse environment, the outdoor environments likely exposed the plants to a soil environment in which root exudation yeilds nutrients for uptake.
-Root exudation of organic acids and chelating agents liberates phosphates and silicates from inorganic pools (\cite{de_tombeur_silicon_2021-1,lambers_plant_2008}).
-In the glasshouse environment, we grew plants in standard potting mix, which is high in organic matter with a relatively low inorganic component.
-In this soil context, root exudation is unlikely to yield appreciable nutrient liberation.
-The high manganese content found in the glasshouse plants is most likely due to the watering regime they experienced.
-We watered these plants using municipal tap water.
-In the Vancouver region, this water is sourced from reservoirs in the surrounding mountains, which themselves are filled from ground runoff.
-Nearby groundwater wells have issues with high manganese content (\cite{hu_drinking_2020}).
-A constant supply of soluble manganese in the irrigation the glasshouse plants likely accounts for the elevated manganese content.
+We observed major differences in plant silicon and manganese content between our growth environments (Figure \ref{Fig:environment_elements}).
+This suggests strong environmental influence over plant silicon and manganese; the effects of soil nutrient availability on uptake are likely significantly stronger than the effect of any given gene of a plant’s ability to uptake the nutrients.
+Our findings of a strong environment component to silicon phenotype are similar to those of \textcite{deren2001plant} who found that rice and sugarcane genotypes show considerable variation in their silicon content when comparing on different soils.
+However, unlike Deren’s findings that rank content was stable across environments, we saw significant variation in the relative content of silicon and manganese between our sites.
+
+Our ranks are not perfectly stable across environments, with some genotypes having a rank standard deviation that spans almost the entire range of ranks, while other genotypes have much smaller rank error (Figure \ref{Fig:rank_plots}).
+This seems to suggest strong genotype by environment interactions.
+However, if there were strong environmental interactions with a single or a few markers, we would expect to observe strong peaks in the Manhattan plots in some, but not all, environments.
+Our inability to detect major marker-trait associations for silicon may be driven by polygenic control over silicon uptake and deposition.
+The independent influence of multiple genes on a single trait would overwhelm the statistical power of this study, creating variation in silicon content but preventing us from detecting significant marker-trait associations.
+In a study on the genetic controls over silicon content in rice, \textcite{dai_genetic_2005} were more stringent than us when selecting tissue to phenotype, targeting just single uppermost leaf.
+In our study, we pooled the upper three leaves of each stem.
+It could be that different genes are responsible for the silicon content of the uppermost leaf and other leaves.
+In other traits in wheat, distinct genes control the phenotype of the ultimate and penultimate leaves (\cite{isidro2012quantitative}).
+If this is true for silicon, this would introduce noise into our data and limit our ability to detect significant marker-trait associations.
+which likely reduced our ability to detect genetic contributions to the silicon phenotype.
+
+We detected many significant marker-trait associations for leaf manganese content, though many of these were only found in a single environment.
+Though we detected no marker-trait associations present in all three growing environments, we detected 11 regions that had significant associations in both outdoor environments.
+Unfortunately, only one marker-trait association is in a coding-region of the genome.
+LOC109751197 codes for a CASP-like protein.
+CASP-like proteins are involved in the modification of the cell membrane and intracellular space.
+CASP-like proteins are involved in cell membrane modification and can form transmembrane scaffolds (\cite{roppolo_functional_2014}).
+This leads to a puzzling interpretation, where mutations in a gene involved membrane modification (likely decreasing effectiveness and thus increasing permeability) decrease manganese uptake.
+The uncharacterised nature of the gene in question makes more in depth interpretation of the results difficult but provides a target for future investigations into genomic controls over manganese uptake.
+
+Finding significant marker-trait associations for manganese that were present in both outdoor gardens but absent from the indoor garden suggests the presence of a strong genotype by environment interaction.
+This interaction may arise from differences in the growth media when comparing the glasshouse and outdoor plants.
+The glasshouse plants were grown in SunGro potting mix, which is mostly made up of peat moss with a small amount of dolomitic lime.
+In comparison, the outdoor gardens had soils with much higher fractions of inorganic content.
+In this soil context, with their mineral soils, root exudation of carboxylates is more likely to have an effect silicon and manganese availability.
+Root exudation of organic acids and chelating agents target inorganic pools of nutrients and other elements (\cite{lambers_plant_2008,de_tombeur_silicon_2021-1}).
+In the potting mix of the glasshouse gardens, the small inorganic fraction would provide limited reactions sites for root exudates, yielding low levels of nutrient liberation.
+The high manganese content found in the glasshouse plants could be due to the watering regime they experienced.
+Though we did not test the tap water directly, nearby municipalities struggle with high manganese concentrations in the water supply (\cite{hu_drinking_2020}).
+If this is true for UBC’s water source, a constant supply of soluble manganese in the irrigation to the glasshouse plants would account for the elevated manganese content.
+Further chemical analysis of the water and soils would help to clarify the substrates available for root exudates to interact with, as well as resolve the drivers of our patterns of manganese concentrations.
+Though not included in this thesis, these data will be added prior to the publication of this chapter.
 
 Our observations of a correlation between silicon and manganese content could be evidence of root-exudate driven silicon uptake.
-It is possible that this positive correlation is due to a nutrient effect more generally, where manganese and silicon covary in availability in the soil.
-However, further regressions show that silicon only shows positive correlations with iron and manganese, both of which are mobilized by root exudation (Supplementary Fig [Number]) (\cite{de_tombeur_silicon_2021-1}).
-Unfortunately, only one MTA is located in a coding-region of the genome.
-LOC109751197 codes for a CASP-like protein, involved in the casparian strip.
-The presence of a SNP in this region is negatively correlated with manganese content.
-CASP-like proteins are involved in cell membrane modification, and can form transmembrane scaffolds (\cite{roppolo_functional_2014}).
-This leads to a puzzling interpretation, where defects in a gene involved membrane modification (likely reducing permeability) decrease manganese uptake.
-The uncharacterized nature of the gene in question makes more indepth interpretation of the results difficult, but provides a target for future investigations into genomic controls over manganese uptake.
+It is possible that this positive correlation is due to a nutrient effect more generally, where small variations in the nutrient availability of a given plant’s soil environment contributed to more abundant nutrients (including manganese and silicon) for the plant to uptake.
+Since XRF measures multiple elements at once, we have data on other metals that do and do not interact with root exudates.
+Though our measurements are currently uncalibrated, we could extend our search to these elements, testing to see if silicon shows correlations with other elements mobilized with root exudation.
+Consistent significant positive correlations between elements that are mobilized by root carboxylates, contrasted with insignificant correlations between other elements, could lend additional support to the role of root exudation contributing to silicon content in our plants.
 
-Silicon interacts with manganese in the leaf, binding manganese to the cell wall and reducing manganese concentrations in the cell symplast (\cite{rogalla_role_2002}).
 Root exudation is most important under limited nutrients, and silicon-exudate dynamics have been best characterized in well-weathered soils (\cite{lambers_plant_2008,de_tombeur_shift_2021,de_tombeur_silicon_2021-1}).
@@
 
-Depsite our inability to detect QTLs that reliably influence silicon or manganese content, this field of study still holds immense promise.
+Phenological and morphological variation in the accessions may have contributed to differences in silicon phenotype.
+We observed noticeable variation in the size and shape of plants and their leaves.
+As silicon is primarily deposited in the epidermis, variation in specific leaf area would alter the relative contribution of the epidermis to overall leaf volume (\cite{reidinger_rapid_2012}).
+Since XRF works by measuring elemental content at the surface of a sample, rather than considering overall mass of the sample, specific leaf area variation could drive variation in leaf silicon content.
+This variation would be independent of any genes influencing silicon transport, silica cell density, or phytolith size.
+Our sampling design targeted plants that had reached maturity and we sampled all plants within two days of each other.
+At the time of harvest, some plants had begun to senesce, and the resulting transfer of nutrients out of the leaf likely altered the relative contribution of silicon content to the leaf mass.
+Controlling for both of these factors would greatly increase the complexity of a future study on this topic but may be necessary to properly identify marker-trait associations for silicon content.
+
+In the future, instead of targeting a specific amount of leaf mass, researchers could standardize their silicon measurements to the amount of leaf area, which would help control for the effects of some morphological differences driving XRF-measured silicon content.
+This approach would also focus the search to genes specifically involved in leaf-surface silicon deposition.
+Another approach would be to directly observe the silica cells and resulting phytoliths, using scanning electron microscopy, similar to \textcite{waterman_short-term_2021-1}.
+Controlling for life stage may also yield improved power.
+Over the growing season, plants can vary their silicon content in response to external cues (\cite{hartley_ecology_2016}).
+Sampling leaves at a well-defined, pre-maturity life stage could help to reduce noise introduced by phenologically driven variation in silicon content.
+
+Despite our inability to detect marker-trait associations that are reliably associated silicon or manganese content, this field of study still holds immense promise.
 We implemented a study design that we hoped would allow generalisations about the genetic regions we hoped to find.
-However, the use of multiple environments may have obscured QTLs that we may have otherwise detected in a single environment.
-The use of the diversity panel allowed for a rapid and low-cost initial search for loci asscociated with silicon and manganese content, but may have also limited our power to detect loci.
-The panel we used was a subset of the larger \textit{Aegilops tauschii} panel, and ours was selected specifically for diversity in disease-resistance genes.
-Utilizing the full panel of 384 genotypes could improve power of future studies.
+However, the use of multiple environments may have obscured marker-trail associations that we may have otherwise detected in a single environment.
+We used multiple environments to validate the effects of genetic variation independent of environmental effects on phenotype.
+This approach is widely used in crop breeding but may have imposed constraints that are too stringent for the more exploratory nature of this research.
+The use of the diversity panel allowed for a rapid and low-cost initial search for loci associated with silicon and manganese content but may have also limited our power to detect loci.
+The L2 panel we used is a subset of a larger \textit{Aegilops tauschii} diversity panel and was selected specifically for diversity in disease-resistance genes.
+
+Utilizing the full panel of 384 genotypes could improve the power of future studies.
 Alternatively, future researchers could attempt to create a mapping population.
@@
 This approach might allow for more sensitivity in the analysis by reducing the overall genetic variation between genotypes in population, highlighting the genetically driven variation in silicon content.
+
 \section{Acknowledgements}
 
-\section{Data Availability}
+We thank Chelsea Gowton and Dennus Chiu for assistance in sample harvest, preparation, and analysis.
+Dr.
+Shaun Barker and Dr.
+Brian McNulty advised and assisted with XRF analysis.
+Grace Wang, Vincent Fetterley, Sara Salad, Katherine Buchanan, Martina Clausen, Paul Fisher, and Matt Tsuruda assisted in transplanting.
+Kelly Wang, Grace Wang, and Chelsea Gowton assisted with sample harvesting.
+Dr.
+Daria Reshetniak, Paul Fisher, Lucas Friesen, Katie Pryer, Dr.
+Kinga Treder, Carly MacGregor, and Grace Wang all provided invaluable assistance with sample preparation.
 
-\printbibliography
+%\section{Data Availability}
+\clearpage
 
@@
 The assembly was scanned using the NCBI genome viewer tool.}
-        \label{tab:peaks_and_genes}
+        \label{Tab:peaks_and_genes}
         \begin{tabular}{lrrr}
@@
 
+\begin{figure}[ht]
+        \centering
+        \begin{subfigure}[b]{0.7\textwidth}
+                \centering
+                \includegraphics[width = \textwidth]{images/sienvplot.png}
+        \end{subfigure}
+
+        \begin{subfigure}[b]{0.70\textwidth}
+                \centering
+                \includegraphics[width = \textwidth]{images/mnenvplot.png}
+        \end{subfigure}
+
+        \centering
+        \caption{Mean silicon and manganese contents of \textit{Aegilops tauschii} grown in three common gardens at University of British Columbia.
+Totem Field and South Campus are two outdoor sites, while the glasshouse was indoors.
+We measures manganese and silicon using XRF.
+Error bars are 1 standard error.}
+        \label{Fig:environment_elements}
+\end{figure}
+
 \begin{figure}[h]
-        \includegraphics[scale=0.048]{images/rank_plots.png}
+        \includegraphics[width = \textwidth]{images/rank_plots.png}
         \centering
@@
         \includegraphics[width = 0.8\textwidth]{images/gwas_plots/svgtopng/mn_manhattan_plot.png}
-        \label{Fig:mn_peak_plot}
         \caption{Manhattan plots of the results of our GWAS analysis for leaf manganese content in \textit{Aegilops tauschii}.
@@
 The dotted line represents the bonferroni-corrected significance threshold.}
+        \label{Fig:mn_peak_plot}
 \end{figure}
@@
 My aim with this thesis was to advance our understanding of how silicon-based plant traits may be applied to agriculture.
-To this end, I identified two separate problems that currently limit uptake and dissemination of silicon research into the broader agricultural context: 1) limited applicability of fundamental reserach into inducded silicon uptake to crop plants, and 2) a limited understanding of the genetic basis of variation in silicon content among genotypes of a common species.
-Based on the exciting finidngs of rapid silicon accumulation in \textit{Brachypodion distachyum}, I attemped to replicate this pattern of silicon uptake in cereal crops, in the hopes that this would spur further research into silicon-based crop defenses.
+To this end, I identified two separate problems that currently limit uptake and dissemination of silicon research into the broader agricultural context: 1) limited applicability of fundamental research into induced silicon uptake to crop plants, and 2) a limited understanding of the genetic basis of variation in silicon content among genotypes of a common species.
+Based on the exciting findings of rapid silicon accumulation in \textit{Brachypodium distachyon}, I attempted to replicate this pattern of silicon uptake in cereal crops, in the hopes that this would spur further research into silicon-based crop defences.
 In the growing conditions I used, I did not observe rapid silicon accumulation, but did observe variation in silicon and phenolic compound levels among the four crop species I tested.
-We also documented the significant variation among common cultivars of these cereal species.
+We also documented the significant variation of silicon content among common cultivars of these cereal species.
 
 For my GWA study, we observed significant variation in silicon and manganese content between sites.
-Though we failed to find robsut genotype-phenotype links for either silicon or manganese, we did identify one gene that is associated with variation in leaf manganese content.
+Though we failed to find robust genotype-phenotype links for either silicon or manganese, we did identify one gene that is associated with variation in leaf manganese content.
 Additionally, by working with a previously developed diversity panel, we hope that our findings will be easily integrated into future work on this topic.
@@
 A number of limitations with my study design and experimental methods should be noted.
-Though I was ultimately unsucessful in my attempts to induce rapid silicification, I identified a number of factors that could have led to the discrepancies between previous research and my own.
+Though I was ultimately unsuccessful in my attempts to induce rapid silicification, I identified several factors that could have led to the discrepancies between previous research and my own.
 We utilized a single herbivore species in our treatments.
-Plants modulate their defensive responses based on the identify of the attacker, and it is likely that using different species or multiple species would yeild different outcomes.
+Plants modulate their defensive responses based on the identity of the attacker, and it is likely that using different species or multiple species would yield different outcomes.
 Some of the insects in my treatments did not initiate feeding, despite our best efforts to maximise the likelihood of herbivory, and this reduced our sample size for the experiment.
@@
 This limitation may have prevented rapid uptake of silicon, as there might just not have been enough in the soil to meet the demands of the plant.
-Future reserach should explicitly investigate the silicon availability of commercially available plant growth media, as nearly all plants can respond positively to silicon.
+Future research should explicitly investigate the silicon availability of commercially available plant growth media, as nearly all plants can respond positively to silicon.
 Additionally, we used a single time point to test for rapid silicon accumulation.
 It is possible that we pre-empted silicon uptake in our plants, and had we waited longer, we would have seen a clear signal of silicon uptake in response to our treatments.
-Future studies would provide immense utility to the field if they were to investigate silicon uptake at regular intervals between 6 and 48 hours post induction, as we still lack replicated data on this time interval.
+Future studies would provide immense utility to the field if they were to investigate silicon uptake at regular intervals between 6- and 48-hours post induction, as we still lack replicated data on this time interval.
 
-For the GWA study, we may have limited our ability to detect important QTLs through the use of multiple growing environments.
-Though multiple growing environments are often used to prove the durability of a given QTL through GxE interations, the highly exploratory nature of our work would have benefitted from a deeper and narrower experimental design, where we had more replicates in a single environment.
-We are left with few markers to investigate, but future work could take the genotypes we identified as having high and low silicon, and develop a mapping population using experimental crosses.
-The increased level of control this design would bring might allow for the discovery of QTLs associated with silicon content.
-Similar studies in wheat and rice have yeilded promising results, and there is no reason to believe that \textit{Aegilops tauschii} would not have similar genes that influence silicon uptake.
+For the GWA study, we may have limited our ability to detect important marker-trait associations through the use of multiple growing environments.
+Though multiple growing environments are often used to prove the durability of a given marker-trait association through genotype by environment interactions, the highly exploratory nature of our work would have benefited from a deeper and narrower experimental design, where we had more replicates in a single environment.
+We are left with few markers to investigate, but future work could take the genotypes we identified as having high and low silicon and develop a mapping population using experimental crosses.
+The increased level of control this design would bring might allow for the discovery of markers associated with silicon content.
+Similar studies in wheat and rice have yielded promising results, and there is no reason to believe that \textit{Aegilops tauschii} would not have similar genes that influence silicon uptake.
+
+For both experiments, adding soil chemistry data would help to improve our understanding of our observations.
+In both chapters, we posit that the potting mix may have low-levels of plant-available silicon.
+Without testing this directly, this remains speculative, and prevents concrete conclusions about our results.
+In the GWA study, having soil chemistry data to compare and contrast the three growing sites would help to clarify to what extent differences in phenotype are driven by the environment, genotypic effects, or an interaction of the two.
+An additional limitation of the GWA experiment is our lack of XRF calibrations for manganese and other metals.
+Adding these data would allow further regressions of silicon and other root carboxylate mobilized metals, which may help strengthen our link between silicon uptake and manganese content in the plants.
+
+\printbibliography
 

### 2023-08-23 src/R_linear_model.R

@@ -13,7 +13,7 @@ library(emmeans)
 library(car)
 library(Hmisc)
 library(lmerTest)
-library(rstanarm)
+#library(rstanarm)
 options(contrasts = c("contr.helmert", "contr.poly"))
 si_data <- as.data.frame(read.csv("data/real_data/biomass_si_data.csv"))
 si_filtered  <- si_data[si_data$isDamaged != "Undamaged",]
@@ -24,6 +24,7 @@ si_mod <- lmerTest::lmer(log(Si_ppm/10000) ~ Induction * Species + mass_g + (1|G
 si_mod <- lme4::lmer(log(Si_ppm) ~ Induction * Species + mass_g + (1|Genotype), si_filtered)
 anova(si_mod, type = 3)
 e1 <- emmeans(si_mod, ~ Species * Induction)
+
 e2 <- contrast(e1, interaction = c("pairwise", "pairwise"))
 my.aes = list(point = list(size = 4))
 e3 <- pwpp(e1, aes = my.aes) + ggplot2::theme_bw( base_size =12) + ggplot2::labs(y = "Contrast")
@@ -61,6 +62,7 @@ my_emm <- pairs(emmeans(si_mod, ~ Induction * Species))
 foo <- pwpm(my_emm, by = NULL, adjust = "bonferroni")
 xtable::xtable(foo)
 emmeans(si_mod, list(pairwise ~ Species|Induction), adjust = "tukey", type = "response")
+emmeans(si_mod, list(pairwise ~ Induction), adjust = "tukey", type = "response")
 ggplot(si_filtered, aes(x = Induction, y = Si_ppm)) +
 geom_col() +
 theme_classic(base_size = 22)

### 2023-08-23 src/cereal_induction.jl

@@ -258,7 +258,7 @@ begin
     png(biomass_si_scatter, "images/biomass_regression")
 end
 
-
+begin
 scale_vals = DataFrame()
 ads_centered = rescalecols(df=analysis_data, collist=[:Si_ppm], centers = scale_vals)
 full_model = fullcenteredglm(df=analysis_data)
@@ -285,11 +285,40 @@ test_df = DataFrame(var1 = repeat(["A", "B", "C"], inner = 3, outer =3), var2 = 
 tplot = @df test_df groupedboxplot(:var1, :yvals, group = :var2)
 
 filtered_analysis_data = filter(row -> ismissing(row.isDamaged) || row.isDamaged  ≠ "Undamaged", analysis_data)
+end
+#Compare Induction Treatments in Graph
 
+g_by_induction = groupby(filtered_analysis_data, [:Induction])
+si_by_induction = combine(g_by_induction, [:Si_ppm, :Induction] => ((si, in) -> (grpMEANsi = mean(si/10000), grpSEsi = standarderror(si/10000), ind = first(in))) => AsTable)
+sort!(si_by_induction, [:ind])
+begin
+    induction_only_plot = plot(si_by_induction.Induction, 
+    si_by_induction.grpMEANsi .± si_by_induction.grpSEsi,
+    seriestype = :scatter,
+    markersize = 10,
+    color = :black,
+    legend = false
+    )
+
+    
+
+    #plot!(xticks = ([2.5, 6.5, 10.5], unique(si_induction_summary.ind)))
+    plot!(xtickfontsize =16 ,xlabelfontsize = 20, ytickfontsize = 16, ylabelfontsize = 20, legendfontsize = 16)
+    plot!(left_margin=5mm, bottom_margin=2mm, right_margin = 10mm)
+    #plot!(xrotation =45)
+
+    plot!(xlabel = "Treatment Type", ylabel = "Leaf Silicon Content (%)")
+    plot!(grid=false)
+    plot!(size = (800,600), dpi = 600)
+end
+savefig(induction_only_plot, "./manuscript/images/induction_only_bytrt.png")
+begin
 g_analysisdata = groupby(filtered_analysis_data, [:Induction, :Species])
[…]
